Wound-Induced Syncytia Outpace Mononucleate Neighbors during Drosophila Wound Repair

Version 1

Abstract

Abstract

All organisms have evolved to respond to injury. Cell behaviors like proliferation, migration, and invasion replace missing cells and close wounds. However, the role of other wound-induced cell behaviors is not understood, including the formation of syncytia (multinucleated cells). Wound-induced epithelial syncytia were first reported around puncture wounds in post-mitotic Drosophila epidermal tissues, but have more recently been reported in mitotically competent tissues such as the Drosophila pupal epidermis and zebrafish epicardium. The presence of wound-induced syncytia in mitotically active tissues suggests that syncytia offer adaptive benefits, but it is unknown what those benefits are. Here, we use in vivo live imaging to analyze wound-induced syncytia in mitotically competent Drosophila pupae. We find that almost half the epithelial cells near a wound fuse to form large syncytia. These syncytia use several routes to speed wound repair: they outpace diploid cells to complete wound closure; they reduce cell intercalation during wound closure; and they pool the resources of their component cells to concentrate them toward the wound. In addition to wound healing, these properties of syncytia are likely to contribute to their roles in development and pathology.

Introduction

Injury is a constant reality of life, and survival requires all organisms to repair wounds. Wound-induced cell behaviors like proliferation, migration, and invasion replace missing cells and close wounds [1, 2]. Other cell behaviors are induced around wounds, but their contribution to wound healing is not well understood, e.g., the fusion of cells into syncytia. Syncytia are one type of polyploid cell, and it is generally appreciated that increases in ploidy – number of genomes per cell – is a common response of post-mitotic tissues and cells to injury [3-5]. Wound-induced epithelial syncytia were first observed around epidermal puncture wounds in Drosophila larvae and adults [3, 6], consistent with the idea of polyploidy induction in non-proliferative cells. However, recent studies have observed syncytia around wounds in mitotically competent tissues: around laser-ablation wounds in Drosophila pupal epidermis [7] and in zebrafish epicardium damaged by endotoxin, microdissection, or laser ablation [1]. Further, injury associated with the surgical implantation of biomaterials can cause immune cells to fuse into multinucleated giant cells, which are associated with rejection [8]. Similarly, injury induces bone marrow-derived cells to fuse with various somatic cells to promote repair [9-12]. The many instances of syncytia being induced by wounds raise the possibility that syncytia offer an adaptive benefit. It is not clear, however, what that benefit is.

Syncytia can form either by endomitosis – mitosis without cytokinesis – or by cell-cell fusion. Such fusion is widely observed throughout development in both vertebrates and invertebrates: for example myoblasts fuse into muscles [13, 14], and fusions occur in the lineages of the C. elegans hypodermis [15] as well as vertebrate trophoblast [16], myoblast, and osteoclast [17]. Cell fusions are also observed in disease: pathogen-induced epithelial fusion allows spreading of many viruses including human respiratory syncytial virus and SARS-CoV-2 [18]; and the fusion of cancer cells with bone-marrow derived cells is implicated in metastasis [19].

Here, we use live imaging and clonal analysis to understand the behavior of syncytia following wounding in the Drosophila pupal notum. The unwounded notum is a monolayer epithelium composed of mononuclear diploid cells that are mitotically competent. Nonetheless, during the first several hours after wounding, many of the surrounding cells fuse to form giant syncytia. Some fusions are obvious with apical borders breaking down, while others appear only as shrinking of a cell’s apical surface. All together, fusion is a common fate of cells near wounds: about half the cells fuse to form syncytia within 70 µm of a wound with 30 µm radius. Compared to their mononuclear neighbors, syncytia have dramatically improved wound-repair abilities: they outpace smaller mononuclear cells to the leading edge, they limit the need to negotiate cell intercalations as the wound closes, and they mobilize and transport actin from distal cells to reinforce the wound margin.

Results

A mitotic tissue utilizes cell-cell fusions during wound repair

The pupal notum is a monolayer columnar epithelium composed of diploid cells that undergo regular mitotic cycles [20] (Fig. S1A-C). To analyze cell behaviors around wounds, we live-imaged after laser ablation. Epithelial cell borders were labeled by the adherens junction protein p120ctnRFP [21] and nuclei were labeled with histone His2GFP. Two hours after wounding, we observed prominent syncytial cells around the wound (Fig.1A, Movie S1). Some syncytia appeared to contain over a dozen nuclei within epithelial borders. For both syncytial and mononuclear cells, it was difficult to assign nuclei precisely to cell borders because notum epithelial cells are not rectangular and are not arranged at right angles with respect to the surface; in 2-D projections, a nucleus was frequently observed outside the cell’s apical border (Fig. S1D). Accordingly, using apical area and nuclear density, we estimated the number of nuclei within the 3 largest syncytia in different wounds. The number of nuclei in these syncytia increased over time: at 1 h after wounding the three largest syncytia contained an average 3-13 nuclei, and 2 h after wounding they nearly doubled to 6-20 nuclei (Fig. 1B). Interestingly, larger wounds generated syncytia with larger apical areas and more nuclei, proportional to wound size, suggesting syncytium formation is a dynamic and scalable response to injury (Fig. 1 C-E).

Wounds induce epithelial syncytia via cell fusion.

A) Syncytia form within 2 h post wounding, evident by the clustering of multiple nuclei within cell borders. B) The number of nuclei per syncytia increases over time after wounding. Number of nuclei was estimated based on area and nuclear density (see text) for the 3 largest syncytia of 3 different wounds, mean and SD. C-D) Larger wounds generate larger syncytia. Images are at 3 h post wounding, single syncytium outlined. E) Apical area of three largest syncytia is proportional to initial wound size, mean and SD. F) A time course of six cells fusing within 30 min after wounding. Apical borders are lost (white arrowhead) as syncytia form. Original cells are numbered. G) All borders lost to cell fusion (white) mapped to cells in the first frame after wounding. The leading edge of wound closure will form at dashed line; cells within the shaded area were damaged by the wound and will be dismantled. H) Distance from the wound center vs time for all border breakdowns in 3 wounds. Each symbol represents a cell border. Leading-edge locations indicated by solid lines. I) Cytoplasmic GFP is expressed in cell 1 before wounding and mixes with neighboring cells 2-4 by 2 min after wounding. Cytoplasmic sharing is followed by the lagging fusion indicator of visible border breakdown (white arrow). The fates of cells 3 and 4 are shown at later times in Fig. 2D. Maximum intensity projections in A, C-D, G; single Z slices in F, I. Scales: A-D,G = 20 µm, F,I = 10 µm. W and red star indicate wound.

Live imaging of cells after wounding revealed the gradual loss of p120ctn between some epithelial cells followed by syncytial rounding (Fig. 1F), suggesting epithelial cell fusion. As p120ctn was lost, the epithelial cadherin E-cadherin was also lost (Fig. S1E-H) indicating the disassembly of adherens junctions between the cells. We called this phenomenon “border breakdown”. Border breakdowns were found spatially clustered in the first three to four rows of cells, 30-50 μm from the wound center, and they occurred primarily within the first hour after wounding, some within 10 minutes after wounding (Fig 1G,H). To test whether border breakdowns represented cell fusions, and not loss of adherens junctions caused by epithelial to mesenchymal transitions, we analyzed cytoplasmic mixing. Individual GFP-labeled cells were generated at random locations using the flip-out Gal4 technique [22]. Before wounding, the level of cytoplasmic GFP fluorescence was stable yet exhibited cell-to-cell variability, allowing some differentiation of cells by intensity. Minutes after a nearby laser ablation, GFP was observed to diffuse from labeled cells into neighboring unlabeled cells. GFP mixing between two cells was followed by the eventual loss of their shared p120ctn-labeled cell border, confirming that border breakdown is indeed cell fusion, but that cytoplasmic mixing occurs more than 10 minutes before the border breakdown is first observed to start (Fig 1I, Movie S2). Border breakdowns between labeled and unlabeled cells were always preceded by GFP mixing (n=11). Thus, epithelial fusion is a rapid local response to wounding.

Cell shrinking is a second form of cell-cell fusion occurring later during wound closure

Because border breakdowns occurred mostly within 1 h after wounding, it was unclear how syncytia grew between 1-2 h after wounding. However, an unexpected cell behavior was frequently observed during this time: the apical area of diploid epithelial cells shrank until they disappeared, which we termed “shrinking cells” (Fig. 2A). Although we first expected that shrinking cells were extruding, when we tracked the nuclei of shrinking cells, we were surprised that all nuclei moved laterally to join neighboring syncytia (n=7, Fig. 2B)), suggesting that shrinking cells are fusing cells. To better understand this behavior, we analyzed individual GFP-labeled cells. Like with border breakdowns, GFP mixing preceded the initiation of cell shrinking, although by a longer interval of one or more hours (Fig. 2C, Movie S3). Additionally, X-Z projections through GFP labeled shrinking cells are consistent with cytoplasm moving into neighboring syncytia (Fig. 2D). Shrinking cells were distributed similarly around wounds as border breakdowns but occurred later and were more numerous (Fig. 2E,F). Thus, both border breakdown and cell shrinking are indicative of cell fusion.

Cell fusion often appears as cell shrinking.

A) An epithelial cell (yellow arrow) shrinks in the epithelial plane after wounding. (B) Same sample as A, showing the shrinking cell’s nucleus (asterisk) entering a neighboring syncytium, outlined in Biv. C) Cell 1 expresses actin-GFP before wounding. After wounding, cell 1 fuses with cells 2-6 evidenced by GFP sharing, then cell 1 shrinks. D) Two cells shrink in the X-Y plane (white and yellow arrows in Di-Diii). X-Z projections are shown for each (Div-Dvi for the top yellow line; Dvii-Dix for the lower white line). Rather than extrude, the shrinking cells’ cytoplasm moves to the right, to join with the neighboring wound-proximal syncytium, via basal connections. These frames are a continuation of the sample shown in Fig. 1I. E) All shrinking cells mapped to first frame after wounding. F) All shrinking cells and border breakdowns were tracked in 3 wounds. Border breakdown fusions happen sooner after wounding than shrinking fusions. A, B, C, Di - iii, E show maximum intensity projections. Div − ix show X-Z projections. Scales: A, B, C, Di-iii, E = 10 µm; Div - ix = 5 µm.

To determine what percentage of cells around a wound will fuse, we analyzed over 100 randomly-labeled single GFP cells within the radius of observed fusion (80 µm), tracking them for 6.5 h to assess their fate (Fig. 3A,B): a full quarter of the cells fused (25%), sharing GFP before breaking down borders or shrinking; 67% persisted as diploid cells, most with stable GFP, but infrequently (n=3) with GFP mixing and no subsequent cell fusion; the remaining 7% could not be tracked (Fig. S2A). Fusing cells were strongly skewed toward the center of the wound: within 70 µm, about half the cells (47%) underwent fusion (Fig. 3A, right). Fusion continued for over 300 min after wounding (Fig. S2B,C) whereas border breakdown is concentrated in the first hour after wounding (Fig. 1H, Fig. S2C). Shrinking accounts for more than half of all fusing cells (Fig. 3C), and the spatial distribution of fusing cells that shrank vs. lost borders was similar (compare Figs. 1G and 2E). However, shrinking began later than border breakdowns, continued for several hours after wounding (Figs. 2F, S2C), and took longer to complete (Fig. S2D). Thus, shrinking fusion accounted for continuing syncytial growth after border breakdowns subsided. Further, cell fusion is a persistent behavior over the course of wound closure.

Half the cells near the wound fuse to form syncytia, demonstrated by tracking individual cell fates.

A) All GFP labeled cells in the region of fusion (80 µm) were tracked in 5 wounds over 6.5 h after wounding to determine frequency of fusion (GFP mixing) or persistence (no GFP mixing). Untrackable cells lost GFP, see Fig. S2A. B) Fusion is common within 70 µm. C) Shrinking-fusion and border-breakdown fusion occur at similar distances from the wound.

Syncytia outcompete mononucleate cells at the leading edge of repair

Because about half the cells fused to form syncytia around wounds, we were able to compare the behavior of syncytial to non-syncytial cells. Comparing cell behaviors within the same wound provided a well-controlled environment for assessing how syncytia contribute to wound closure. Using live imaging, we observed that syncytia frequently overtook unfused cells as they moved toward the wound. Figure 4A shows the fusion of seven cells, with non-fusing GFP-labeled cells both distal and proximal to the fusing cells (panel 4Aii, cells 1 and 4 respectively). Later the syncytium advanced beyond both unfused cells toward the wound; it even reached around and past wound-proximal cell 4 to extend the leading edge toward the wound Fig. 4Aiii). This behavior was evident even without GFP labeling: Fig. 4B-C show a group of cells at the leading edge of the wound, recognized by the lack of p120ctn. At 90 min after wounding, this area of the leading edge is composed of three mononuclear cells (the middle one outlined in orange and white) flanked on either side by syncytia (outlined in yellow). In panel 4Biii, both syncytia have pushed out the mononuclear cells from the leading edge. Thus, it appeared that syncytia were able to outpace mononuclear cells toward the leading edge.

Syncytia outpace smaller cells.

A) Before wounding, two clusters of cells are labeled with GFP, 1-2 and 3-4 (white numbers). After wounding, cells 2 and 3 fuse with neighbors 5-9 to form a syncytium (dashed leading edge in Aiii), which advances toward the wound, passing unfused cell 4 (arrow). B-C) Unfused cell (outlined in orange and white) with corresponding nucleus (asterisk in C) is replaced at leading edge by neighboring syncytia (S and yellow outline, yellow arrows). D) All small cells are excluded from leading edge by syncytia well before the wound closes 20-160 min later. E) Images of sample 1 from the graph D. Unfused cells (arrows) were tracked over the course of wound closure and mapped back to the 30 min timepoint. F) At 120 min, the last unfused cell is ejected from leading edge (arrow). The wound closed at 140 min. G) Percent of leading-edge perimeter occupied by syncytia increases to 100%. Images are single Z slices in Ai-Aii and maximum intensity projections in Aiii-F. Scale: A,E-F = 20 µm, B-C = 10 µm.

Further, we noticed that by several hours after wounding, the leading edge of the wound was occupied primarily by syncytia. To investigate how syncytia came to occupy this position at the front lines of wound healing, we expressed MyoIIGFP/Zip-GFP, which along with actin forms the contractile purse string and makes the leading edge visible, along with p120ctnRFP to label cell borders. We analyzed the persistence of all mononuclear and syncytia cells at the leading edge over the course of closure for three wounds, starting when the leading edge was first visible (Fig. 4D-F). As soon as the leading edge formed 30 min after wounding, about 75% of the perimeter was occupied by syncytia (Fig. 4G); the rest of the perimeter was occupied by 10-13 small cells (Fig. 4D). As the wound closed, the syncytia became larger and displaced all the small cells, with the last small cell removed from the leading edge well before closure, which occurred 20-160 min after removal of the last small cell (Fig. 4 D,F). No small cell persisted at the leading edge through wound closure, indicating that syncytia are better able to close wounds than unfused cells.

Syncytia reduce intercalations and move actin to the wound

Why are syncytia better at closing wounds? To address this question, we considered the geometry of fusion. We observed that fusions sometimes occur between two adjacent cells equidistant from the wound, as diagramed in the top panel of Fig. 5A and exemplified in Fig. 1 Iiii-Iiv (cells 1,2) We call this type radial border breakdown, and it produces a syncytium elongated along the wound edge. Alternatively, fusion may occur between adjacent cells at different distances from the wound, as drawn in the lower panel of Fig. 5A and exemplified in Fig. 1F (cells 3,4). We call this type tangential border breakdown, and it produces a spoke-like syncytium pointing into the wound. To analyze the frequency and timing of these two different axes of fusion, we calculated the angles of all lost borders for all border breakdowns in four wounds, binning them into radial and tangential fusions. (The angle of the axis of fusion can be determined for border breakdowns; the axis of fusion for shrinking cells is difficult to analyze, although the geometry of the starting cells must be similar.) In four wounds, 235 border breakdowns were identified: fewer radial borders broke down than tangential borders (39 vs 196), and radial borders broke down a bit later (Fig. 5B, Fig. S3B). However, the two types were similarly distributed around the wounds (Fig. S3A).

Radial border fusions reduce the number of wound proximal intercalations.

A) Illustration of tangential vs radial border breakdown. B) More tangential than radial borders break down after wounding. C) Radial border fusions reduce intercalation at a wound. D) Quantification of intercalations and fusions around the wounds of Fig. 4D,G.

These fusion orientations provide different benefits for wound repair. Radial border fusions reduce the requirement for cell intercalation: as a wound closes, fewer cells can occupy the leading edge, necessitating a rearrangement of cell adhesions to allow cell intercalation, as depicted in the top of Fig. 5C. Indeed, intercalation is known to be a rate-limiting step of wound repair [23]. However, fusion of radial borders remove the need for intercalation, as shown in the bottom of Fig. 5C. To consider the contribution of radial border fusions to wound closure, we analyzed the three wounds from Fig. 4D-G and asked how many leading-edge cells fused radially vs. intercalated: 16-41% of cells removed from the leading edge were removed through fusion, reducing the burden of intercalation substantially (Fig. 5D). Interestingly, the larger the percentage of cells that fused rather than intercalated, the faster the wound closed (Fig. 5D). Thus, one mechanism syncytia use to promote wound closure is radial fusion.

In contrast, tangential border fusions might provide a way for cellular resources that would be trapped in distal cells to move toward the wound to contribute to closure. To test this hypothesis, we generated small flip-out clones expressing actin-GFP and wounded such that unlabeled cells intervened between the labeled cell and the leading edge (Fig. 6). We envisioned that, upon fusion, actin would equilibrate throughout the newly fused cells, pooling their resources, and indeed we did observe actin-GFP to equilibrate between cells soon after fusing. For syncytia that did not have access to the leading edge (n =2), actin-GFP levels remained uniform (Fig. 6A,B). In contrast, syncytia with access to the leading edge (n = 10) first uniformly distributed actin (Fig. 6Cii), but once the leading edge was contacted, they redistributed actin-GFP to it (Fig. 6C-D). Kymographs of actin-GFP confirm that regardless of location, actin equilibrates between fusing cells within 5 minutes (Fig. 6Biii, Div); however, nearly all actin that originated in the distal cell is redistributed to the leading edge in a syncytium positioned there (Fig. 6Div). In one striking instance, actin-GFP appeared to travel through three cells to arrive at the leading edge from its initial location three rows back (Fig. 6E-G, Movie S4). Importantly, although actin is labeled from only one of the fusing cells, it likely represents the total actin from all fusing cells, explaining why syncytia are better able to occupy the leading edge. We envision that other resources would also be concentrated as needed by syncytia. We conclude that syncytia formed by cell fusion are able to outcompete their mononuclear neighbors by reducing intercalation and by concentrating the collected resources of many cells.

Syncytia concentrate pooled resources at the leading edge.

Random scattered cells expressing Actin-GFP were generated by heat-shock mediated flip-out expression of Gal4. A-B) Labeled actin is expressed in cell 1 before wounding (Ai). Actin-GFP rapidly equilibrates between cells 1 and 2 by 28 minutes after wounding, demonstrating cytoplasmic fusion. At 180 min, cell 2 shrunk into cell 1. The resulting syncytiuym has no access to the leading edge, and actin remains equilibrated, as shown in the kymograph (Biii) generated from actin-GFP intensity over time at the yellow line in Bii. C-D) Labeled actin is expressed in cell 1 before wounding (Ci) and equilibrates between cells 1 and 2 by 6 minutes after wounding, demonstrating cytoplasmic fusion (Cii). The resulting syncytium contacts the leading edge, and by 28 min after wounding actin from cell 1 is redistributed to the wound margin by 28 min after wounding, as shown in the kymograph (Div) of actin intensity over time at the yellow line in Diii. At 55 min after wounding, cell 1 shrunk into cell 2. E-F) Before wounding, actin-GFP in cell 1 is three cells away from the future leading edge. After wounding, fusion of cells 1-4 allows actin-GFP to move to and accumulate at the leading-edge actin cable. Border breakdown occurs between cells 1-2 (Ev) which later shrink into cells 3,4 (see Movie S4). G) Mean profile plot of actin-GFP comparing the syncytia in Fv at 30 min after wounding (dark line) with the cells in Fi before wounding (dotted line) demonstrating that nearly all actin-GFP has been redistributed to the leading edge from its starting position 20-30 µm away. Single Z slices for Ei - ii, Fi - ii; maximum intensity projections for A-D, Eiii - v, Fiii - v. Scale A-Bii, Ci-Diii = 5 µm, E-F = 10 µm.

Discussion

Previous work established that polyploid cells – both cells with multiple nuclei and cells with single enlarged nuclei – are important for closing epithelial wounds in post-mitotic tissues [3]. Here we report that in a mitotic tissue, the epithelial monolayer of the pupal notum, syncytia form around wounds by cell fusion at a remarkably high rate, with almost half the epithelial cells within 5 cells of the wound fusing with neighbors over the course of hours. Syncytial size increases with wound size, indicating syncytia formation is a dynamic and scalable response to wounding. Syncytia form via two temporally distinct processes, a rapid breakdown of epithelial borders within 40 min of wounding, and later cell shrinking which persists 30 min to 2 h after wounding. We confirmed that both border breakdown and cell shrinking are cell fusion events, rather than epithelial to mesenchymal transitions or extrusions, by generating small clones of cells labeled with cytoplasmic GFP and observing the diffusion of GFP from a source cell into neighboring cells after wounding, indicating cell fusion. Strikingly, although the resulting syncytia are fewer in number than persisting diploid cells, they completely displace unfused cells at the leading edge of the wound such that wounds are closed entirely by syncytia.

We identified several factors that endow syncytia with wound closing abilities. Live imaging indicated that syncytia are faster than smaller cells at extending toward the wound, and once there, they maintain their positions at the leading edge, forcing out smaller cells. Epithelial wounds close by the cinching of an actin cable, sometimes called the purse string [24-27]; as the cable tightens and shortens, the leading edge becomes smaller, with room for fewer and fewer cells. Cells are removed from the leading edge by intercalation, a process that requires them to remodel their adhesions. A previous study found that intercalation is the rate-limiting step of wound closure; further, the greater the tension in cellular adhesions, the slower the wound closure, a finding both predicted by computational modeling and verified experimentally. Thus, fluidity promotes closure [23]. By fusing, cells reduce the need for adhesion remodeling and intercalation as the leading edge becomes smaller. Moreover, the resulting larger cell has significantly more fluidity and less epithelial tension than the diploid progenitors, as recently demonstrated in syncytia formed by age-induced epidermal cell fusion [28].

In addition to promoting wound closure by reducing the burden of intercalation, syncytia can redirect toward the wound cellular resources that would be trapped in individual diploid cells. We demonstrated this ability by visualizing labeled actin from one cell as it fused with an unlabeled cell. As expected after fusion, labeled actin diffuses and equilibrates throughout the new large cell. Remarkably however, when the syncytium is in contact with the leading edge, labeled actin from the distal cell is concentrated at the leading edge, even if the original cell is several cells away from the leading edge. This result suggests that syncytia can apply up to N times more actin to the leading edge, where N represents the number of cells that fused; considering that we observed syncytia with dozens of nuclei, this could represent a significant enhancement of actin at the leading edge. Increased actin can explain the ability of syncytia to outcompete diploid cells at the leading edge. In addition to the actin purse-string, actin also forms filopodia and lamellipodia important for migration, offering an explanation for how syncytia extend more quickly than diploid neighbors toward the wound, and these structures also participate in closing the wound [27, 29]. Presumably, other resources such as mitochondria and ribosomes could also be pooled and concentrated by syncytia at cellular locations where they promote wound healing. It is known that mitochondrial fragmentation promotes the repair of single-cell wounds, with more fragmentation causing faster repair [30]. Further, fragmentation is localized to the site of cellular injury [31], raising the possibility that syncytia may increase the local concentration of fragmented mitochondria. Ribosomes have been observed to be localized in wounded tissue, accumulating at the tips of severed neurons [32], and it is possible that syncytia may increase the localized pool of ribosomes in epidermal wounds. By pooling the cellular resources of component cells, syncytia may also allow lethally damaged cells to survive by providing them with needed survival factors originating in cells further from the wound. Thus, the concept of resource sharing that we demonstrate with actin may have ramifications for many resources.

We observed two cellular behaviors that attended fusion, border breakdowns and cell shrinking. Border breakdowns occurred sooner after wounding than shrinking and appeared to be a faster process, as shrinking lasted for hours. These differences in appearance and timing suggest that the mechanisms behind these fusions may be somewhat independent. For shrinking cells, we envision that the original site of cytoplasmic fusion, the fusion pore, occurs in the basolateral membrane, whereas border-breakdown fusion pores are nearer the apical adherens junctions observed to break down. The adherens junctions would be under more tension than the basolateral membranes, and this may explain why fusion proceeds more quickly there; we recently found that although epithelial tension drops after laser wounding, it is restored within about 10 min [33], consistent with the timing of apical border breakdown.

It is unclear what triggers either type of wound-induced epithelial fusion. Many developmentally programmed cell fusions are mediated by fusogens, cell surface proteins that bring opposing membranes into close contact with each other, as in the C. elegans hypodermis [34-40]. For other cell fusions, the fusogen is elusive and may not exist, for example in Drosophila myoblast fusions, which occur when a fusion competent myoblast generates actin-rich podosome-like membrane protrusions that invade a founder cell [41-45], but we did not observe these structures in fusing epidermal cells. Wound-induced fusion may be a response to the plasma membrane damage that occurs around wounds; indeed, plasma membrane damage has been documented around both laser wounds and puncture wounds [46-48].

Polyploidy as a wound response has begun to get increased recognition. In adult Drosophila, epithelial puncture wounds are repaired by both endoreplication and syncytia formation [3, 49-53]. In the zebrafish epicardium, genetic ablation is repaired by a wavefront of multinucleated polyploid cells formed by endomitosis, and these lead diploid cells to encompass the heart [1]. In adult mammalian cardiomyocytes, polyploidy may be an adaptive response to maintain growth after the cardiomyocytes lose their ability to complete mitosis. Mouse cornea endothelial cells endoreplicate to increase polyploidy to restore tissue ploidy following genetic ablation [52]. Mammalian hepatocytes are known to become increasingly polyploid with age [54, 55] and in response to various types of injury and disease [56-61]. All mechanisms that promote polyploidy – fusion, endomitosis, endoreplication – result in larger cells with the potential to localize more resources; of these, fusion would act the fastest after wounding because there is no need for DNA replication. Interestingly, in Drosophila embryos, wounds induce the surrounding cells to become larger by increasing their volume alone and not their ploidy [62], suggesting that simply an increase in size is important. Although many examples of wound-induced polyploidy exist, it is still likely to be an underreported phenomenon, as endpoint analysis might miss a transient polyploid response to injury; live imaging is the surest way to identify a polyploid wounding response.

Another polyploid response to injury can occur after surgical implantation of biomaterials for the purpose of guiding regeneration, as reviewed previously [8]. In some cases, only mononuclear cells of the immune system respond to the implant, and in these cases the biomaterial is integrated into the body; in other cases, the implant triggers the fusion of immune cells into multinucleated giant cells, and in these cases the material is degraded and rejected. These multinucleated giant cells seem to share properties with the syncytia of the pupal notum, as they are formed by fusion in response to an environmental trigger and they have an aggressive ability to protect the animal in response to wounding. As these studies highlight, understanding the formation, maintenance, and regulation of polyploid cells may improve our ability to successfully implant biomaterials to aid tissue regeneration.

It is often noted that wound responses are similar to cancer cell behaviors. This similarity extends to wound-induced syncytia and their counterparts, polyploid giant cancer cells, as both cell types are highly aggressive and invasive. Chemotherapeutics induce the formation of polyploid giant cancer cells [63-66], and some studies indicate that they can form through cell-cell fusion in tumors [67-71]. Once formed polyploid giant cancer cells are hypothesized to escape further chemotherapy treatments due to increased resistance to genotoxic stress [72]. These polyploid giant cancer cells and their progeny also exhibit increased migration and invasion potential [73, 74]. The parallels between the behaviors of polyploid giant cancer cells and the wound-induced syncytia of the pupal notum highlight the importance of understanding wound induced syncytia formation in a highly reproducible system, as a basic understanding of how these cells form in the Drosophila notum could inform how they become dysregulated in cancer.

Supplemental Movie 1: Syncytial cells form after wounding in the Drosophila pupal notum. Epithelial cell borders in red (p120ctnRFP) and nuclei in green (HistoneGFP). White box on first frame denotes field of view in Fig. 1A. Movie begins before wounding and extends to 2 h after wounding.

Supplemental Movie 2: GFP mixing precedes border breakdown after wounding. Arrow in first frame points to cell border between cells that will fuse after wounding. GFP diffusion into the unlabeled cell precedes visible border breakdown. w, wound region. Cell borders are labeled with p120ctnRFP. Movie begins before wounding and extends to 2 h 10 min after wounding. Same cells as Fig. 1I and 2D.

Supplemental Movie 3: Shrinking cells contribute to syncytia. Arrow in first frame points to an individual cell labeled with Actin-GFP that fuses by shrinking after wounding. This cell contributes its actin-GFP to neighboring cells minutes after wounding then shrinks much later; shrinking is first evident about 1.5 h after wounding and is nearly complete by 3.5 h after wounding. w, wound region. Cell borders are labeled with p120ctnRFP. Same cells as Fig. 2C.

Supplemental Movie 4: Syncytia pool actin and concentrate it at the leading edge of repair. An individual cell labeled with actin-GFP fuses with wound proximal cells and contributes its actin to the leading edge of the syncytium. w, wound region. The original source cell and its immediate neighbor go on to shrink into wound proximal cells. Cell borders are labeled with p120ctnRFP. Same cells as Fig. 6E,F.

Supporting information

eLife Assessment

This work addresses an important biological question: what is the cellular basis of wound healing? Using the Drosophila pupal notum as a model, the paper provides an elegant, thorough, descriptive characterisation of syncytia-driven wound closure using state-of-the-art confocal live imaging of the pupal notum. The authors meticulously characterize the cell-cell fusion events during wound healing, but without any mechanisms to inhibit cell fusion, it is incomplete, since it remains unclear whether cell fusion is required or not for speeding wound healing and/ or increasing the level of actin resources at the leading edge.

Reviewer #1 (Public Review):

Summary:

This study aims to understand how cell fusion contributes to wound healing using a laser-induced injury in the notum epithelium of a developing fruit fly. The authors meticulously characterize the epithelial fusion events using a live imaging approach and report that syncytia arise by 'border breakdown' and 'cell shrinking'. The syncytial epithelial cells also appear to outcompete mononucleated cells and preferentially dissolve their tangential borders, which correlates with the accumulation of actin at the leading edge.

Strengths:

The strength of this study is the authors' live imaging approach to capture these dynamic fusion events that are a fundamental, yet poorly understood biological process.

Weaknesses:

A major weakness is that all the authors' conclusions are based on descriptive studies, in which the role of cell fusion is not directly tested. This is particularly important because other models of wound-induced polyploidization have demonstrated that another cytoskeletal protein, myosin, was upregulated and dependent on endoreplication, and not cell fusion. Therefore it remains unclear to what extent cell fusion, endoreplication, or both are required to outcompete mononucleated cells as well as pool actin as described in this study.

Reviewer #2 (Public Review):

Summary:

Overall, this study provides a thorough description of the formation of syncytia following wounding of the proliferation-competent diploid epithelium of the pupal notum. While this phenomenon has already been described briefly for this particular tissue by the Galko lab in Wang et al 2015, the authors provide a much more detailed description and characterisation of the process providing some novel insights (radial versus tangential border breakdown, cell shrinkage, timings, syncytia outcompeting mononucleated cells, etc.).

Strengths:

This paper provides an elegant, thorough, descriptive characterisation of syncytia-driven wound closure using state-of-the-art confocal live imaging of the pupal notum. The authors show that laser-induced wounding of this diploid, proliferation-competent epithelium results in the formation of syncytia of various sizes in the first few cell rows around the wound edge, which progressively become bigger as healing proceeds. This results in ~50% of cells becoming part of these syncytia. The cell fusion events were convincingly demonstrated by showing the disappearance of p120ctnRFP and E-Cadherin-GFP from cell-cell borders as well as cytoplasmic GFP mixing of GFP-positive cells with a GFP-negative cell.

Apart from cell-cell fusion by border breakdown that mostly happens in the first 2h following wounding, the authors also found that at later stages of wound healing cell shrinkage following cytoplasmic mixing contributed to sycytia formation.

Next, the authors provided some convincing evidence that syncytia outcompete mononuclear cells for being positioned in the first cell row around the wound.

The authors then show that radial border breakdown occurs much less frequently than tangential border breakdown. They suggest that radial border breakdown reduces the requirement for cell-cell intercalations. They also hypothesise that tangential border breakdown might allow fused cells to share resources and provide more resources to be used near the wound edge, e.g. for actomyosin cable formation. To test this, the authors generate single-cell clones that overexpress Actin-GFP. They then show convincingly how a single Actin-GFP-positive cell in the second cell row fuses with one GFP-negative cell in the first cell row. The Actin-GFP signal then spreads in the fused cell and labels some previously unlabelled actin-rich structure near the wound edge which most likely is the actomyosin cable. This provides some evidence for resource sharing by cytoplasmic mixing following fusion.

Weaknesses:

The authors provide some convincing evidence that syncytia outcompete mononuclear cells for being positioned in the first cell row around the wound. The authors suggest that the syncytial cells might be better able to close the wound. However, some genetic studies would need to be done to establish this more convincingly. E.g. Could the authors genetically block syncytia formation and then show that these wounds now heal slower?

The authors suggest that radial border breakdown reduces the requirement for cell intercalation. While this might be true it also raises the question of how the various syncytia facing the wound border change shape to allow the shrinkage of the first cell row over time to allow wound closure. None of the four movies included in the study shows the whole wound healing process until the later stages, making it hard to assess this. It would be good to include one such movie showing the syncytia in the whole wound and comment on this point.

The authors hypothesise that tangential border breakdown might allow fused cells to share resources and provide more resources to be used near the wound edge, e.g. for actomyosin cable formation. They show convincingly through the fusion of a single Actin-GFP-positive cell in the second cell row with a GFP-negative cell in the first cell row that Actin-GFP spreads in the fused cell and labels the previously unlabelled actomyosin cable. While the hypothesis of resource sharing to improve healing is intriguing and makes sense, this experiment doesn't necessarily prove the benefit of resource sharing. It does show cytoplasmic mixing following fusion, now allowing the GFP-labelled actin to diffuse and be incorporated into the actomyosin cable. In a wild-type condition, fusion would not increase the total concentration of resources, although it would increase the total amount of resources within this bigger fused cell. The question is whether resource sharing without increasing the protein concentration is beneficial and increases the efficiency of certain wound healing mechanisms. There might be a benefit of cell fusion, if for example certain resources were only present in limited amounts or if protein transport could increase the concentration locally. To provide better evidence for the hypothesis that resource sharing improves wound healing, maybe the authors could look at the actomyosin cable in a wounded epithelium (such as in Figure 4E, F), in which all cells express MyoII-GFP. The authors could compare the average intensity of the actomyosin cable at the wound edge in mononucleated cells versus in syncytia. If resource sharing is indeed beneficial, it might be that the actomyosin cable is stronger/brighter in syncytia or it forms quicker.

The biggest limitation of this study is that the authors don't address how the formation of these syncytia is regulated. While the manuscript in its current form provides some valuable new insights into syncytial-driven wound closure, it would be much more informative if it also provided some mechanistic details. The authors could test if some of the mechanisms shown to regulate syncytial formation in other types of syncytia-driven wound healing are also involved here. E.g. Yorkie was shown to negatively regulate cell fusion in adult syncytial-driven wound closure (Losick et al 2013). The authors could test for the effect of Yorkie-RNAi in the epithelium on wound closure and syncytia formation. Expression of the dominant negative RacN17 also blocked cell fusion in adult syncytial-driven wound closure (Losick et al 2013).

Moreover, JNK activation was shown to be needed in larval syncytial-driven wound closure (Galko and Krasnow 2004). The authors could test JNK pathway reporters to assess pathway activation or test if the JNK pathway is needed for syncytial-driven wound closure by expressing a dominant-negative form of Basket JNK in the epithelium.

Or could syncytia formation be regulated by changes in Integrin-mediated adhesion as shown by the Galko lab in Wang et al 2015? They show that wounding provoked a striking relocalization of PINCH and ILK, indicating the disassembly of functional FA complexes concomitant with syncytium formation. Maybe the authors could investigate some of these.

Another general question that the authors raise but don't address enough is whether syncytia-driven wound closure in proliferation-competent epithelia is any different from the one in post-mitotic, polyploid epithelia. Since the mechanism regulating the former is not known, this remains unclear.

Finally, it is not clear, whether syncytia in these proliferation-competent epithelia get resolved after wound healing. Do they get removed and replaced by mononucleated proliferation-competent cells or do the syncytia stay in the epithelium like a scar? The authors should provide some images of wound areas a few hours after wound closure is complete and comment on this.

Minor points:

Figure 3: It would be better to have the microcopy images alongside the quantifications.

Figure 4A: The syncytium at the wound edge here doesn't look straight but wavy. Does it not form an actomyosin cable that straightens the front? Or are there lamellipodia/filopodia?

248: The authors suggest an interesting hypothesis that mitochondria or ER could be pooled in fused cells. It would be nice to see some evidence: e.g. by labeling mitochondria and assessing where they are in syncytia versus mononucleated cells and whether they are concentrated around the wound edge.

141-145 (Figure 4B and C) This example is not completely convincing. First, it is hard to see where the wound edge is. Second, it would be good to include an even later time point when the cell is clearly no longer at the wound edge.

Reviewer #3 (Public Review):

Summary:

White et al. described laser-induced wound healing of the Drosophila pupal notum. They found that the epithelial monolayer is dynamically induced to form syncytia by cell-cell fusion as an important part of repair. They reveal two processes: cell shrinking and border breakage that occur as part of syncytia formation. Expression of GFP in the cytoplasms of some epithelial cells reveals that cytoplasmic contents mix following injury and the GFP rapidly diffuses between cells. Using live imaging they observe that syncytia expand towards the wound, maintain their positions close to the leading edge, and apparently displace smaller cells. They propose that syncytia redistribute cellular components towards the wound facilitating repair and show that labelled actin becomes concentrated at the leading edge.

Strengths:

The manuscript is interesting and on an important and emerging topic of wound healing in a genetically tractable organism. The manuscript is very well written.

Weaknesses:

There are three major issues that the authors must address:

(1) Is cell-cell fusion sufficient to enhance/facilitate wound healing?

(2) Characterization of "border breakdown"; Is this phenomenon disassembly of apical junctions following membrane fusion?

(3) Are cells really shrinking or is it only the apical domains that "shrink" as the cells join the syncytium?

Author response:

The following is the authors’ response to the original reviews.

Public Reviews:

Reviewer #1 (Public Review):

Summary:

This study aims to understand how cell fusion contributes to wound healing using a laser-induced injury in the notum epithelium of a developing fruit fly. The authors meticulously characterize the epithelial fusion events using a live imaging approach and report that syncytia arise by 'border breakdown' and 'cell shrinking'. The syncytial epithelial cells also appear to outcompete mononucleated cells and preferentially dissolve their tangential borders, which correlates with the accumulation of actin at the leading edge.

Strengths:

The strength of this study is the authors' live imaging approach to capture these dynamic fusion events that are a fundamental, yet poorly understood biological process.

Weaknesses:

A major weakness is that all the authors' conclusions are based on descriptive studies, in which the role of cell fusion is not directly tested. This is particularly important because other models of wound induced polyploidization have demonstrated that another cytoskeletal protein, myosin, was upregulated and dependent on endoreplication, and not cell fusion. Therefore it remains unclear to what extent cell fusion, endoreplication, or both are required to outcompete mononucleated cells as well as pool actin as described in this study.

We thank the reviewer for appreciating our live imaging and meticulous approach. In this revision we have identified that the gene Atg1 is required for wound-induced fusion in the pupal notum: when Atg1 is knocked down, there is a reduction in wound-induced cell fusions, both border breakdown and cell shrinking. Analysis of Atg1 knockdown shows that the wounds close more slowly. This is a direct test of the role of cell fusion in speeding wound closure, presented in new Fig. 4.

Reviewer #2 (Public Review):

Summary:

Overall, this study provides a thorough description of the formation of syncytia following wounding of the proliferation-competent diploid epithelium of the pupal notum. While this phenomenon has already been described briefly for this particular tissue by the Galko lab in Wang et al 2015, the authors provide a much more detailed description and characterisation of the process providing some novel insights (radial versus tangential border breakdown, cell shrinkage, timings, syncytia outcompeting mononucleated cells, etc.).

Strengths:

This paper provides an elegant, thorough, descriptive characterisation of syncytia-driven wound closure using state-of-the-art confocal live imaging of the pupal notum. The authors show that laserinduced wounding of this diploid, proliferation-competent epithelium results in the formation of syncytia of various sizes in the first few cell rows around the wound edge, which progressively become bigger as healing proceeds. This results in ~50% of cells becoming part of these syncytia. The cell fusion events were convincingly demonstrated by showing the disappearance of p120ctnRFP and E-Cadherin-GFP from cell-cell borders as well as cytoplasmic GFP mixing of GFPpositive cells with a GFP-negative cell.

Apart from cell-cell fusion by border breakdown that mostly happens in the first 2h following wounding, the authors also found that at later stages of wound healing cell shrinkage following cytoplasmic mixing contributed to sycytia formation.

Next, the authors provided some convincing evidence that syncytia outcompete mononuclear cells for being positioned in the first cell row around the wound.

The authors then show that radial border breakdown occurs much less frequently than tangential border breakdown. They suggest that radial border breakdown reduces the requirement for cell-cell intercalations. They also hypothesise that tangential border breakdown might allow fused cells to share resources and provide more resources to be used near the wound edge, e.g. for actomyosin cable formation. To test this, the authors generate single-cell clones that overexpress Actin-GFP. They then show convincingly how a single Actin-GFP-positive cell in the second cell row fuses with one GFP-negative cell in the first cell row. The Actin-GFP signal then spreads in the fused cell and labels some previously unlabelled actin-rich structure near the wound edge which most likely is the actomyosin cable. This provides some evidence for resource sharing by cytoplasmic mixing following fusion.

Weaknesses:

The authors provide some convincing evidence that syncytia outcompete mononuclear cells for being positioned in the first cell row around the wound. The authors suggest that the syncytial cells might be better able to close the wound. However, some genetic studies would need to be done to establish this more convincingly. E.g. Could the authors genetically block syncytia formation and then show that these wounds now heal slower?

We now present such data in new Fig. 4, which describes knocking down Atg1, previously shown by the Leptin lab to promote wound-induced fusions in larval epidermis. We quantify the resulting reduction in fusion in the pupal notum and show that the leading edge advances more slowly to heal the wound.

The authors suggest that radial border breakdown reduces the requirement for cell intercalation. While this might be true it also raises the question of how the various syncytia facing the wound border change shape to allow the shrinkage of the first cell row over time to allow wound closure. None of the four movies included in the study shows the whole wound healing process until the later stages, making it hard to assess this. It would be good to include one such movie showing the syncytia in the whole wound and comment on this point.

In response to the reviewer's request, we now extend Supplemental Video S1 out through 8 hours after wounding (same video as included previously but extended longer). In this video, as in many of the wounds, it is hard to determine the exact moment of closure because a syncytium extends across the wound whereas the nuclei do not. However, during the process of closure, one can clearly observe the large syncytia becoming more wedge-shaped – drastically reducing the section of their perimeter remaining in contact with the wound’s leading edge.

In addition, we now explore how syncytia reduce the need for intercalation in a computational model, presented in new Fig. 7 and Supplemental Videos S5 and S6. One can observe the modeled syncytia becoming similarly wedge-shaped. The modeling shows that the presence of syncytia and their ability to reshape can speed closure by about 1/3 even if the syncytia have no special properties aside from their relative size.

In both the experiments and models, some syncytia are also removed from the leading edge by intercalation, but the presence of syncytia reduces the total number of intercalations needed.

The authors hypothesise that tangential border breakdown might allow fused cells to share resources and provide more resources to be used near the wound edge, e.g. for actomyosin cable formation. They show convincingly through the fusion of a single Actin-GFP-positive cell in the second cell row with a GFP-negative cell in the first cell row that Actin-GFP spreads in the fused cell and labels the previously unlabelled actomyosin cable. While the hypothesis of resource sharing to improve healing is intriguing and makes sense, this experiment doesn't necessarily prove the benefit of resource sharing. It does show cytoplasmic mixing following fusion, now allowing the GFPlabelled actin to diffuse and be incorporated into the actomyosin cable. In a wild-type condition, fusion would not increase the total concentration of resources, although it would increase the total amount of resources within this bigger fused cell. The question is whether resource sharing without increasing the protein concentration is beneficial and increases the efficiency of certain wound healing mechanisms. There might be a benefit of cell fusion, if for example certain resources were only present in limited amounts or if protein transport could increase the concentration locally. To provide better evidence for the hypothesis that resource sharing improves wound healing, maybe the authors could look at the actomyosin cable in a wounded epithelium (such as in Figure 4E, F), in which all cells express MyoII-GFP. The authors could compare the average intensity of the actomyosin cable at the wound edge in mononucleated cells versus in syncytia. If resource sharing is indeed beneficial, it might be that the actomyosin cable is stronger/brighter in syncytia or it forms quicker.

We agree with the reviewer that we have not "proved the benefit of resource sharing". Because we cannot inhibit resource sharing while still allowing cell fusion, we can think of no rigorous way to test this hypothesis. We appreciate the reviewer's suggestion of quantifying the myosin at the leading edge cable, but we can imagine too many caveats to the interpretation to make it worthwhile. Rather, we accept the limitation that this is an untested, perhaps untestable, hypothesis -- but nevertheless intriguing.

We do want to clarify ideas about the concentration of resources after fusion. We agree that the overall concentration of a given resource (mass/volume) throughout a syncytium would be the same as the overall concentration in the unfused progenitor cells; however, a syncytium would have a larger total resource mass to direct subcellularly, allowing for local subcellular concentration to be greater in a syncytium vs. an unfused cell. We demonstrate this subcellular localization of actin in a syncytium twice, in Fig. 7C and E (previously Fig. 6C,E), which we think is evidence for increased local concentration.

The biggest limitation of this study is that the authors don't address how the formation of these syncytia is regulated. While the manuscript in its current form provides some valuable new insights into syncytial-driven wound closure, it would be much more informative if it also provided some mechanistic details. The authors could test if some of the mechanisms shown to regulate syncytial formation in other types of syncytia-driven wound healing are also involved here. E.g. Yorkie was shown to negatively regulate cell fusion in adult syncytial-driven wound closure (Losick et al 2013). The authors could test for the effect of Yorkie-RNAi in the epithelium on wound closure and syncytia formation. Expression of the dominant negative RacN17 also blocked cell fusion in adult syncytial-driven wound closure (Losick et al 2013).

Moreover, JNK activation was shown to be needed in larval syncytial-driven wound closure (Galko and Krasnow 2004). The authors could test JNK pathway reporters to assess pathway activation or test if the JNK pathway is needed for syncytial-driven wound closure by expressing a dominantnegative form of Basket JNK in the epithelium.

Or could syncytia formation be regulated by changes in Integrin-mediated adhesion as shown by the Galko lab in Wang et al 2015? They show that wounding provoked a striking relocalization of PINCH and ILK, indicating the disassembly of functional FA complexes concomitant with syncytium formation. Maybe the authors could investigate some of these.

We investigated the role of JNK in fusion by expressing bskDN on one side of the wound. Comparing the numbers of border-loss fusion on each side, we did not find a significant difference in our seven-sample cohort (see Author response image 1). If we had increased the sample size, we may have found a significant difference with a small effect size, but because of the small difference in fusions on each side we did not think this was worth pursuing. Instead, we include data that the autophagy gene Atg1 is required for cell fusion in new Fig. 4, which begins to address mechanism, and relates the wound-induced fusion described here in pupae to wound-induced fusion shown in larvae. A complete mechanism for wound-induced fusion is outside the scope of this paper, as we focus on the function of syncytia in healing wounds.

Another general question that the authors raise but don't address enough is whether syncytia-driven wound closure in proliferation-competent epithelia is any different from the one in post-mitotic, polyploid epithelia. Since the mechanism regulating the former is not known, this remains unclear.

We now include a paragraph on this question in the discussion.

Finally, it is not clear, whether syncytia in these proliferation-competent epithelia get resolved after wound healing. Do they get removed and replaced by mononucleated proliferation-competent cells or do the syncytia stay in the epithelium like a scar? The authors should provide some images of wound areas a few hours after wound closure is complete and comment on this.

To answer the reviewer’s question: some but not all syncytia do get removed during wound closure by remarkable apoptotic/extrusion events. This will be the subject of a future manuscript, as it is outside the scope of this paper focusing on the function of syncytia in promoting wound healing.

Minor points:

Figure 3: It would be better to have the microcopy images alongside the quantifications.

The images in Figs. 1 and 2 show the border breakdown and shrinking cells, and we do not see benefit in adding them in Fig. 3.

Figure 4A: The syncytium at the wound edge here doesn't look straight but wavy. Does it not form an actomyosin cable that straightens the front? Or are there lamellipodia/filopodia?

We assume the reviewer is asking about the wavy edge outlined at 400 min after wounding (now Fig. 5A). As shown by Jacinto and colleagues in the first pupal wounding paper (JCB 2013), the actin cable forms quickly, within 15 minutes; much later actin protrusions extend from the leading edge to close the wound. This result is consistent with the wavy edge 400 min after wounding.

248: The authors suggest an interesting hypothesis that mitochondria or ER could be pooled in fused cells. It would be nice to see some evidence: e.g. by labeling mitochondria and assessing where they are in syncytia versus mononucleated cells and whether they are concentrated around the wound edge.

Although we don't think that exploring mitochondria or ER is central to this manuscript, we agree it would be an interesting question for the future.

141-145 (Figure 4B and C) This example is not completely convincing. First, it is hard to see where the wound edge is. Second, it would be good to include an even later time point when the cell is clearly no longer at the wound edge.

We have revised this figure, now Fig. 5B,C, to include a later image at 360 min after wounding healing, and this additional panel clarifies that the smaller cell leaves the wound edge. As noted in the text, the wound edge is indicated by the cell borders lacking p120ctn.

Reviewer #3 (Public Review):

Summary:

White et al. described laser-induced wound healing of the Drosophila pupal notum. They found that the epithelial monolayer is dynamically induced to form syncytia by cell-cell fusion as an important part of repair. They reveal two processes: cell shrinking and border breakage that occur as part of syncytia formation. Expression of GFP in the cytoplasms of some epithelial cells reveals that cytoplasmic contents mix following injury and the GFP rapidly diffuses between cells. Using live imaging they observe that syncytia expand towards the wound, maintain their positions close to the leading edge, and apparently displace smaller cells. They propose that syncytia redistribute cellular components towards the wound facilitating repair and show that labelled actin becomes concentrated at the leading edge.

Strengths:

The manuscript is interesting and on an important and emerging topic of wound healing in a genetically tractable organism. The manuscript is very well written.

Weaknesses:

There are three major issues that the authors must address: 1. Is cell-cell fusion sufficient to enhance/facilitate wound healing? 2. Characterization of "border breakdown"; Is this phenomenon disassembly of apical junctions following membrane fusion? 3. Are cells really shrinking or is it only the apical domains that "shrink" as the cells join the syncytium.

We thank the reviewer for recognizing the importance of this topic. Our responses to the specific weaknesses are below.

Recommendations for the authors:

Reviewer #1 (Recommendations For The Authors):

Major Components:

(1) For syncytia measurements the nuclei are labeled with histone-GFP which is expressed in all cell types. How do you know the nuclei within the cell junctions are epithelial and not another cell type, such as immune cells recruited to the injury site? It would be helpful to verify the number of nuclei per cell using an epithelial-specific nuclear marker as well. This could be via epithelial Gal4-specific expression of a UAS-nls-GFP.

This is an interesting point. In response to the reviewer's question, we investigated by doing the converse experiment, labeling immune cells with hml-Gal4, UAS-GFP, and observing what they do after wounding (analyzing six wounded pupae). They do get recruited to the wound, but they remain either in the wound center or at the basal side of the leading edge. Because they are labeled with cytoplasmic GFP, we would be able to ascertain whether they fused with epithelial cells because they would share their GFP with epithelial cells in the epithelial plane, and they did not. Thus we are confident that the many syncytial nuclei are not derived from immune cells. Our live tracking throughout the manuscript, and specifically of GFP-labeled clones, also supports our interpretation that syncytial nuclei derive from epithelial cells.

(2) The manuscript focuses on cell fusion, but other mechanisms of cell enlargement have been observed to occur during wound healing via endoreplication. To what extent do epithelial cells in pupae notum endocycle or endomitosis post injury? It is unclear if the increase in syncytia size during a 1-2hr period could also be due to endomitosis, which would also increase nuclear number.

Since the first submission of this manuscript, we published our results demonstrating limited wound-induced endoreplication after this type of explosive laser injury to the pupal notum (White et al, 2024, PMID: 38495588). We chose to publish this work separately because we could not offer the same degree of depth for endoreplication as we could for fusion: our pupal notum injury model is extremely well-suited to analyzing cell fusion and wound closure by live imaging; however, it is not particularly well-suited for analyzing endoreplication in fixed tissue. With respect to reviewer's question about endomitosis -- i.e. nuclear divisions that are not accompanied by cell divisions -- even after many years we have not observed an endomitosis event, which would be visible by live imaging, whereas we frequently and easily observe mitosis of diploid cells.

(3) One of the major conclusions of this study is that cell fusion is necessary to pool resources at the leading edge. Therefore it is critical that authors identify a mechanism to inhibit cell fusion to test this assumption.

We now include new Fig. 4, an analysis of the role of Atg1 in promoting wound-induced fusion and wound closure. These results build on the finding of the Leptin lab (Kakanj et al, 2022) that autophagy genes are required for fusion. Our results are consistent with the model that syncytia speed wound closure.

(4) There is evidence that myosin increases in endoreplicating cells during wound healing hence it is, maybe equally - if not more - probable that the increase in resources (here actin-GFP) at the leading edge is dependent on endoreplication instead of cell fusion.

Some of the new data we provide for this manuscript is a correlation between cell size and distance traveled, showing that larger cells travel more within the wound (Fig. 4F,G). Endoreplication would certainly be expected to contribute to increasing cell size, and our published 2024 data indicates that there can be one extra S-phase induced by these types of wounds. Doubling the genome is not a significant contribution to cell size compared to the 10s of nuclei we observe in syncytia from fusion. Nevertheless, we do not claim that actin is the only important resource that can be pooled subcelluarly for the benefit of the cell; we use it only as a proof-of-principle. Finally, we discuss the work on myosin in wound-induced endoreplicating cells (Losick and Duhaime, 2021).

Reviewer #3 (Recommendations For The Authors):

Major comments

(1) Can induction of epithelial fusion enhance wound healing?

Different epithelial cell-cell fusion processes have been well-characterized: i) Trophoblast fusion in the placenta mediated by Syncytins. ii) Viral induced cell-cell fusion mediated by diverse viral glycoproteins (e.g. gp41 from HIV, Hemaglutinin from Influenza, GP from Ebola, and G glycoprotein from VSV). iii) Epidermal, myoepithelial, and other epithelial cell-cell fusion in C. elegans mediated by EFF-1 and AFF-1. iv) Cell-cell fusion in the eye lens (unknown fusogens). The authors may want to compare and discuss the temporal dynamics and intermediates observed in the diverse processes of epithelial cell-cell fusion with the characterization of syncytia formation during wound healing of the Drosophila pupal notum. Since some of these characterized cell-cell fusogens can fuse heterologous cells, including Drosophila S2 cells (Shilagardi et al., 2013; https://pubmed.ncbi.nlm.nih.gov/23470732/), the authors may consider expressing these fusogens in Drosophila pupal notum before, during and after injury. This could determine whether syncytia formation is sufficient to stimulate efficient wound healing.

We thank the reviewer for the suggestion of comparing and discussing temporal dynamics and intermediates observed in the many types of epithelial fusion that are well understood. Regretfully, we do not think this article is the right venue for such a complex discussion, especially since we have little by way of comparison in our own wound-induced fusion data. As for overexpression of fusogens, it is an intriguing idea to force cell fusion with a heterologous fusogen such as EFF-1 and then investigate any resulting changes in wound healing. However, since half the cells within 70 µm of the wound already fuse even without a heterologous fusogen, it seems unlikely we could meaningfully increase the level of cell fusion unless we expressed the fusogen universally, forcing the fusion of nearly all the epithelial cells as well as other cells throughout the body that express pnr-Gal4. Because the overexpression of EFF-1 in C .elegans results in lethality (PMID: 26854231), a widespread induction of fusion would be expected to cause other types of physiological problems that would interfere with the interpretation of wound closure rates. Further, the conditional expression tools in Drosophila allow excellent spatial control, but temporal control is still somewhat low-resolution, so that we would have difficulty expressing EFF-1 before, during, and after wounding at times that would be relevant to understanding wound healing.

(2) The phenomenon of "border breakdowns" described here is not clear. The authors are probably studying the disassembly of the apical junctions following the initiation of membrane fusion and pore expansion. This should be clarified by using membrane labels to directly observe membrane fusion. Researchers have used electron microscopy and membrane fluorescent probes to follow cell-cell fusion. For example, GPI-mCherry, FM4-64, lipid-modified-GFPs (e.g. PH-domain fluorescently labeled proteins) DiO, DiI, and many others. See for example: Markosyan et al., 2016; https://pubmed.ncbi.nlm.nih.gov/26730950/; Mohler et al., 1998; https://pubmed.ncbi.nlm.nih.gov/9768364/; Meng et al., 2020; https://pubmed.ncbi.nlm.nih.gov/32668210/.

We agree completely with the reviewer, that border breakdowns represent the disassembly of apical junctions following initiation of membrane fusion and pore expansion. Direct evidence for this order of events is found in the video stills of Figure 1 panel I and video S2, which show that cytoplasmic GFP is transferred to the fusion partner 14 minutes before there is a visible decrease in the apical adherens junction marker p120ctn. The reproducibility of this order of events is documented in Fig. 3: among 107 GFP-labeled cells, 30 of them first visibly shared GFP with a fusion partner, and then 11/30 displayed border breakdown, 16/30 displayed cell shrinking, and 3/30 did not fuse. This last category is consistent with a fusion pore that closed rather than expanded productively. Although we have obtained TEM images of wound-induced fusion pores, these are included in another manuscript currently in revision and so cannot be included here, and further these EM images do not shed light on border breakdown per se, as only live imaging can establish the relationship between border breakdown and pore formation (GFP-sharing).

(3) The observation of cell shrinking may be misleading. The process the authors describe as "cell shrinking" may involve shrinking of the apical domain, maintaining the cell volume. To clarify this process, the authors may simultaneously label the apical and basolateral domains. It is possible that fusion pore formation occurs in the basolateral, apical, or both domains. The apical shrinking could reflect the migration of the apical junctions following fusion. A similar process has been described in epidermal and vulval cells of C. elegans and other nematodes (Mohler et al., 1998; https://pubmed.ncbi.nlm.nih.gov/9768364/; Sharma-Kishore et al., 1999; https://pubmed.ncbi.nlm.nih.gov/9895317/; Kolotuev and Podbilewicz 2008; https://pubmed.ncbi.nlm.nih.gov/18031720/).

We thank the reviewer for pointing out these examples of cell fusion in nematodes, and we now compare our findings to Mohler et al, 1998. In Fig. 2D, we specifically investigated what happened to the cell volume of these shrinking cells, and we hope we have now clarified both the text and the annotations on the figure to make our findings more clear. In the X-Z plane, the entire cell volume of two shrinking cells is visible from cytoplasmic GFP labeling. For both cells, the cytoplasmic volume moves laterally into the neighboring syncytia, appearing to initiate the movement from the basal-most area of the cell so that 150 minutes after wounding, both cells have a reduced apical footprint and only a whisp of apically-oriented cytoplasm, with the remainder of the cytoplasm having moved into the syncytia. These images make it clear that fusion is occuring, and that when the apical area disappears the corresponding cytoplasm has also moved into the territory of the neighboring syncytium. In response to the reviewer's suggestion, we did try labeling basolateral domains, but the fluorescent proteins we examined are not restricted to the basolateral domain and are difficult to interpret.

Minor comments

(1) Lines 40-43. Repair of injuries has also been observed in non-proliferative syncytial epidermal cells and involves cell-cell fusogens. The authors may want to include this reference: Meng et al., 2020; https://pubmed.ncbi.nlm.nih.gov/32668210/.

We thank the reviewer for the suggestion, and we have included this reference in the Discussion paragraph about fusogens.

(2) Lines 128-130. Is "Shrinking fusion" an "artefact"?

The apical junction shrinks not the cell. I suggest following basolateral membranes to see whether the cell is indeed shrinking as it fuses. The authors may want to share whether the cell volume is maintained but spills into an existing syncytium; the apical junction shrinks because it disappears/disassembles (see also Major comment 3).

As discussed in Major comment 3, we do provide evidence that the cell cytoplasm spills into an existing syncytium. Perhaps the reviewer finds the term "shrinking cell" to be misleading, as we all agree that the cell contents do not disappear. We have updated the manuscript to use the term "apical shrinking" throughout.

(3) Lines 157-159. Are these small cells or instead they are small apical junctions? The interpretation should include basolateral domains of the small cells to determine their size! It is also possible that some small cells have fused with the syncytia but on the basolateral domain without apical junction disassembly.

We appreciate the reviewer's rigor. As noted above, we were not able to analyze the basolateral domains of these cells. Because our all analyses are live-imaging videos, we are able to identify the cells are undergoing apical shrinking and clearly delineate those from stable diploid cells. We now realize that the term "small cells" is confusing and can be mixed up with apical shrinking. These cells are not "small" but normal sized, small only in comparison with the gigantic syncytia around them. We have removed the term "small" from this description.

(4) Lines 204-206. Many genes required for myoblast fusion in Drosophila have been shown to play a role in different stages of cell-cell fusion. Do they play roles in epithelia fusion during wound closure in the pupal notum?. For example, actin polymerization? Dynamin? Ig-domain and integrin cell adhesion machineries?

We now provide a new Fig. 4 that shows that the autophagy gene Atg1 reduces wound-induced cell fusion, as it does in larvae (Kakanj et al, 2022), and importantly these wounds close more slowly. We have not analyzed mutants in actin polymerization because we are confident they would interrupt many aspects of wound healing. The Galko lab has identified that integrins suppress wound-induced cell fusion in larval epidermis, but we have not tested these. We have a manuscript in revision demonstrating a requirement for Dynamin and other endocytosis genes in wound-induced fusion, and without dynamin-mediated fusion, these wounds close more slowly.

Version 2

Abstract

In response to injury, cells proliferate, migrate and invade to replace missing cells and close wounds. However, the role of other wound-induced cell behaviors is not understood, including the formation of syncytia (multinucleated cells). Here, we use in vivo live imaging to analyze wound-induced syncytia in mitotically competent Drosophila pupae. We find that almost half the epithelial cells near a wound fuse to form large syncytia. When the autophagy gene Atg1 is knocked down, fewer syncytia form, and wounds close more slowly. Further, a computational model of tissue fluidity indicates that cell fusion speeds wound closure time by about one third. Syncytia use several routes to speed wound repair: they outpace diploid cells at the wound margin to lead the initial resealing of the wound; they reduce the need for intercalation as the tissue reshapes during closure; and they pool resources of their component cells to concentrate them toward the wound margin. In addition to wound healing, these properties of syncytia are likely to contribute to their roles in development and pathology.

Introduction

Injury is a constant reality of life, and survival requires all organisms to repair wounds. Wound-induced cell behaviors like proliferation, migration, and invasion replace missing cells and close wounds (Martin et al., 2024). Other cell behaviors are induced around wounds, but their contribution to wound healing is not well understood, e.g., the fusion of cells into syncytia. Syncytia are one type of polyploid cell, and it is generally appreciated that increases in ploidy – number of genomes per cell – is a common response of post-mitotic tissues and cells to injury (Losick et al., 2013; Nandakumar et al., 2020; Tamori and Deng, 2013). Wound-induced epithelial syncytia were first observed around epidermal puncture wounds in Drosophila larvae and adults (Galko and Krasnow, 2004; Losick et al., 2013), consistent with the idea of polyploidy induction in non-proliferative cells. However, recent studies have observed syncytia around wounds in mitotically competent tissues: around laser-ablation wounds in Drosophila pupal epidermis (Wang et al., 2015) and in zebrafish epicardium damaged by endotoxin, microdissection, or laser ablation (Cao et al., 2017). Further, injury associated with the surgical implantation of biomaterials can cause immune cells to fuse into multinucleated giant cells, which are associated with rejection (Al-Maawi et al., 2017). Similarly, injury induces bone marrow-derived cells to fuse with various somatic cells to promote repair (Alvarez-Dolado et al., 2003; Corbel et al., 2003; Davies et al., 2009; Nygren et al., 2004). The many instances of syncytia being induced by wounds raise the possibility that syncytia offer an adaptive benefit. It is not clear, however, what that benefit is.

Syncytia can form either by endomitosis – mitosis without cytokinesis – or by cell-cell fusion. Fusion is widely observed throughout development in both vertebrates and invertebrates: for example myoblasts fuse into muscles (Kim et al., 2015a; Lehka and Rędowicz, 2020), and fusions occur in the lineages of the C. elegans hypodermis (Podbilewicz and White, 1994) as well as vertebrate trophoblast (Renaud and Jeyarajah, 2022) and osteoclast (Søe, 2020). Cell fusions are also observed in disease: pathogen-induced epithelial fusion allows spreading of many viruses including human respiratory syncytial virus and SARS-CoV-2 (Leroy et al., 2020); and the fusion of cancer cells with bone-marrow derived cells is implicated in metastasis (Pawelek and Chakraborty, 2008).

Here, we use live imaging and clonal analysis to understand the behavior of syncytia following laser wounding in the Drosophila pupal notum. The unwounded notum is a monolayer epithelium composed of mononuclear diploid cells that are mitotically competent. Nonetheless, during the first several hours after wounding, many of the surrounding cells fuse to form giant syncytia. We previously reported that after wounding the pupal notum, some wound-adjacent cells undergo a single round of endoreplication, increasing their ploidy two-fold (White et al., 2024), but fusion increases ploidy far more, with some wound-induced syncytia observed to have more than a dozen nuclei. Some cell fusions are obvious with apical borders breaking down between cells, while others surprisingly appear as shrinking of a cell’s apical surface. Combined, fusion is a common fate of cells near wounds: about half the cells fuse to form syncytia within 70 µm of a wound with 30 µm radius. To understand the role of fusion-mediated syncytia in wound repair, we undertook two types of analysis, first analyzing an autophagy mutant that reduces fusion after wounding, and second comparing the behaviors of syncytial and mononuclear cells within the same wound. Compared to their smaller mononuclear neighbors, syncytia outpace mononuclear cells to the leading edge. We propose that syncytia accelerate wound repair because cell fusion limits the need to negotiate cell intercalations as the wound closes, a hypothesis supported by computational modeling; and because clonal analysis demonstrates that syncytia mobilize and transport cell resources such as actin from distal cells to the wound margin.

Results

A mitotic tissue utilizes cell-cell fusions during wound repair

The pupal notum is a monolayer columnar epithelium that will become the epidermis of the adult fly. At the time of our analysis, 12-18h after puparium formation (APF), it sits atop a basement membrane that is still developing (Mehaffey et al., 2024), and it is composed of diploid cells that undergo regular mitotic cycles (Guirao et al., 2015) (Figure 1—figure supplement 1A-C). To analyze cell behaviors around wounds, we live-imaged after laser ablation. Epithelial cell borders were labeled by the adherens junction protein p120ctnRFP (Ogura et al., 2018) and nuclei were labeled with histone His2GFP. Two hours after wounding, we observed what appeared to be prominent syncytial (multi-nucleated) cells around the wound (Figure 1A); experiments described below demonstrated that these were indeed syncytia. Some appeared to contain over a dozen nuclei within epithelial borders. For both syncytial and mononuclear cells, it was difficult to assign nuclei precisely to cell borders because notum epithelial cells are not oriented at right angles with respect to the surface, and in 2-D projections, a nucleus was frequently observed outside the cell’s apical border (Figure 1—figure supplement 1D).

Wounds induce epithelial syncytia via cell fusion.

A) Syncytia form within 2 h post wounding, evident by the clustering of multiple nuclei within cell borders. B) The number of nuclei per syncytia increases over time after wounding. Number of nuclei was estimated based on area and nuclear density (see text) for the 3 largest syncytia of 3 different wounds, mean and SD. C-D) Larger wounds generate larger syncytia. Images are at 3 h post wounding, single syncytium outlined. E) Syncytial apical area is proportional to initial wound size. Each dot represents a wound, made with either low or high laser energy as sown. For each wound, the mean area of the three largest syncytia is plotted. Bars represent SD. F) A time course of six cells fusing within 30 min after wounding. Apical borders are lost (white arrowhead) as syncytia form. Original cells are numbered. G) All borders lost to cell fusion (white) mapped to cells in the first frame after wounding. The leading edge of wound closure will form at dashed line; cells within the shaded area were damaged by the wound and will be dismantled. H) Distance from the wound center vs time for all border breakdowns in 3 wounds. Each symbol represents a cell border. Leading-edge locations indicated by solid lines. I) Cytoplasmic GFP is expressed in cell 1 before wounding and mixes with neighboring cells 2-4 by 2 min after wounding. Cytoplasmic sharing is followed by the lagging fusion indicator of visible border breakdown (white arrow). The fates of cells 3 and 4 are shown at later times in Figure 2D. Maximum intensity projections in A, C-D, G; single Z slices in F, I. Scale bars: A,C,D,G = 20 µm, F,I = 10 µm. W and red star indicate wound.

Accordingly, using apical area and nuclear density, we estimated the number of nuclei within the 3 largest syncytia in different wounds. The number of nuclei in these syncytia increased over time: at 1 h after wounding the three largest syncytia contained an average 3-13 nuclei, and 2 h after wounding they nearly doubled to 6-20 nuclei (Figure 1B). Interestingly, larger wounds generated syncytia with larger apical areas and more nuclei, proportional to wound size, suggesting syncytium formation is a dynamic and scalable response to injury (Figure 1 C-E).

Live imaging of cells after wounding revealed the gradual loss of p120ctn along the border between two epithelial cells followed by rounding of the resulting syncytium (Figure 1F,I, and Video 2), such that the initial two-cell morphology gradually became a one-cell morphology. We termed the visible loss of p120ctn “border breakdown”. We determined that as p120ctn was lost, the adherens junction protein E-cadherin was also lost (Figure 1—figure supplement 1E-H) indicating adherens junctions disassembly along the border between the cells. Border breakdowns were found spatially clustered in the first three to four rows of cells, 30-50 μm from the wound center, and they occurred primarily within the first hour after wounding, some within 10 minutes after wounding (Figure 1G,H). These events could be explained by wound-induced cell fusion or by wound-induced epithelial-to-mesenchymal transition. To ask whether border breakdowns represented cell fusions, we analyzed cytoplasmic mixing. Individual GFP-labeled cells (clones) were generated at random locations using the flip-out Gal4 technique (Pignoni and Zipursky, 1997). Before wounding, the level of cytoplasmic GFP fluorescence was stable yet exhibited cell-to-cell variability, allowing some differentiation of cells by intensity (Figure 1Ii). Minutes after a nearby laser ablation, GFP was observed to diffuse from labeled cells into neighboring unlabeled cells. GFP mixing between two cells was followed by the eventual loss of their shared p120ctn-labeled cell border, confirming that border breakdown is indeed cell fusion, but that cytoplasmic mixing occurs more than 10 minutes before the border breakdown is first observed to start (Figure 1I, Video 2). GFP mixing always preceded border breakdowns (n=11). Cytoplasmic GFP mixing resulted in a consistent level of fluorescence across the two cells, indicating that GFP mixing is a reliable indicator of plasma membrane fusion. The order of events demonstrates that plasma membrane fusion occurs many minutes before the visible disassembly of the apically located adherens junctions. Thus, epithelial fusion is a rapid local response to wounding, and border breakdown is a lagging indicator of cell fusion.

Apical cell shrinking also represents cell-cell fusion, evident later during wound closure

Border breakdowns, representing cell fusion events, occurred mostly within 1 h after wounding, so it was unclear how syncytia grew in size and nuclear number between 1-2 h after wounding. However, an unexpected cell behavior was frequently observed during this time: the apical area of diploid epithelial cells shrank until they disappeared from the epithelial sheet, a phenomenon we termed “apical shrinking” (Figure 2A). Although the appearance of apical shrinking suggested cell extrusion, we were surprised at its high frequency around wounds (Figure 2E), as wounds have already lost many cells and would be expected to prioritize cell survival to speed barrier repair. To understand the fate of apical-shrinking cells, we tracked their nuclei and determined that they did not extrude; rather, all apical-shrinking cell nuclei moved laterally within the plane of the epithelium to join neighboring syncytia (n=7, Figure 2B), suggesting that apical shrinking is an indicator of cell fusion. To better understand this behavior, we analyzed individual GFP-labeled cells so that we could observe the fate of the cytoplasm.

Cell fusion often appears as apical shrinking.

A) The apical footprint of an epithelial cell shrinks in the epithelial plane after wounding (yellow arrow). B) Same sample as A, showing that the nucleus (asterisk) of the apically shrinking cell enters a neighboring syncytium, outlined in Biv. C) Cell 1 expresses actin-GFP before wounding. After wounding, cell 1 fuses with cells 2-6 evidenced by GFP sharing, then the apical surface of cell 1 shrinks. D) Two cells undergoing apical shrinking are identified in the X-Y plane with white and yellow arrows, shown at three times after wounding. Below, two X-Z projections are shown for each, at Y1 and Y2 as indicated, with the Z-plane of the adherens junctions (ZAJ) indicated by dotted lines. Rather than extrude, cytoplasm moves to the right as the cell depth in the Z-axis diminishes, so that cytoplasm joins with the neighboring wound-proximal syncytium. These video frames are a continuation of the sample shown in Figure 1I. E) All cells displaying apical shrinking mapped to first frame after wounding. F) All apical shrinking and border breakdowns were tracked in 3 wounds. Border-breakdown fusions happen sooner after wounding than apical-shrinking fusions. A, B, C, D (top), and E show maximum intensity projections. Lower panels in D show X-Z projections. Scale bar for all panels A and B shown in Biv. A, B, C, D (top),E = 10 µm; D (X-Z projections) = 5 µm.

Like with border breakdowns, GFP mixing in the X-Y plane of the epithelium was observed to precede the initiation of apical shrinking, although the interval between GFP mixing and apical shrinking was longer, one or more hours (Figure 2C, Video 3). In X-Z projections, GFP labeled cytoplasm associated with apical shrinking shifted laterally to join with a neighboring cell as shown for two apical-shrinking cells in Figure 2D. For the cell labeled with the yellow arrow, before apical shrinking at 60 min post wound, its GFP-labeled cytoplasm extended from the plane of the adherens junctions (ZAJ) basally over 8 µm; when apical shrinking was underway at 120 min post wound, its cytoplasm extended less than 5 µm down from the apical plane; and when shrinking was almost complete at 150 min post wound, it extended only 1-2 µm down from the apical plane, appearing as a whisp of cytoplasm connecting the shrinking apical surface with the neighboring syncytia. A similar loss of basal depth over time is observed in the cell labeled with the white arrow. In both these cells, as the cell lost its basal volume, cytoplasm moved laterally to join the neighboring syncytia. Thus, analysis of both nuclei and cytoplasm demonstrated that apical shrinking represented cell fusion events. The spatial distribution of apical shrinking around wounds was similar to border breakdowns, but fusions associated with apical shrinking occurred later and were more numerous (Figure 2E,F). Thus, cell fusion is evident as either border breakdown or apical shrinking, which both occur after GFP sharing.

To determine what percentage of cells around a wound will fuse, we analyzed over 100 single epithelial cells randomly-labeled with cytoplasmic GFP within the radius of observed fusion (80 µm), tracking them for 6.5 h to assess their fate (Figure 3A,B): a full quarter of the cells fused (25%), sharing GFP before borders breakdown or apical shrinking; 67% persisted as diploid cells, most with stable GFP, but infrequently (n=3) with GFP mixing and no subsequent cell fusion; the remaining 7% could not be tracked (Figure 3—figure supplement 1A). All GFP-labeled cells with apical shrinking had previously shared cytoplasm, indicating that all apical shrinking represents fusion. The observation of rare diploid cells with previous cytoplasmic mixing confirms that plasma membrane fusion occurs before the later steps of border breakdown/apical shrinking, as these cells represent fusion initiation events (fusion pore) but were unable to productively stabilize and expand the site of fusion and so returned to the diploid state. Fusing cells were strongly skewed toward the center of the wound: within 70 µm, about half the cells (47%) underwent fusion (Figure 3A, left), and fusion continued for over 300 min after wounding (Figure 3—figure supplement 1B,C). Apical shrinking accounts for more than half of all fusing cells (Figure 3C), and the spatial distribution of fusing cells that shrank vs. lost borders was similar (compare Figures 1G and 2E). However, apical shrinking began later than border breakdowns (Figure 2F, Figure 3—figure supplement 1C), continued for several hours after wounding (Figure 2F, Figure 3—figure supplement 1C), and took longer to complete (Figure 3—figure supplement 1D). Thus, fusion associated with apical shrinking accounted for continuing syncytial growth after border breakdowns subsided. Further, cell fusion is a persistent behavior over the course of wound closure.

Half the cells near the wound fuse to form syncytia, demonstrated by tracking individual cell fates.

A) All GFP labeled cells in the region of fusion (80 µm) were tracked in 5 wounds over 6.5 h after wounding to determine frequency of fusion initiation (GFP mixing) and later morphological changes into syncytia. All morphological fusion events were preceded by GFP sharing, and 90% of GFP sharing events were followed by morphological fusion. Untrackable cells lost GFP, see Figure 3—figure supplement 1A. B-C) Pre-wound location of persisting and fusing cells from panel A is shown with respect to the wound center. Panel B shows that fusion is common within 70 µm. Panel C shows that apical-shrinking fusion and border-breakdown fusion occur at similar distances from the wound.

Our individual cell tracking indicated that syncytia were the result of epithelial cell fusions. However, phagocytic hemocytes are known to migrate to the sites of pupal wounds (Antunes et al., 2013). To determine if hemocytes fused with epithelial syncytia around wounds, we labeled hemocytes with cytoplasmic GFP (hml>GFP) and observed their behavior in live imaging. In six wounds, we observed GFP-labeled hemocytes at the basal side of the wound margin; however, we did not observe any events where GFP was transferred from hemocytes to the epithelium, ruling out the fusion of hemocytes with epithelia in pupal wounds (Supplemental Video 4) .

Syncytia close wounds faster

To evaluate the role of cell fusion and syncytia in repairing wounds, we wanted to analyze a mutant that blocked syncytia formation. It has been reported that several individual autophagy (Atg) genes are required for syncytia formation around small laser ablation wounds in the polyploid Drosophila larval epithelium (Kakanj et al., 2022), so we asked if autophagy promotes fusion around laser wounds in the diploid pupal epithelium. Using a dsRNA line known to disrupt syncytia in larvae, we knocked down Atg1 on one side of the wound expressing pnr-Gal4, as illustrated in Figure 4A, reserving the other side of the wound as an internal control, modifying an internally controlled wounding system we have previously reported (O’Connor et al., 2021b). Visualizing Ecad-labeled cell borders across the wound, we found that indeed Atg1 is required for syncytia formation in pupal wounds, like it is in larval wounds (Figure 4B).

Atg1 promotes wound-induced fusion and faster wound closure.

A) Overview: Atg1 was knocked down in central pnr domain (pink), visualized by nls-mCherry (not shown). A laser pulse was targeted to the edge of the pnr domain (right or left) so that the wound extended into both Atg1 knockdown and control domains, allowing the symmetry of response to be analyzed. B-C) Inhibition of Atg1 decreased the frequency of wound-induced fusion in the pnr domain. For border loss, n = 11 pupae, p < 0.0001. For apical shrinking, n=9 pupae, p = 0.0008 (paired T-test). D-E) Wound closed more slowly when Atg1 was knocked down. (D) Radius of the wound was calculated by the distance between the wound center and the leading edge in pnr or control domains. (E) Difference in the wound radius between pnr and control domains over time (i.e., wound asymmetry) is shown. Each dot represents one wound. With no genetic manipulation in either domain (black), wound radius was slightly larger in pnr than in control domain. When Atg1 was knocked down in the pnr domain, the asymmetry grew over time indicating slower wound closure on the Atg1 knockdown side. p = 0.0031 (5 hr) calculated from two-way ANOVA, fit full model, comparing each cell mean with the other cell mean in that row. error bar = 1 +/- SEM. F-G) Larger cells (i.e. syncytia) move further during wound closure, evident in control wounds (F) or with Atg1 knockdown (G). Each data point represents a cell or syncytium present at the leading edge when the wound is half closed. The curve of square root area was overlaid for comparison, as it represents the amount of movement expected based on reshaping of a cell/syncytium from an initial round state to one strongly elongated towards the wound.

Specifically, we observed a reduction in cell fusion, both reductions of border loss and apical shrinking (Figure 4C). We next compared wound closure rates on the two sides of the wound (Figure 4D,E). Because leading edge migration is slightly delayed on the pnr-Gal4 side of the wound even in control wounds (pnr>+), we compared the difference in wound radii on the two sides, comparing control wounds to pnr>Atg1RNAi wounds. By 5h after wounding, the pnr>+ side lagged behind the other side by only about 10 µm, yet the pnr>Atg1RNAi side lagged behind the other side by an average of about 25 µm, a reproducible, significant, and large reduction in wound closure when syncytia formation was reduced (Figure 4E). These results suggest that syncytia speed wound closure. It is difficult, however, to disentangle the roles of autophagy and syncytia in this experiment, as the Atg1 knockdown cells may move more slowly because they are defective in the essential cellular function of autophagy and/or lacking syncytia. To address this question, we found a strong correlation between the size of the cell/syncytium and the distance it had moved from the start, measured at the time when the wound was half-closed. This correlation was true for cells/syncytia in control wounds as well as for those when Atg1 was knocked down (Figure 4F,G respectively). These data indicate that syncytia formed through cell fusion speed wound closure.

Syncytia outcompete mononucleate cells at the leading edge of repair

To understand this phenomenon better, we further investigated control wounds. Because about half the cells fused to form syncytia around wounds, we were able to compare the behavior of syncytial to non-syncytial cells within the same wound, providing a well-controlled environment for assessing how syncytia contribute to wound closure. In live imaging, we observed that syncytia frequently moved faster than mononucleated cells toward the wound, overtaking unfused cells as the wound closed. Figure 5A tracks four GFP-labeled cells near a wound, with cell 1 furthest from and cell 4 closest to the wound (Figure 5Ai). Middle cells 2 and 3 fuse with five others into a large syncytium, which advances toward the wound faster than both non-fusing cells, and by 400 min after wounding the syncytium extends well beyond the unfused cells (Figure 5Aiii). This behavior was evident even without GFP labeling: Figure 5B-C show a group of cells at the leading edge of the wound (the wound is recognized by the lack of p120ctn). At 90 min after wounding, this region of the leading edge comprises three mononuclear cells flanked on either side by syncytia (outlined in yellow); the middle of the three mononuclear cells is outlined in orange (5B) and white (5C). By 150 min, the syncytia have pushed out all three mononuclear cells from the leading edge, and at 360 min two of these mononuclear cells are visible about 20 µm back from the wound edge. Indeed, we observed that several hours after wounding, the entire leading edge was occupied by syncytia. To investigate how syncytia came to occupy this position at the front lines of wound healing, we expressed MyoIIGFP/Zip-GFP, which along with actin forms the contractile purse string and makes the leading edge visible, along with p120ctnRFP to label cell borders. We analyzed the persistence of all mononuclear and syncytial cells at the leading edge over the course of closure for three wounds, starting when the leading edge was first visible about 30 min after wounding (Figure 5D-G). At the start, 75% of the perimeter was occupied by syncytia with the rest occupied by 10-13 mononuclear cells (Figure 5D). As the wound closed, the syncytia became larger and displaced all the mononuclear cells, with the last mononuclear cell removed from the leading edge 20-160 min before closure (Figure 5E-G). No mononuclear cell persisted at the leading edge through wound closure, indicating that syncytia outcompete unfused cells to close the wound.

Syncytia outpace mononuclear cells.

A) Before wounding, two clusters of cells are labeled with GFP, 1-2 and 3-4 (white numbers). After wounding, cells 2 and 3 fuse with neighbors 5-9 to form a syncytium (dashed leading edge in Aiii), which advances toward the wound, passing unfused cell 4 (arrow). B-C) Unfused cell (outlined in orange and white) is replaced at leading edge by neighboring syncytia (S and yellow outline). D) Leading-edge perimeter was analyzed over time for three wounds, also analyzed in panel G. Percent leading edge occupied by syncytia increased over time to 100%. E) Images of sample 1 from graph D. Unfused cells (arrows) were tracked over the course of wound closure; 30 min after wounding, 10 unfused cells at the wound leading edge are indicated by arrows. F) At 120 min, the last unfused cell was ejected from leading edge (arrow). The wound closed at 140 min. G) The loss of unfused cells was analyzed over time in the three wounds from panel D. All unfused cells are excluded from leading edge by syncytia well before each wound closed 20-160 min later. Images are single Z slices in Ai-Aii and maximum intensity projections in Aiii-F. Scale bar for B,C shown in Civ and for E,F shown in F. Scale bars in A,E,F = 20 µm, B,C = 10 µm.

Radial border fusion reduces intercalation, the rate-limiting step of wound closure

Why do syncytia move faster? To address this question, we considered the geometry of fusion because different fusion orientations could provide different benefits for wound repair. We observed that fusions sometimes occur between two adjacent cells equidistant from the wound, as diagramed in the top panel of Figure 6A and exemplified in Figure 1Iiii-Iiv (cells 1,2). This breakdown of a radial border produces a syncytium elongated along the wound edge.

Cells fuse along different axes at different frequencies, reducing the number of wound-proximal cell intercalations.

A) Illustration of tangential vs radial border breakdown. B) More tangential than radial borders break down after wounding. Data are from four wounds, with a total of 235 border breakdowns: 39 radial and 196 tangential. Individual counts are shown per wound with mean and SEM. C) Fusions across radial borders reduce intercalation at a wound. D) Quantification of intercalations and fusions around the three wounds of Figure 5D,G.

Alternatively, fusion may occur between adjacent cells at different distances from the wound, as diagramed in the lower panel of Figure 6A and exemplified in Figure 1F (cells 3,4). This breakdown of a tangential border produces a spoke-like syncytium pointing into the wound. To analyze the frequency and timing of these two different axes of fusion, we calculated the orientations of all lost borders in four wounds. (The angle of the lost border can be determined for border breakdowns; it is difficult to determine for apical shrinking because the other fusion partner is hard to identify). In four analyzed wounds, we identified 235 border breakdowns: 39 radial borders and 196 tangential borders (Figure 6B). Thus, there were about five-fold more tangential borders lost to fusion than radial. Nonetheless, the two types had similar temporal and spatial distributions around the wounds (Figure 6—figure supplement 1).

As a wound closes, fewer cells can occupy the ever-smaller leading edge, necessitating a rearrangement of cell adhesions to allow cell intercalation, as depicted in the top of Figure 6C. Indeed, when cells do not fuse after wounding, as in Drosophila wing discs, the impact of intercalations on tissue fluidity is known to be a rate-limiting step of wound repair (Tetley et al., 2019). As shown in the bottom of Figure 6C, the requirement for cell intercalation would be reduced specifically by radial border fusion.

To consider the contribution of radial border fusions to wound closure, we took three approaches. First, we analyzed existing data from the three wounds shown in Figure 5D-G, and asked how many leading-edge cells fused radially vs. intercalated: 17-41% of cells removed from the leading edge were removed through fusion, reducing the burden of intercalation substantially (Figure 6D). Interestingly, the larger the percentage of cells that fused rather than intercalated, the faster the wound closed (Figure 6D).

Second, we analyzed tissue fluidity as measured by the cell shape index, ! = #⁄√%, where P is cell perimeter and A is cell area. This dimensionless index characterizes the cell jamming transition in vertex models of epithelia: for π < 3.81, a modeled isotropic epithelium is in a jammed state with energy barriers that limit intercalations, giving rise to a solid-like tissue; but for π > 3.81, the energy barriers vanish, yielding an unjammed fluid-like tissue (Bi et al., 2016; Park et al., 2015). In real epithelia, the solid-fluid transition may take place at different values of π (Wang et al., 2020), but higher π still implies greater tissue fluidity. We segmented images of the pupal notum before and after wounding and calculated the distributions of cell shape index for both syncytia and non-syncytial cells (Figure 7A-B). The mean cell shape index increased significantly after wounding with syncytia having a higher shape index than nearby post-wound non-syncytial cells (p = 0.001) and both having higher shape indices than pre-wound cells (p < 10-6). The post-wound epithelium is thus substantially more fluid-like than it was before wounding

Computational modeling indicates that fusion speeds the rate of wound closure because of reduced cell intercalation.

(A-B) Experimental observations of cell shape index (perimeter divided by square root of area). By two hours after wounding, the cell shape index has increased for both syncytia and non-syncytial cells within two rows of the wound margin. These increases indicate an increase in tissue fluidity. (A) Cells and syncytia were segmented, and cell shape index calculated, as shown for this image at 115 min after wounding; cell shape index indicated by color-coding. (B) Pre-and post-wound cell-shape distributions were compiled from 6 pupae. The red dashed line represents the critical cell shape parameter, 3.81, for the solid-to-fluid transition in vertex models of epithelia. Grey boxes span the interquartile range with a line at the median; whiskers extend out 1.5x the median-to-closest-quartile range; and outliers beyond the whiskers are plotted as individual points. Distributions include 1304 pre-wound mononucleate cells, 168 post-wound mononucleate cells, and 52 post-wound syncytia; **** denotes p < 10-6 level; *** denotes p < 0.001. (C-E) Vertex model simulations of wound closure with cell fusion allowed or suppressed. When the spatial and temporal probabilities for cell fusion match experimental observations, the model yielded 55 ± 22 fusions per wound (mean ± standard deviation), with 15 ± 6 of these occurring at radially aligned edges. As shown in C, fusions in the model speed wound closure: four simulations per condition; shaded regions denote standard deviation; time normalized to the average time to closure when fusion is suppressed. Panels D-E show a series of still frames from a matched pair of simulations in which fusion is allowed or suppressed. Corresponding videos are available as Video 5.

Third, we adapted the computational vertex model from Tetley et al. (2019) to include probabilistic cell-cell fusions to directly investigate the role of cell fusion in tissue fluidity and wound closure. We generated four initial cell configurations with similarly sized wounds and used these to run four matched pairs of simulations – one that proceeded to close without any cell fusions and the other that allowed fusions to occur with spatiotemporal probabilities that matched our experimental observations (Figure 6B, Figure 6—figure supplement 1A). On average, the probabilistic fusion-allowed simulations had 68 ± 14 (mean ± SD) fusions and closed the wound in about 1/3 less time than those without fusions (Figure 7C). Selected frames from one of the matched simulation pairs are shown in Figure 7D-E and full videos of this pair are available as Video 5. Importantly, the modeled syncytia had no special properties: same contractility as other cells, same target shape index, and same target area as the sum of the fused pair. Just the presence of these larger cells reduced the need for intercalations during wound closure and thus sped up the process.

Tangential border fusions allow resources from distant cells to be mobilized to the wound edge

In contrast to radial border fusions, tangential border fusions might provide a way for cellular resources that would be trapped in distal cells to move toward the wound to contribute to closure. To test this hypothesis, we generated small flip-out clones expressing actin-GFP and wounded such that unlabeled cells intervened between the labeled cell and the leading edge (Figure 8). Like cytoplasmic GFP in Figures 1I and 2C, we observed actin-GFP to equilibrate between cells soon after fusing. For syncytia that did not have access to the leading edge (n =2), actin-GFP levels remained uniform (Figure 8A,B). In contrast, syncytia with access to the leading edge (n = 10) first uniformly distributed actin (Figure 8Cii), but once the leading edge was contacted, they redistributed actin-GFP to it (Figure 8C,D). Kymographs of actin-GFP confirm that regardless of location, actin equilibrates between fusing cells within 5 minutes (Figure 8Biii, Div); however, nearly all actin that originated in the distal cell is redistributed to the leading edge in a syncytium positioned there (Figure 8Div). In one striking instance, actin-GFP appeared to travel through three cells to arrive at the leading edge from its initial location three rows back (Figure 8E-G, Video 6). Importantly, although actin is labeled from only one of the fusing cells, it likely represents the location of actin from all fusing cells, explaining why syncytia are better able to occupy the leading edge. We envision that other resources would also be concentrated in subcellular regions as needed by syncytia. This ability of syncytia to concentrate pooled actin to the leading edge may allow syncytia to outcompete their smaller mononuclear neighbors and close wounds more rapidly.

Syncytia concentrate pooled resources at the leading edge.

Random scattered cells expressing Actin-GFP were generated by heat-shock mediated flip-out expression of Gal4. A-B) Labeled actin is expressed in cell 1 before wounding (Ai,Bi). These cells are not close to the wound (inset in Aii). By 28 min after wounding, actin-GFP equilibrated between cells 1 and 2 (Aii,Bii), demonstrating cytoplasmic fusion. The resulting syncytium had no access to the leading edge, and actin remained equilibrated, as shown in the kymograph (Biii) generated from actin-GFP intensity over time at the yellow line in Bii. Fusion was confirmed at 180 min after wounding by apical shrinking of cell 2 (not shown). C-D) Labeled actin is expressed in cell 1 before wounding (Ci,Di) and equilibrates between cells 1 and 2 by 6 minutes after wounding, demonstrating cytoplasmic fusion (Cii,Dii). The resulting syncytium contacts the leading edge, and by 28 min after wounding actin from cell 1 is redistributed to the wound margin (Ciii,Diii), as shown in the kymograph (Div) of actin intensity over time at the yellow line in Diii. Fusion was confirmed at 55 min after wounding, by apical shrinking of cell 1 (not shown). E-F) Before wounding, actin-GFP in cell 1 is three cells away from the future leading edge. After wounding, fusion of cells 1-4 allows actin-GFP to be subcellularly localized to the leading-edge actin cable. Border breakdown is visible between cells 1-2 at 30 min after wounding (Ev) and apical shrinking occurs later (see Video 6). G) Mean profile plot of actin-GFP comparing the syncytia in Fv at 30 min after wounding (dark line) with the cells in Fi before wounding (dotted line) demonstrating that nearly all actin-GFP has been relocalized to the leading edge from its starting position 20-30 µm away. Single Z slices for Ei - ii, Fi - ii; maximum intensity projections for A-D, Eiii - v, Fiii - v. Scale bar for panels in A-B shown in Bii, 10 µm. Scale bar for panels in C-D shown in Diii, 5 µm. Scale bar for panels E-F shown in Fv, 10 µm.

Discussion

Previous work established that polyploid cells – both cells with multiple nuclei and cells with enlarged nuclei – are important for closing epithelial wounds in post-mitotic tissues (Losick et al., 2013). Here we have used live-imaging to analyze the response to wounds of a diploid mitotic tissue, the epithelial monolayer of the pupal notum. We found that syncytia form around wounds by cell fusion at a remarkably high rate, with almost half the epithelial cells within 5 cells of the wound fusing with neighbors over the course of hours. Syncytial size increases with wound size, indicating syncytia formation is a dynamic and scalable response to wounding.

Several lines of evidence demonstrate that syncytia are faster at closing wounds than unfused cells. Knockdown of the autophagy gene Atg1 reduces fusion around pupal wounds as was previously reported for larval wounds; Atg1 knockdown also reduces the rate of wound closure. In both Atg1 knockdowns and in controls, there was a consistent observation that the bigger the syncytium, the faster it moved. Within control wounds, syncytia completely displace unfused cells of the same genotype at the leading edge of the wound, such that wounds are closed entirely by syncytia, even though there is a smaller number of syncytia than unfused cells. Finally, computational modeling of wound closure with and without cell fusion shows that wounds with fusion-based syncytia close faster because of the reduced need for cell intercalations.

Thus, wound-induced polyploidy is a generalized strategy for wound closure, used by both mitotic and non-mitotic tissues. Although one might have wondered if polyploidy was a wound response limited to tissues that cannot use mitosis to increase nuclear content, both endoreplication and cell fusion into syncytia occur after wounding in this mitotically active diploid tissue (Wang et al., 2015; White et al., 2024; this study). The requirement for Atg1 for fusion in both polyploid larval tissue and in diploid pupal tissue suggests that the fusion mechanisms are common, despite the altered cell cycle of the starting cells. Further, our finding that larger syncytia travel faster suggests that syncytia – and by extension perhaps cell enlargement via endoreplication –occur as adaptations to close wounds faster.

We identified two unique properties of syncytia that may improve their wound-closing ability. First, syncytia are larger than neighboring diploid cells and their presence enhances tissue fluidity. Previous studies have determined that the rate-limiting step in epithelial wound healing is the process of cell intercalation or rearrangement (Tetley et al., 2019). As the wound closes and the leading edge encircling the wound gets smaller, fewer cells can occupy the leading edge. The cells leaving the leading edge must remodel their cellular junctions, requiring time and energy. Overall, tissue solidity slows closure, so enhancing fluidity promotes faster closure. By fusing into larger syncytia, cells reduce the need for adhesion remodeling and intercalation and instead just change shape as the leading edge becomes smaller. Moreover, the resulting larger cells can have significantly more fluidity and less epithelial tension than their diploid progenitors, as recently demonstrated in syncytia formed by age-induced epidermal cell fusion (Dehn et al., 2023). As our computational modeling demonstrates, fusion can speed wound closure by ∼33% even when the resulting syncytia have no special properties besides being larger than neighboring diploid cells. Making the syncytia individually more fluid would be expected to speed closure even more.

The second unique property of syncytia is their ability to pool and concentrate the cellular resources of the many constituent cells into a subcellular location within the large syncytium. We explored this concept by monitoring the localization of actin within syncytia. Actin is important for wound healing, as epithelial wounds close through the action of two actin-dependent processes: the cinching of a multicellular actin cable at the leading edge that surrounds the wound, sometimes called a purse string (Abreu-Blanco et al., 2012; Bement et al., 1993; Martin and Lewis, 1992; Tamada et al., 2007), and the protrusion of actin-rich projections into the wound (Abreu-Blanco et al., 2012; Farooqui and Fenteany, 2005). Pupal wounds have been shown to use both processes, with the actin cable developing early and actin protrusions appearing later in wound closure (Antunes et al., 2013). By labeling actin in individual cells (clones), we were able to follow it redistribution in syncytia after fusion. As expected, at first labeled actin diffuses and equilibrates throughout the new large cell. Remarkably however, when the syncytium is in contact with the leading edge, labeled actin is concentrated at the leading edge, even if the actin originated several cells away from the leading edge. Without fusion, the labeled actin would have been trapped in a distal cell without access to the leading edge, but after fusion, the syncytium can concentrate the actin of its many component cells at the wound front. If N represents the number of cells that fused, our results suggests that syncytia can apply up to N times more actin to the leading edge; considering that we observed syncytia with dozens of nuclei, this could represent a significant enhancement of actin at the leading edge. Increased actin might explain the ability of syncytia to outcompete diploid cells at the leading edge.

Presumably, other resources such as mitochondria and ribosomes could also be pooled and concentrated by syncytia at cellular locations where they promote wound healing. It is known that mitochondrial fragmentation is localized to the site of cellular injury (Horn et al., 2020), and that ribosomes are localized at the tips of severed neurons (Noma et al., 2017), and syncytia would have access to a larger pool of mitochondria and ribosomes than their component cells. By pooling the cellular resources of component cells, syncytia may also allow lethally damaged cells to survive by providing them with needed survival factors originating in cells further from the wound. Thus, the concept of resource sharing that we demonstrate with actin may have ramifications for many resources. Interestingly, it has been demonstrated that wound-induced endoreplicating cells also increase protein levels of some wound-repair resources, such non-muscle myosin, presumably possible because of the many copies present in the polyploid genome (Losick and Duhaime, 2021).

Two visibly distinct cellular behaviors attended fusion in the apical plane of the adherens junctions: border breakdowns and apical shrinking. Both processes were confirmed as fusion by expressing GFP in clones and analyzing GFP sharing at fusion. Border breakdowns occurred sooner after wounding than apical shrinking and appeared to be a faster process, as apical shrinking started later and took hours to complete. It is unclear whether these differences in appearance and timing indicate different mechanisms, as both require the autophagy gene Atg1. For border-breakdown, it is simple to envision that fusion initiates near the apical adherens junctions, which are under more tension than the basolateral membranes, and this may explain why fusion proceeds more quickly there. We recently found that although epithelial tension drops after laser wounding, it is restored within about 10 min (Han et al., 2023), consistent with the timing of apical border breakdown. Also consistent with this model, in the C. elegans hypodermis, developmental cell fusion has been observed to initiate at the apical side and travel in the apical-to-basal direction (Mohler et al., 1998). It is also possible that fusion initiates at a more basal site, and that fusion proceeds apically, with adherens junction disassembly occurring later. Regardless of the site of fusion initiation, fusion seems to be a relatively equal process when border breakdown is observed, with the resulting syncytium occupying the same footprint that the two fusion partners did, at least initially. For apical shrinking, however, one cell appears to transfer its cytoplasmic contents and nucleus into another recipient cell, with both cytoplasm and nucleus moving laterally to the location of the fusion recipient, while the apical surface of the donor cell shrinks. It is possible that apical shrinking may indicate a different location for the initial fusion pore below the plane of the adherens junctions; it is also possible that this may reflect a difference in the starting pressure of the two fusion partners, so that one subsumes the other.

It is unclear what triggers either border breakdown or apical shrinking fusion. Many developmentally programmed cell fusions are mediated by fusogens, cell surface proteins that bring opposing membranes into close contact with each other, as in the C. elegans hypodermis (Chernomordik and Kozlov, 2008; Iosilevskii and Podbilewicz, 2021; Markvoort and Marrink, 2011; Mohler et al., 2002; Podbilewicz et al., 2006; Sapir et al., 2007; Shemer et al., 2004). For other cell fusions, the fusogen is elusive and may not exist, as none has been identified yet in Drosophila myoblast fusions, which occur when a fusion competent myoblast generates actin-rich podosome-like membrane protrusions that invade a founder cell (Kim et al., 2015b; Lee and Chen, 2019; Petrany and Millay, 2019; Rushton et al., 1995; Sens et al., 2010). Wound-induced fusion may be a response to the plasma membrane damage that occurs around wounds; indeed, plasma membrane damage has been documented around both laser wounds and puncture wounds (McNeil and Steinhardt, 2003; Shannon et al., 2017). Interestingly, in the large syncytial cells of the C. elegans hypodermis, plasma membrane repair after a puncture wound requires the fusogen EFF-1 (Meng et al., 2020).

Polyploidy as a wound response has begun to get increased recognition. In adult Drosophila, epithelial puncture wounds are repaired by both endoreplication and syncytia formation (Bailey et al., 2020; Besen-McNally et al., 2021; Grendler et al., 2019; Losick, 2016; Losick et al., 2013; Losick et al., 2016). In the zebrafish epicardium, genetic ablation is repaired by a wavefront of multinucleated polyploid cells formed by endomitosis, and these lead diploid cells to encompass the heart (Cao et al., 2017). In adult mammalian cardiomyocytes, polyploidy may be an adaptive response to maintain growth after the cardiomyocytes lose their ability to complete mitosis. Mouse cornea endothelial cells endoreplicate to increase polyploidy to restore tissue ploidy following genetic ablation (Losick et al., 2016). Mammalian hepatocytes are known to become increasingly polyploid with age (Carriere, 1969; Wheatley, 2008) and in response to various types of injury and disease (Gentric et al., 2015; Madra et al., 1995; Muramatsu et al., 2000; Sigal et al., 1999; Toyoda et al., 2005; Wilkinson et al., 2019). All mechanisms that promote polyploidy – fusion, endomitosis, endoreplication – result in larger cells with the potential to localize more resources; of these, fusion would act the fastest after wounding because there is no need for DNA replication. Interestingly, in Drosophila embryos, wounds induce the surrounding cells to become larger by increasing their volume alone and not their ploidy (Scepanovic et al., 2021), suggesting that simply an increase in size is important. Although many examples of wound-induced polyploidy exist, it is still likely to be an underreported phenomenon, as endpoint analysis might miss a transient polyploid response to injury; live imaging is the surest way to identify a polyploid wounding response.

Another polyploid response to injury can occur after surgical implantation of biomaterials for the purpose of guiding regeneration, as reviewed previously (Al-Maawi et al., 2017). In some cases, only mononuclear cells of the immune system respond to the implant, and in these cases the biomaterial is integrated into the body; in other cases, the implant triggers the fusion of immune cells into multinucleated giant cells, and in these cases the material is degraded and rejected. These multinucleated giant cells seem to share properties with the syncytia of the pupal notum, as they are formed by fusion in response to an environmental trigger and they have an aggressive ability to protect the animal in response to wounding. As these studies highlight, understanding the formation, maintenance, and regulation of polyploid cells may improve our ability to successfully implant biomaterials to aid tissue regeneration.

It is often noted that wound responses are similar to cancer cell behaviors. This similarity extends to wound-induced syncytia and their counterparts, polyploid giant cancer cells, as both cell types are highly aggressive and invasive. Chemotherapeutics induce the formation of polyploid giant cancer cells (Illidge et al., 2000; Mosieniak et al., 2015; Ogden et al., 2015; Wang et al., 2013), and some studies indicate that they can form through cell-cell fusion in tumors (Melzer et al., 2018; Noubissi et al., 2015; Powell et al., 2011; Song et al., 2021; Zhang et al., 2021). Once formed polyploid giant cancer cells are hypothesized to escape further chemotherapy treatments due to increased resistance to genotoxic stress (Weihua et al., 2011).

These polyploid giant cancer cells and their progeny also exhibit increased migration and invasion potential (Qu et al., 2013; Zhao et al., 2021). The parallels between the behaviors of polyploid giant cancer cells and the wound-induced syncytia of the pupal notum highlight the importance of understanding wound induced syncytia formation in a highly reproducible system, as a basic understanding of how these cells form in the Drosophila notum could inform how they become dysregulated in cancer.

Experimental model and subject details

Drosophila melanogaster

Drosophila lines used in this study are in Table S1. All Drosophila lines were maintained on standard cornmeal-molasses media supplemented with dry yeast. All flies, except those used in heat-shock flip clonal analysis, were raised at room temperature. For clonal analysis experiments flies were raised at 18 degrees Celsius until the 3rd instar stage when they were heat shocked in a circulating water bath at 37°C for 3 minutes. They were then allowed to develop at room temperature to 15-18hr after puparium formation (APF) before wounding experiments (described below) were conducted. To knock down Atg1, Atg1RNAi was crossed to the stock for internally controlled wounding analysis, expressing Ecad-GFP and TubP-Gal80ts ubiquitously and pnr-Gal4, UAS-nls-mCherry in the central region of the notum, similar to our previous system (O’Connor et al., 2021b). Crosses were moved to 29° for 6 days to allow expression of the RNAi construct and dissected between 12-15 hr APF.

Method details

Pupal mounting

At room temperature, white prepupae were identified and marked within plastic food vials. 15-18 APF pupae were removed from the vial onto a piece of double-sided tape (Scotch brand, catalogue #665) applied to a microscope slide. Using fine forceps, the anterior pupal case was removed exposing the head and notum of all pupae applied to the tape, as previously described (O’Connor et al., 2022; Shannon et al., 2017). The tape was carefully removed from the microscope slide and inverted onto a pre-prepared cover glass (Corning 2980-246, 24 mm x 64 mm) (O’Connor et al., 2022). The pupae were carefully pressed down onto the cover glass by adhering the section of tape above the pupal head. Once the notum was visibly pressed onto the cover glass an oxygen permeable membrane (YSI, standard membrane kit, cat#1329882) was applied to prevent the pupae from drying out during imaging.

Pupal survival

Following imaging, pupae were kept mounted as described above and allowed to continue to develop and eclose for 3-4 days. Pupae that continued developing until they were able to crawl out of the partially dissected case were classified as ‘survived’ and their data acquired form these samples were used for analysis. If a pupae did not survive to eclosion, the associated datasets were not used in the study.

Live imaging

Images were collected using a Nikon Ti2 Eclipse with X-light V2 spinning disc (Nikon, Tokyo, Japan) with a 40X 1.3 NA oil-immersion objective or 60X 1.4 NA oil-immersion objective.

Unless otherwise noted samples were imaged pre-wounding, immediately after wounding, every 2 min for 30 min, and then every 10 min for 6 h. Images were pre-processed in NIS-Elements using combinations of background subtraction, rolling ball correction, local contrast, and Denoise a.i. Assembly of figure panels was done using Affinity Designer and frames were centered on the entity in focus, compensating for frame shift due to wounding.

Laser ablation

A single pulse of a 3rd harmonic (355 nm) Q-switched Nd:YAG laser (5 ns pulse-width, Continuum Minilite II, Santa Clara, CA) was used for laser ablation. Laser pulse energies were kept to 1.9 μJ +/- 0.1 μJ, increased from our previous report (O’Connor et al., 2021b) to keep the wound size similar between old and new ablation rig.

Border breakdown and tangential vs radial assignment analysis

Individual border breakdown events were manually observed using FIJI by identifying syncytia late in the video and back-tracking to determine which borders broke down to form each syncytium. These borders were traced back to the first frame after wounding to develop the map on Fig 1G. For each border breakdown event, distance from the center of the wound and time after wounding were recorded in Microsoft Excel. These events were categorized as tangential or radial based on the orientation of the border-to-be-lost relative to a vector pointing outward from the center of the wound. Specifically, the line tool in FIJI was used to measure both the angle of the border with respect to the horizon, 8, and the angle of the line from the center of the wound to the center of the border with respect to the horizon, α. If |cos(, − .)| ≥ cos 45°, then the border was classified as radial; otherwise, the border was classified as tangential.

Shrinking cell initiation / duration analysis

Shrinking cells were identified in live microscopy videos by beginning at the end of the video and playing backwards in FIJI. Backwards, a shrinking cell appears to bloom from the epithelial layer, characterized by the expansion of a bright puncta of p120ctnRFP. Each cell that underwent this behavior was marked on a single frame of the video, then the distance from the center of the wound when shrinking started was denoted as well as the time that the shrinking started and completed. All cells that shrank were then manually traced back to the first frame after wounding to develop the map of shrinking cells (Fig 2E).

Approximating nuclei per syncytia

Nuclei and cell borders do not align in Z-projections of the pupal notum as the cells are non-prismatic. To estimate the number of nuclei per syncytium, the pre-wounding density of nuclei per unit area was determined for the circular region where syncytia form after wounding. Next, the apical area of the three largest syncytia/cells around a wound was measured at 0h, 1 h and 2 h post wounding. The area of each syncytia was multiplied by the nuclear density to yield the approximate number of nuclei per syncytia. The number of nuclei in each of the three syncytia was averaged to give a value for each of three samples in Figure 1B.

Measuring apical area of syncytia across varied wound sizes

The three largest syncytia were determined by eye in FIJI for six samples, three ablated at 1.9 µJ and three ablated at 3 µJ ablation. The apical area of syncytia was measured 3 h post wounding using the p120ctnRFP signal. Initial wound size was calculated by measuring Myosin II marked leading edge when it became apparent 30-160 min post wounding.

Unfused cells at leading edge: count and percent analysis

Unfused cells at the leading edge were identified using FIJI by a lack of border breakdowns. Each unfused cell was manually observed over the duration of wound closure and the time at which it departed from the leading edge was noted. A count of unfused cells at the leading edge was created in Excel and formal graphs were generated using Prism 9. To measure percent of the leading edge comprised of unfused vs. syncytial cells, each unfused cell’s leading edge contact was measured in FIJI using the polygon line tool. The total circumference of the wound was measured using the polygon line tool and unfused cell measurements were subtracted from the total to infer the syncytial occupancy at the leading edge. Formal histograms were generated using Prism 9.

Analyzing GFP labeled cells

107 individually labeled GFP cells were analyzed across 5 wounds over 6.5 h. The position of the ablation was optimized to place as many individually labeled cells within 40-80 µm from the center of the wound as possible. After wounding it was possible to identify a mixing event by the decrease in intensity from the source cell with a corresponding increase in intensity of a previously unlabeled neighbor. Intensity differences made it possible to distinguish instances where two source cells were adjacent to each other but only one had a mixing event. However, large patches of source cells were not evaluated because inter-patch mixing was not distinguishable. To evaluate if border breakdowns were preceded by mixing, the 11 individually labeled cells that had border breakdowns were tracked back to the start of the video and confirmed to have a mixing event. There was never an instance where a labeled source cell had a border breakdown without a prior mixing event occurring.

Wound closure analysis

A pigmented scar forms at the site of laser ablation making identifying the exact moment a wound is closed difficult. Since the scar is approximately the same size in each sample, the time point at which the Myosin II signal disappeared below the scar was used as a proxy for closure.

Fusion analysis

A cell border was counted as a location of wound-induced fusion if it met three criteria: (1) adherens junctions along the cell border were lost; (2) the lost adherens junctions were not restored; and (3) the loss of adherens junctions was accompanied by a morphological change of the neighboring cells, for example, the moving apart of tri-cellular junctions previously connected by the lost cell border. Since border-breakdown fusions are observed most frequently within 1 hr after wounding, we counted all such fusions occurring within 1 hr 40 min of wounding.

When comparing the fusion frequency in pnr versus control domains, we used a landscape-oriented rectangle centered on the wound with its height defined by the inner radius of the zone of nuclear membrane damage visualized by the release of nls-mCherry from the nucleus, which is approximately the same as the zone of lysis (O’Connor et al., 2021a).

Replacing the annular region with an extended rectangle excludes possible heterotypic fusions along the pnr-control border just above and below the wound. To calculate the frequency of fusion, in addition to the raw number of fusion borders, the total number of cell borders was also estimated by area-based scaling of the border count within a randomly selected 60-x-60-pixel (16.8-x-16.8-µm) region in the pre-wound image.

Since fusions can also appear as apical shrinking rather than border breakdown, we also quantified the number of apical-shrinking cell fusions in pnr versus control domains. Apical-shrinking cell fusions were defined as diploid cells that decreased in surface area over time and fully disappeared within a 5-hr movie. As with fusion analysis, cells near the pnr-control border were excluded. Apical shrinking cells were counted manually and their spatial locations marked using FIJI’s “Multi-point” tool.

Determining the effect of syncytia on wound closure

The distance between the leading edge and the wound center over time was used to calculate the speed of wound closure. The distance was calculated using the coordinates of the wound center and fixed points on the leading edge. To identify fixed positions on the leading edge that are unbiased and consistent over time, four reference lines, all crossing the center of the wound, were drawn using the red channel with only the nuc-mCherry signal: line 1 along the edge of the pnr domain labeled by nuc-mCherry, line 2 perpendicular to line 1, and line 3 and 4 are 30° from line 2. These references lines were then overlaid on the green channel with the E-cadherin signal.

To determine the speed of wound closure, the intersection of E-cadherin marked leading edge and reference lines 2-4 were used as the position of the leading edge. That is, at each time point, a total of 6 positions of the leading edge were recorded, among which 3 will be in the pnr domain and the other 3 in the ctrl domain. The average distance (d) between the leading edge and the center of the wound were calculated at each time point for both the pnr domain (dpnr) and the control domain (dctrl). The difference between control and pnr distance (Δd = dpnr – dctrl) was calculated over time for both the control pupae with no genetic manipulation in the pnr domain (pnr > +) and experimental animals with Atg1 knocked down in the pnr domain (pnr > Atg1RNAi). To determine whether Δd is statistically different between control and experimental pupae at each time point, two-way ANOVA (fit using full model) was used, comparing each cell mean with the other cell mean in that row.

To determine the distance syncytia covered during wound healing, syncytia were analyzed (1) at the leading edge when wound is half closed and (2) within the bow-tie region between reference line 3 and 4. ROIs were drawn along syncytia of interest using the “Freehand selections” tool of FIJI. In addition to syncytia location when wound is half closed, the starting position of cells that later fuse to form syncytia were also recorded in live-imaging videos, tracking the syncytia back in time. In addition to area, the coordinates of ROIs outlining syncytia at different time points were obtained using a self-written FIJI Macro. Within each ROI, the coordinate closest to the center of the wound was recorded using python, as the location of the syncytia. The distance a certain syncytium moved was calculated by subtracting the distance between the syncytium and the wound center when the wound was half closed from the distance between the syncytium and the wound center immediately after injury.

Calculating intercalation in live imaging

The change in the number of cells when the leading edge forms at 30 min (Nstart) to when the wound is closed (Nend) is equal to the number of intercalations plus the number of radial fusion events. Thus, (intercalations = ΔN – radial fusions). For the same three samples used in Fig 4D,G, we determined ΔN from Nstart at 30 min and Nend when the wound had closed. Radial fusions were tallied by manually observing border breakdown events between the leading-edge cells and intercalations were calculated. Each radial fusion was counted as one prevented intercalation event.

Calculating cell shape index

Images of pupae labeled with Ecad-GFP were segmented into individual cells with Cellpose (Stringer et al., 2021); cells that were not readily segmented were excluded from subsequent analysis. To avoid perimeter inflation due to pixelation, the boundaries of segmented cells were smoothed by taking the midpoint between every other pair of boundary pixels. Smoothing and calculation of perimeters, areas and cell shape indices (#⁄√%) were performed in Mathematica (Wolfram Research Inc., Champaign, IL).

Vertex models with cell fusion

Simulations of wound healing were run using vertex models implemented in Mathematica (Wolfram Research Inc., Champaign, IL) following the methods of Tetley et al. (2019)and using parameters that match experimental observations from the Drosophila pupal notum as detailed in Han et al. (2024) with the following exceptions: the mean line tension on each edge (Λ) was specified in terms of a target shape parameter (ρ0) such that Λ=−2ρ0ΓA0, where the contractility (Γ) and target cell area (A0) were taken from Han et al. (2024) and ρ0 = 3.745 was chosen so that the actual cell shape index matched experimental pre-wound measurements; and the line tension variability (αm) was decreased from 1.0 x 10-2 to 9.3 x 10-4 so that the intercalation rate in unwounded cell sheets matched experimental observations. Pre-wound configurations were generated by performing a Voronoi tessellation of a circular area with 320 randomly distributed seed points and then equilibrating the cell configuration without stochastic fluctuations until it reached a steady state. Wounds were introduced by removing all cell borders within a circular patch ∼8 cells across and removing the wound’s contributions to the energy Hamiltonian. To handle wound closure and fusion, the following additions were made to the model: (1) purse-string tension on the leading edge of the wound was incorporated as in Tetley et al. (2019) with a normalized strength of 0.115; (2) traction/crawling forces of cells on the leading edge of the wound were incorporated as a uniform tensile stress acting inwardly on the edges and vertices of the wound area (similar to the uniform stress applied to the outer edge of the patch to represent tension from surrounding cells); this inward stress had a normalized strength of Σin⁄KA0 = 0.13, where K is the area elasticity modulus; (3) the cell-cell interfaces at which fusion would occur were selected stochastically with probabilities that depended on time, distance from the wound, and the radial or tangential orientation of the interface – all chosen to match experimental observations reported here; and (4) the fusion of two cells was handled by eliminating the interface between those cells and setting the target area of the fused cell to the sum of the target areas of its two progenitors. All cells, fused or not, had the same contractility and target shape parameter.

Kymograph and plot profile analysis

Actin-GFP intensity was analyzed using the kymograph tool in FIJI after drawing a 11-pixel line through the middle of the syncytia. Profile plot values were exported from NIS elements to Excel and graphs were generated using Prism.

Resource availability

Lead contact

Requests for fly lines, reagents, and additional questions should be directed to Dr. Andrea Page-McCaw (andrea.page-mccaw@vanderbilt.edu).

Materials availability

Fly lines generated in this study are available from the Bloomington Drosophila Stock Center or from the lead contact.

eLife Assessment

This important work addresses a very relevant biological question: what is the cellular basis of wound healing? Using the Drosophila pupal notum as a model, the paper provides an elegant, thorough, descriptive characterization of syncytia-driven wound closure using state-of-the-art confocal live imaging of the pupal notum. The authors meticulously characterize the cell-cell fusion events during wound healing and inhibit cell fusion to show to that it is necessary to speed wound closure. In addition, the study provides convincing evidence that cell fusion allows actin resources at be partitioned to the leading edge.

Reviewer #1 (Public Review):

Summary:

This study aims to understand how cell fusion contributes to wound healing using a laser-induced injury in the notum epithelium of a developing fruit fly. The authors meticulously characterize the epithelial fusion events using a live imaging approach and report that syncytia arise by 'border breakdown' and 'cell shrinking'. The syncytial epithelial cells also appear to outcompete mononucleated cells and preferentially dissolve their tangential borders, which correlates with the accumulation of actin at the leading edge.

Strengths:

The strength of this study is the authors' live imaging approach to capture these dynamic fusion events that are a fundamental yet poorly understood biological process.

Comments on revised version.

The manuscript overall is significantly improved and authors addressed majority of my concerns. The addition of the computational vertex model (Figure 7) as well as Atg1 RNAi (Figure 4) to inhibit cell fusion provide more mechanistic insight to their study. However, the analysis of Atg1 RNAi wound assay falls short as it does directly measure changes in syncytium frequency nor size to confirm that cell fusion is reduced. The authors should quantify the number of nuclei per syncytium over the 2hr wound healing period as performed for WT in Figure 1C. It would have been ideal if they could have also performed the Act-GFP spreading assay in WT and Atg1 RNAi strains to determine if Act-GFP movement is dependent on cell fusion as purposed. At the least, further quantification of Atg1 RNAi phenotype is warranted to support their conclusions.

Reviewer #2 (Public Review):

Summary:

Overall, this study provides a thorough description of the formation of syncytia following wounding of the proliferation-competent diploid epithelium of the pupal notum. While this phenomenon has already been described briefly for this particular tissue by the Galko lab in Wang et al 2015, the authors provide a much more detailed description and characterisation of the process providing some novel insights (radial versus tangential border breakdown, cell shrinkage, timings, syncytia outcompeting mononucleated cells, etc.).

Strengths:

This paper provides an elegant, thorough, descriptive characterisation of syncytia-driven wound closure using state-of-the-art confocal live imaging of the pupal notum. The authors show that laser-induced wounding of this diploid, proliferation-competent epithelium results in the formation of syncytia of various sizes in the first few cell rows around the wound edge, which progressively become bigger as healing proceeds. This results in ~50% of cells becoming part of these syncytia. The cell fusion events were convincingly demonstrated by showing the disappearance of p120ctnRFP and E-Cadherin-GFP from cell-cell borders as well as cytoplasmic GFP mixing of GFP-positive cells with a GFP-negative cell.

Apart from cell-cell fusion by border breakdown that mostly happens in the first 2h following wounding, the authors also found that at later stages of wound healing cell shrinkage following cytoplasmic mixing contributed to syncytia formation.

Next, the authors provided some convincing evidence that syncytia outcompete mononuclear cells for being positioned in the first cell row around the wound.

The authors then show that radial border breakdown occurs much less frequently than tangential border breakdown. They suggest that radial border breakdown reduces the requirement for cell-cell intercalations. They also hypothesise that tangential border breakdown might allow fused cells to share resources and provide more resources to be used near the wound edge, e.g. for actomyosin cable formation. To test this, the authors generate single-cell clones that overexpress Actin-GFP. They then show convincingly how a single Actin-GFP-positive cell in the second cell row fuses with one GFP-negative cell in the first cell row. The Actin-GFP signal then spreads in the fused cell and labels some previously unlabelled actin-rich structure near the wound edge which most likely is the actomyosin cable. This provides some evidence for resource sharing by cytoplasmic mixing following fusion.

Comments on revised version:

The authors have extended their original manuscript by adding two key parts. First, they show a role of Atg1 in mediating cell fusion (Figure 4). Second, they provide additional evidence for a contribution of radial border fusions to wound closure through its effect on tissue fluidity and through computational modelling (Figure 7).

This new version of the manuscript is greatly improved and provides significant new insights into the role of syncytia in aiding wound repair. There are just a few minor, yet important, additions needed to back up Figure 4 which should not require new experiments.

Minor but important points:

The authors show a role of Atg1 in mediating syncytia formation in Figure 4. However, since the Pnr>+ side of the wound closes slower than the non-Pnr side (control side), a few additions to this figure would be important and should not require additional experiments.

(1) The authors should show, similar to the data shown in Figure 4D of the wound radius over time for control versus Pnr>Atg1RNAi, also the same type of data for control versus Pnr>+.

(2) Since Pnr>+ also slows down wound healing, albeit to a lesser extent than Pnr>Atg1, the authors should also show an extra graph that provides evidence that Pnr>Atg1RNAi reduces syncytia formation more than Pnr>+ does. E.g. Two graphs could be added that show individual cell size at 4 or 5h post wounding for control versus Pnr>Atg1RNAi as well as for control versus Pnr>+ and also another graph with the same data but comparing cell size between Pnr>+ and Pnr>Atg1RNAi. Otherwise, if the expected minimum cell size for a syncytium is easy to estimate, a graph could be added that shows the percentage of cells that are above this threshold (e.g. above 100 square micron) for control versus Pnr>Atg1RNAi and control versus Pnr>+ and Pnr>+ versus Pnr>Atg1RNAi.

Reviewer #3 (Public Review):

In this revised manuscript, White et al. aimed to understand the wound-induced syncytia formation behavior in wound repair of Drosophila melanogaster pupal notum. For this purpose, the authors characterized two different types of adherens junctions' outcomes during syncytia formation around the wound region - border breakdown versus apical shrinking which appear to happen in different time points and for different time durations. The authors characterized cell-cell fusion events using cytoplasmic, junctional and nuclear markers. They determined that about half of the cells within 70 um radii from the wound undergo cell-cell fusion. They studied wound induction on the border between control epithelia and pnr domain suggesting that Atg1 is required for post-wound syncytia formation and wound closure. They showed that during wound closure syncytia gradually invade the wound leading edge mostly by radial fusion events. The data suggests that intercalation of cells from the leading edge slows down the wound closure process. They propose that cell fluidity of syncytial cells plays a role in wound closure speed. Finally, the authors showed that actin is concentrated to the front edge of syncytia located in the wound leading edge. The authors described some aspects of syncytia formation during wound closure using different approaches. Some clarifications are needed as described below.

Major suggestions:

(1) Introduction, page 4. The examples of developmental syncytia formation of invertebrates and vertebrates are confusing. The authors may want to make the examples clear and add additional examples. Currently, readers may assume that C. elegans cell fusions occur only in the hypodermis - other structures can be mentioned like the vulva, pharyngeal muscles, glia, tail. In addition, the authors may want to add injury-induced fusions like the C. elegans' PLM and PVD neurons (Ghosh-Roy et al., 2010; Newman et al., 2015; Oren-Suissa et al., 2017).

(2) In cases where it is not clear whether fusion has occurred or whether mononucleated cells were ejected from the leading edge, membrane markers can be used. Page 6. Lines 96-99. The authors may want to use a membrane marker like RFP-PH driven by the epithelial cell promoter.

(3) Pages 8-10. The authors may want to clearly explain that apical junctions shrinking is a post fusion event. That the apical shrinking is caused by the expansion of fusion pores and the migration of apical junctions towards the basolateral domain. This is something that was clearly shown during physiological epidermal cell-cell fusion in C. elegans by Mohler et al., 1998 and 2002. A cartoon showing the process of cell-cell fusion, pore expansion and apical junction dynamics would make the manuscript much clearer.

(4) Page 9. Line 170. "...as these cells represent fusion initiation events (fusion pore) but were unable to productively stabilize and expand the site of fusion and so returned to the diploid state." The authors may want to make clear that this is an assumption that needs to be tested. Live imaging using a membrane marker may resolve whether a reversible fusion pore was generated.

(5) Page 11. It is not clear whether Atg1 is directly required for cell fusion, or that autophagy is required for efficient cell fusion or both Atg1 and autophagy participate in the fusion process.

(6) Page 12. Line 235. "Indeed, we observed that several hours after wounding, the entire leading edge was occupied by syncytia." This observation is based only on the adherens junction marker. Can they test basal cell membrane marker? Is it possible that the mononucleate cell in the leading edge is under the two syncytia?

Author response:

The following is the authors’ response to the previous reviews

Public Reviews:

Reviewer #1 (Public Review):

Summary:

This study aims to understand how cell fusion contributes to wound healing using a laser-induced injury in the notum epithelium of a developing fruit fly. The authors meticulously characterize the epithelial fusion events using a live imaging approach and report that syncytia arise by 'border breakdown' and 'cell shrinking'. The syncytial epithelial cells also appear to outcompete mononucleated cells and preferentially dissolve their tangential borders, which correlates with the accumulation of actin at the leading edge.

Strengths:

The strength of this study is the authors' live imaging approach to capture these dynamic fusion events that are a fundamental yet poorly understood biological process.

Comments on revised version.

The manuscript overall is significantly improved and authors addressed majority of my concerns. The addition of the computational vertex model (Figure 7) as well as Atg1 RNAi (Figure 4) to inhibit cell fusion provide more mechanistic insight to their study. However, the analysis of Atg1 RNAi wound assay falls short as it does directly measure changes in syncytium frequency nor size to confirm that cell fusion is reduced. The authors should quantify the number of nuclei per syncytium over the 2hr wound healing period as performed for WT in Figure 1C. It would have been ideal if they could have also performed the Act-GFP spreading assay in WT and Atg1 RNAi strains to determine if Act-GFP movement is dependent on cell fusion as purposed. At the least, further quantification of Atg1 RNAi phenotype is warranted to support their conclusions.

In response to the reviewer's comment, we have repeated the analysis of Fig 1C and generated a new panel, Fig. 4C, which is directly comparable to the control and shows that syncytial size is dramatically reduced in the Atg1 knockdown area. Unfortunately, we cannot perform the second analysis of actin-GFP spreading in the Atg1 knockdown cells because we need Gal4 for labeling individual cells and for knocking down Atg1, and we can't do both at the same time.

Reviewer #2 (Public Review):

Summary:

Overall, this study provides a thorough description of the formation of syncytia following wounding of the proliferation-competent diploid epithelium of the pupal notum. While this phenomenon has already been described briefly for this particular tissue by the Galko lab in Wang et al 2015, the authors provide a much more detailed description and characterisation of the process providing some novel insights (radial versus tangential border breakdown, cell shrinkage, timings, syncytia outcompeting mononucleated cells, etc.).

Strengths:

This paper provides an elegant, thorough, descriptive characterisation of syncytia-driven wound closure using state-of-the-art confocal live imaging of the pupal notum. The authors show that laser-induced wounding of this diploid, proliferation-competent epithelium results in the formation of syncytia of various sizes in the first few cell rows around the wound edge, which progressively become bigger as healing proceeds. This results in ~50% of cells becoming part of these syncytia. The cell fusion events were convincingly demonstrated by showing the disappearance of p120ctnRFP and E-Cadherin-GFP from cell-cell borders as well as cytoplasmic GFP mixing of GFP-positive cells with a GFP-negative cell.

Apart from cell-cell fusion by border breakdown that mostly happens in the first 2h following wounding, the authors also found that at later stages of wound healing cell shrinkage following cytoplasmic mixing contributed to syncytia formation.

Next, the authors provided some convincing evidence that syncytia outcompete mononuclear cells for being positioned in the first cell row around the wound.

The authors then show that radial border breakdown occurs much less frequently than tangential border breakdown. They suggest that radial border breakdown reduces the requirement for cell-cell intercalations. They also hypothesise that tangential border breakdown might allow fused cells to share resources and provide more resources to be used near the wound edge, e.g. for actomyosin cable formation. To test this, the authors generate single-cell clones that overexpress Actin-GFP. They then show convincingly how a single Actin-GFP-positive cell in the second cell row fuses with one GFP-negative cell in the first cell row. The Actin-GFP signal then spreads in the fused cell and labels some previously unlabelled actin-rich structure near the wound edge which most likely is the actomyosin cable. This provides some evidence for resource sharing by cytoplasmic mixing following fusion.

Comments on revised version:

The authors have extended their original manuscript by adding two key parts. First, they show a role of Atg1 in mediating cell fusion (Figure 4). Second, they provide additional evidence for a contribution of radial border fusions to wound closure through its effect on tissue fluidity and through computational modelling (Figure 7).

This new version of the manuscript is greatly improved and provides significant new insights into the role of syncytia in aiding wound repair. There are just a few minor, yet important, additions needed to back up Figure 4 which should not require new experiments.

Minor but important points:

The authors show a role of Atg1 in mediating syncytia formation in Figure 4. However, since the Pnr>+ side of the wound closes slower than the non-Pnr side (control side), a few additions to this figure would be important and should not require additional experiments.

(1) The authors should show, similar to the data shown in Figure 4D of the wound radius over time for control versus Pnr>Atg1RNAi, also the same type of data for control versus Pnr>+.

The data the reviewer requests is available in our bioRxiv manuscript, in Fig. 6B (Hua, Krystofiak, Pumford, Page-McCaw, and Hutson, https://doi.org/10.64898/2026.05.31.728998). These experiments were all done at the same time. As you can see, the difference in closure rate is quite subtle in control wounds.

(2) Since Pnr>+ also slows down wound healing, albeit to a lesser extent than Pnr>Atg1, the authors should also show an extra graph that provides evidence that Pnr>Atg1RNAi reduces syncytia formation more than Pnr>+ does. E.g. Two graphs could be added that show individual cell size at 4 or 5h post wounding for control versus Pnr>Atg1RNAi as well as for control versus Pnr>+ and also another graph with the same data but comparing cell size between Pnr>+ and Pnr>Atg1RNAi. Otherwise, if the expected minimum cell size for a syncytium is easy to estimate, a graph could be added that shows the percentage of cells that are above this threshold (e.g. above 100 square micron) for control versus Pnr>Atg1RNAi and control versus Pnr>+ and Pnr>+ versus Pnr>Atg1RNAi.

In response to this comment and the comment from reviewer 1, we have now added new Fig. 4C, which addresses the reviewer's question about the comparative frequency of fusion in pnr>Atg1RNAi and pnr>+. These graphs show that Atg1 knockdown significantly reduces the size of syncytia.

Reviewer #3 (Public Review):

In this revised manuscript, White et al. aimed to understand the wound-induced syncytia formation behavior in wound repair of Drosophila melanogaster pupal notum. For this purpose, the authors characterized two different types of adherens junctions' outcomes during syncytia formation around the wound region - border breakdown versus apical shrinking which appear to happen in different time points and for different time durations. The authors characterized cell-cell fusion events using cytoplasmic, junctional and nuclear markers. They determined that about half of the cells within 70 um radii from the wound undergo cell-cell fusion. They studied wound induction on the border between control epithelia and pnr domain suggesting that Atg1 is required for post-wound syncytia formation and wound closure. They showed that during wound closure syncytia gradually invade the wound leading edge mostly by radial fusion events. The data suggests that intercalation of cells from the leading edge slows down the wound closure process. They propose that cell fluidity of syncytial cells plays a role in wound closure speed. Finally, the authors showed that actin is concentrated to the front edge of syncytia located in the wound leading edge. The authors described some aspects of syncytia formation during wound closure using different approaches. Some clarifications are needed as described below.

Major suggestions:

(1) Introduction, page 4. The examples of developmental syncytia formation of invertebrates and vertebrates are confusing. The authors may want to make the examples clear and add additional examples. Currently, readers may assume that C. elegans cell fusions occur only in the hypodermis - other structures can be mentioned like the vulva, pharyngeal muscles, glia, tail. In addition, the authors may want to add injury-induced fusions like the C. elegans' PLM and PVD neurons (Ghosh-Roy et al., 2010; Newman et al., 2015; Oren-Suissa et al., 2017).

We appreciate the suggestions and have included the additional examples of C. elegans vulva and PLM and PVD neurons. We are limiting ourselves to those because we don't want to focus too heavily on C. elegans examples, as that's not the direction this paper is heading.

(2) In cases where it is not clear whether fusion has occurred or whether mononucleated cells were ejected from the leading edge, membrane markers can be used. Page 6. Lines 96-99. The authors may want to use a membrane marker like RFP-PH driven by the epithelial cell promoter.

At this point in the manuscript, we are introducing syncytia and are not concerned yet with their origin.

(3) Pages 8-10. The authors may want to clearly explain that apical junctions shrinking is a post fusion event. That the apical shrinking is caused by the expansion of fusion pores and the migration of apical junctions towards the basolateral domain. This is something that was clearly shown during physiological epidermal cell-cell fusion in C. elegans by Mohler et al., 1998 and 2002. A cartoon showing the process of cell-cell fusion, pore expansion and apical junction dynamics would make the manuscript much clearer.

Apical shrinking cannot be caused by the "migration of apical junctions towards the basolateral domain" because that is not what we observed -- rather, we observed labeled adherens junctions remaining at the apical surface while the area they enclose becomes smaller (shrinks). Further, despite close reading of the Mohler papers, it is not clear how similar the apical shrinking events of this manuscript are to the fusion events described there. Finally, we do not want to include a schematic describing this process because that would suggest certainty that we do not have. Unlike in C. elegans development, wound-induced cell fusion is stochastic, not stereotyped; with cells that display apical shrinking, the fusion partner of a labeled cell is difficult to identify because it is often not a neighboring cell. These factors make it difficult to describe this process in detail, but we have sufficient data to conclude that these are indeed cell fusion events.

(4) Page 9. Line 170. "...as these cells represent fusion initiation events (fusion pore) but were unable to productively stabilize and expand the site of fusion and so returned to the diploid state." The authors may want to make clear that this is an assumption that needs to be tested. Live imaging using a membrane marker may resolve whether a reversible fusion pore was generated.

Thank you for the suggestion; we have updated this text to make it clear that this is an interpretation.

(5) Page 11. It is not clear whether Atg1 is directly required for cell fusion, or that autophagy is required for efficient cell fusion or both Atg1 and autophagy participate in the fusion process.

Our data show that Atg1 is required for cell fusion. The work that inspired this experiment, Kakanj et al 2022, concluded from their more comprehensive studies that the process of autophagy was required. We have clarified the text.

(6) Page 12. Line 235. "Indeed, we observed that several hours after wounding, the entire leading edge was occupied by syncytia." This observation is based only on the adherens junction marker. Can they test basal cell membrane marker? Is it possible that the mononucleate cell in the leading edge is under the two syncytia?

Unfortunately, there are not good basal markers -- the recently reported basal spot markers also label adherens junctions. Nonetheless, we are confident that the mononuclear cell is not under the syncytia because we image Z-stacks and thus can detect cell overlap.

Recommendations for the authors:

Reviewer #3 (Recommendations For The Authors):

Minor suggestions:

(1) Figure 1. The authors may want to add an image immediately after laser ablation of the actual wound and the area around the wound. Add an arrow to mark the wound.

With this wounding modality, the extent of the wound is unclear for ~30 min. As we reported in O'Connor et al, PLoS One, 2021, there is a gradient of damage emanating out from the center of the wound, and cells with greater amounts of damage die while those with less damage repair and survive. Immediately after laser ablation, very little visible damage is evident by 120ctn-RFP and Histone-GFP (the markers in Fig. 1) until the cells die and the surrounding cells respond.

(2) Page 6. Line 86. "A mitotic tissue utilizes cell-cell fusions during wound repair." replace "during wound repair" with "after wound induction" since in this section the authors do not show that this process is part of wound repair.

Thank you for the suggestion - we reworded this heading to remove "wound repair".

(3) Page 6. Line 92. The authors may want to be consistent with the terms used in the text and in the figure - His2GFP in the text versus Histone GFP in the figures.

Thank you for the suggestion, we have revised for consistency.

(4) Figure 1 - supplement figure 1D. The "v" of Div panel moved below D.

Thank you, we have corrected it.

(5) Figure 1 - supplement figure 1G. add "i" to second Gii to make it Giii.

Thank you, we have corrected it.

(6) Page 24. Figure 1H legend. 3 or 4 wounds?

Thank you for catching this error - 4 wounds.

(7) Page 7. Line 124. "GFP mixing always preceded border breakdowns (n=11)" instead of "always" use "in all observed cases".

We have made this change.

(8) Figure 2. Switch the writing "Apical Shrinking: Nuclear Transfer" since apical shrinking represented in panel 2A and Nuclear Transfer in panel 2B. If this description applies only to panel 2B, make it clear.

We consider this heading to apply to panels A and B together (as they show the same sample, just different channels).

(9) Figure 2C. Is ActinGFP a cytoplasmic GFP driven by actin promoter or Actin-bound GFP? Cytoplasmic GFP versus membrane-cortex GFP?

It is a transgene expressing an actin-GFP fusion protein, as noted in the key reagents table and discussed in Fig. 8. We corrected the manuscript to ensure it is always referred to now as Actin-GFP in the text, figures, and legends.

(10) Video 3 - Impressive movie!

Thank you!

(11) Page 9. Line 155. "In both these cells, as the cell lost its basal volume, cytoplasm moved laterally to join the neighboring syncytia." It seems that the apical shrinking cells' cytoplasm joined the neighboring syncytia even before.

Because both indicated cells (yellow and white arrows) and the neighboring syncytium are all labeled with GFP, it is not possible to determine precisely when the cells' cytoplasm joined the syncytium.

(12) Page 9. Line 158. "...but fusions associated with apical shrinking occurred later and were more numerous." Did the fusion occur later or the apical shrinking itself as was mentioned before and shown in Figure 2F?

We have changed the wording, as for many apical shrinking events we cannot tell exactly when the fusions were initiated.

(13) Page 25. Figure 3A legend. What is the meaning of morphological fusion? Border breakdown and apical shrinking? The authors may want to define it.

We have defined it now in the legend.

(14) Page 26. Figure 3B-C legend. "Panel C shows that apical shrinking fusion and border-breakdown fusion occur at similar distances from the wound." It seems that fusion by apical shrinking mostly occurs within 60-70 um from wound center and fusion by breakdown occurs equally at all distances up to 80 um.

We don't disagree with your comment, but we feel the dataset is too small to make such a statement. The data is presented so the interested reader can make their own conclusion.

(15) Page 9. Line 165. "...but infrequently (n=3) with GFP mixing and no subsequent cell fusion..." Does this mean that there were GFP mixing without border breakdown or apical shrinking?

Yes, that is correct. We assume that in this case a fusion pore opened and then closed again. We have added a phrase to clarify.

(16) Page 9. Line 175. "...the spatial distribution of fusing cells that shrank vs. lost borders was similar (compare Figures 1G and 2E)." Even though the visual comparison suggests similar spatial distribution, the carefully quantified distribution in figure 3C suggests more fusion by shrinkage at 60-70 um from wound center of the 5 tested wounds.

As we noted to comment 14, we feel the data set is too small to make such a statement. The data is presented so the interested reader can make their own conclusion.

(17) Figure 3. The shown pies sum the results from 5 wounds. It would be interesting to add a graph comparing the percentage of fused and persisted cells per wound to see the variability, if exists.

Unfortunately, the number of fused/persisting cells in each wound is greatly affected by the heat-shock conditions that generate the labeled clones; even the ratio of these fates would be heavily influenced by noise because the numbers are small in each animal. Further, the frequency of fusion is determined by the wound size as shown in Fig. 1. Because of these variables, such data could be easily misinterpreted.

(18) Figure 3 - figure supplement 1D. Even though it was mentioned that the duration of some border breakdown is finished within minutes it is worth comparing it with shrinking duration on one graph.

Unlike apical shrinking, it is difficult to identify exactly when border breakdown concludes, so this data is difficult to compare. We have provided several examples of border breakdown in the manuscript that give an overview of the process.

(19) Video 1 is not mentioned in the main text.

Thank you for catching that omission. We now refer to it in the first paragraph of the results.

(20) Figure 4B. The difference between the treated group and the control group is unclear. Add arrows.

We have added some arrows to Fig. 4B.

(21) Figure 4C. For consistency use percentage for both border breakdown and shrinking cells.

In response to the reviewer's comment, we now provide the consistent metric of number of lost borders and number of shrinking cells.

(22) Page 11. Did the authors try other wound types (e.g. mechanical/chemical wounds)? May other wound causes besides laser ablation result in different response? This may help to answer whether there is a causation between syncytia formation and speed wound closure.

There are reports of puncture and pinch wounds inducing cell fusion. Perhaps the reviewer is suggesting that we might be able to identify a wounding method that does not induce cell fusion and then compare the rate of wound closure. However, another type of wound would probably inflict different amounts of cell damage and so would be hard to compare. Overall, we think the half-and-half system of comparing responses on the two sides of the wound is the best, most controlled comparison.

(23) Figure 4F-G. It was mentioned that there is less syncytia formation in Atg KD cells, however the difference in cell area between control, WT and Atg KD is not obvious. The authors may want to mark the dots that represent syncytia to distinguish them from mononucleated cells.

The point we are trying to make (now Fig. 4G-H) is that cell area is related to distance moved, regardless of how cell area is determined. We do not have the ability to count nuclei in the control sides (nuclei are labeled only on the pnr side), and further, we have reported separately (White et al, 2024) that there is a limited amount of endocycling in these cells, which should also increase area.

(24) Figure 5G. y axis. The authors may want to change "small cells" to "mononucleate cells".

We changed it to "unfused cells" which is the term we used in the legend. In these wounds we were unable to visualize nuclei.

(25) Page 12. Line 243. (Figure 5D,G) instead (Figure 5D).

We changed it to read (Figure 5D,G).

(26) Page 12-13. Lines 241-246. The description of "mononuclear cells removed" and "syncytia outcompete unfused cells" may be clearer if explained here as mononuclear cells joining the syncytium by cell-cell fusion.

Here we are describing a different phenomenon - not that fusion is removing all the smaller cells but rather that the syncytia are faster/better/more effective at wound closure than the smaller cells. This is illustrated in Fig. 5Cii-Ciii.

(27) Page 13. Line 261. "Thus, there were about five-fold more tangential borders lost to fusion than radial" Is this conclusion also true when analyzing each wound individually?

This is a reproducible finding, that there is more fusion across tangential borders than across radial borders. The ratio of tangential-border loss: radial-border loss for each wound is as follows:

wound 1, 63:12

wound 2, 39: 11

wound 3, 44:8

wound 4, 50:8

(28) Page 31. Figure 6 - figure supplement 1 legend, Line 652. Make "B)" bold.

Done.

(29) Page 14. Lines 270-275. If there is an advantage to radial fusion versus cell intercalation for wound closure speed, how do the authors explain that the percentage of radial fusion is lower than the percentage of intercalation? (Figure 6D) How does the wound affect the molecular level (fusogen expression?) of the surrounding cells?

We expect that radial fusion specifically reduces the need for intercalation at the leading edge, as shown in Fig. 6C. Both would speed closure, however, as any increase in cell area will allow more efficient redistribution of resources such as actin and will also reduce the total number of junctions needing to be remodeled as the wound closes. Since we don't know the fusogen, we can't say how the wound affects its distribution.

(30) Page 14. It is not clear where the experimental data ends and the model starts. For example, in line 276, it would be clearer to describe the "tissue fluidity as measured" or is it more precise to write instead "as estimated/calculated". The fusion between observations and model is confusing and maybe this should be unfused.

This text, referring to the analysis in Fig. 7A, B, is not a computational model but rather a quantitative analysis of tissue fluidity as measured by a pre-existing metric, the shape index. This is experimental data. The computational model begins in the next paragraph, accompanying Fig. 7C, D. We edited the language slightly in this paragraph to clarify.

(31) Figure 8 versus Figure 2C. Actin-bound GFP versus cytoplasmic GFP? Both mentioned as Actin GFP. Make it clear.

They are indeed the same thing, actin protein fused to GFP, as described in the text and legend, and they are labeled identically.

(32) Figure 8Biii, Div. Nice presentation of signal distribution between the cells.

Thank you!

(33) Page 30. Figure 8G legend. Lines 623-628. Is the shown mean profile plot based on specific images shown in Fi and Fv or just the cells represented there? Since Fi is a single z slice and Fv is maximum intensity projection which are not comparable.

In response to the reviewer's question, we reanalyzed the image. Fig. 8G compares Z-projections.

(34) Page 15. Line 303-304. "Tangential border fusions allow resources from distant cells to be mobilized to the wound edge." Does not this leading-edge actin localization happen in radial fusions close to the region of the wound?

Fusions along radial borders, as shown in the top panel of Fig. 6A, would not offer the opportunity to move actin from distant cells to cells nearer to the wound.

(35) Page 15. Did the authors test any predictions from the simulations of the model experimentally?

This isn’t so much a predictive model as an exploratory model that addresses one question: is it plausible that the presence of syncytia can speed closure by reducing the need for intercalations, even if the syncytia have no other special properties. The only prediction would be that inhibiting fusion would slow wound closure.

(36) Page 17. It would be interesting to discuss the following questions: (A) Is autophagy required for fusion. (B) Is Atg1 required for epithelial cell fusion? (C) Is autophagy required for wound repair? Are any of the combinations correct (A&B, A&C, B&C, A&B&C)

The role of autophagy in wound-induced cell fusion was thoroughly explored in the 2022 EMBO J paper from Maria Leptin's lab, "Autophagy-mediated plasma membrane removal promotes the formation of epithelial syncytia" by Kakanj et al. We merely knockdown a gene they discovered to be important for wound-induced epithelial fusion, Atg1, as one means of investigating how syncytia contribute to wound closure. Our results don't add to their findings, and the role of autophagy is not what we want to focus on in our Discussion.

(37) Page 19. Line 381-384. "If N represents the number of cells that fused, our results suggests that syncytia can apply up to N times more actin to the leading edge; considering that we observed syncytia with dozens of nuclei, this could represent a significant enhancement of actin at the leading edge. Increased actin might explain the ability of syncytia to outcompete diploid cells at the leading edge." To enhance this suggestion, the authors may want to compare actin signal in the leading edge of different size syncytia.

We thought a lot about this experiment because reviewer 2 asked for it in the previous round of review, but as we said then, we can imagine too many caveats to the interpretation to make it worthwhile.

(38) Page 22. Line 447-448. "...fusion would act the fastest after wounding because there is no need for DNA replication." There may be a potential need for protein (fusogen) synthesis.

The timing of fusion, which we report here begins within 10 minutes after wounding, suggests that if there is a fusogen, it is already present in the cells before wounding.

(39) Page 34. Line 716. Add "C" to "29{degree sign}".

Done.

(40) Page 38. "Wound closure analysis" part. Can the wound closure be visualized using brightfield?

The scar also impedes imaging through bright-field microscopy.

Version 3

Abstract

In response to injury, cells proliferate, migrate and invade to replace missing cells and close wounds. However, the role of other wound-induced cell behaviors is not understood, including the formation of syncytia (multinucleated cells). Here, we use in vivo live imaging to analyze wound-induced syncytia in mitotically competent Drosophila pupae. We find that almost half the epithelial cells near a wound fuse to form large syncytia. When the autophagy gene Atg1 is knocked down, fewer syncytia form, and wounds close more slowly. Further, a computational model of tissue fluidity indicates that cell fusion speeds wound closure time by about one third. Syncytia use several routes to speed wound repair: they outpace diploid cells at the wound margin to lead the initial resealing of the wound; they reduce the need for intercalation as the tissue reshapes during closure; and they pool resources of their component cells to concentrate them toward the wound margin. In addition to wound healing, these properties of syncytia are likely to contribute to their roles in development and pathology.

Introduction

Injury is a constant reality of life, and survival requires all organisms to repair wounds. Wound-induced cell behaviors like proliferation, migration, and invasion replace missing cells and close wounds (Martin et al., 2024). Other cell behaviors are induced around wounds, but their contribution to wound healing is not well understood, e.g., the fusion of cells into syncytia. Syncytia are one type of polyploid cell, and it is generally appreciated that increases in ploidy – number of genomes per cell – is a common response of post-mitotic tissues and cells to injury (Losick et al., 2013; Nandakumar et al., 2020; Tamori and Deng, 2013). Wound-induced epithelial syncytia were first observed around epidermal puncture wounds in Drosophila larvae and adults (Galko and Krasnow, 2004; Losick et al., 2013), consistent with the idea of polyploidy induction in non-proliferative cells. However, recent studies have observed syncytia around wounds in mitotically competent tissues: around laser-ablation wounds in Drosophila pupal epidermis (Wang et al., 2015) and in zebrafish epicardium damaged by endotoxin, microdissection, or laser ablation (Cao et al., 2017). Further, injury associated with the surgical implantation of biomaterials can cause immune cells to fuse into multinucleated giant cells, which are associated with rejection (Al-Maawi et al., 2017). Similarly, injury induces bone marrow-derived cells to fuse with various somatic cells to promote repair (Alvarez-Dolado et al., 2003; Corbel et al., 2003; Davies et al., 2009; Nygren et al., 2004), and injury induces fusions of some C. elegans neurons (Ghosh-Roy et al., 2010; Oren-Suissa et al., 2017). The many instances of syncytia being induced by wounds raise the possibility that syncytia offer an adaptive benefit. It is not clear, however, what that benefit is.

Syncytia can form either by endomitosis – mitosis without cytokinesis – or by cell-cell fusion. Cell fusion is widespread during animal development in both vertebrates and invertebrates. For example, myoblasts fuse to form muscle in both vertebrates and in Drosophila (Kim et al., 2015a; Lehka and Rędowicz, 2020), epithelial cells fuse during the development of several C. elegans tissues including the hypodermis (Podbilewicz and White, 1994) and vulva (Sharma-Kishore et al., 1999), and fusion gives rise to vertebrate syncytiotrophoblasts (Renaud and Jeyarajah, 2022) and multinucleated osteoclast (Søe, 2020). Cell fusions are also observed in disease: pathogen-induced epithelial fusion allows spreading of many viruses including human respiratory syncytial virus and SARS-CoV-2 (Leroy et al., 2020); and the fusion of cancer cells with bone-marrow derived cells is implicated in metastasis (Pawelek and Chakraborty, 2008).

Here, we use live imaging and clonal analysis to understand the behavior of syncytia following laser wounding in the Drosophila pupal notum. The unwounded notum is a monolayer epithelium composed of mononuclear diploid cells that are mitotically competent. Nonetheless, during the first several hours after wounding, many of the surrounding cells fuse to form giant syncytia. We previously reported that after wounding the pupal notum, some wound-adjacent cells undergo a single round of endoreplication, increasing their ploidy two-fold (White et al., 2024), but fusion increases ploidy far more, with some wound-induced syncytia observed to have more than a dozen nuclei. Some cell fusions are obvious with apical borders breaking down between cells, while others surprisingly appear as shrinking of a cell’s apical surface. Combined, fusion is a common fate of cells near wounds: about half the cells fuse to form syncytia within 70 µm of a wound with 30 µm radius. To understand the role of fusion-mediated syncytia in wound repair, we undertook two types of analysis, first analyzing an autophagy mutant that reduces fusion after wounding, and second comparing the behaviors of syncytial and mononuclear cells within the same wound. Compared to their smaller mononuclear neighbors, syncytia outpace mononuclear cells to the leading edge. We propose that syncytia accelerate wound repair because cell fusion limits the need to negotiate cell intercalations as the wound closes, a hypothesis supported by computational modeling; and because clonal analysis demonstrates that syncytia mobilize and transport cell resources such as actin from distal cells to the wound margin.

Results

A mitotic tissue undergoes cell-cell fusions after wounding

The pupal notum is a monolayer columnar epithelium that will become the epidermis of the adult fly. At the time of our analysis, 12-18h after puparium formation (APF), it sits atop a basement membrane that is still developing (Mehaffey et al., 2024), and it is composed of diploid cells that undergo regular mitotic cycles (Guirao et al., 2015) (Figure 1—figure supplement 1A-C). To analyze cell behaviors around wounds, we live-imaged after laser ablation. Epithelial cell borders were labeled by the adherens junction protein p120ctn-RFP (Ogura et al., 2018) and nuclei were labeled with GFP-labeled Histone2Av (Histone-GFP). Two hours after wounding, we observed what appeared to be prominent syncytial (multi-nucleated) cells around the wound (Figure 1A, Video 1); experiments described below demonstrated that these were indeed syncytia. Some appeared to contain over a dozen nuclei within epithelial borders. For both syncytial and mononuclear cells, it was difficult to assign nuclei precisely to cell borders because notum epithelial cells are not oriented at right angles with respect to the surface, and in 2-D projections, a nucleus was frequently observed outside the cell’s apical border (Figure 1—figure supplement 1D). Accordingly, using apical area and nuclear density, we estimated the number of nuclei within the 3 largest syncytia in different wounds. The number of nuclei in these syncytia increased over time: at 1 h after wounding the three largest syncytia contained an average 3-13 nuclei, and 2 h after wounding they nearly doubled to 6-20 nuclei (Figure 1B). Interestingly, larger wounds generated syncytia with larger apical areas and more nuclei, proportional to wound size, suggesting syncytium formation is a dynamic and scalable response to injury (Figure 1 C-E).

Wounds induce epithelial syncytia via cell fusion.

A) Syncytia form within 2 h post wounding, evident by the clustering of multiple nuclei within cell borders. B) The number of nuclei per syncytia increases over time after wounding. Number of nuclei was estimated based on area and nuclear density (see text) for the 3 largest syncytia of 3 different wounds, mean and SD. C-D) Larger wounds generate larger syncytia. Images are at 3 h post wounding, single syncytium outlined. E) Syncytial apical area is proportional to initial wound size. Each dot represents a wound, made with either low or high laser energy as sown. For each wound, the mean area of the three largest syncytia is plotted. Bars represent SD. F) A time course of six cells fusing within 30 min after wounding. Apical borders are lost (white arrowhead) as syncytia form. Original cells are numbered. G) All borders lost to cell fusion (white) mapped to cells in the first frame after wounding. The leading edge of wound closure will form at dashed line; cells within the shaded area were damaged by the wound and will be dismantled. H) Distance from the wound center vs time for all border breakdowns in four wounds. Each symbol represents a cell border. Leading-edge locations indicated by solid lines. I) Cytoplasmic GFP is expressed in cell 1 before wounding and mixes with neighboring cells 2-4 by 2 min after wounding. Cytoplasmic sharing is followed by the lagging fusion indicator of visible border breakdown (white arrow). The fates of cells 3 and 4 are shown at later times in Figure 2D. Maximum intensity projections in A, C-D, G; single Z slices in F, I. Scale bars: A,C,D,G = 20 µm, F,I = 10 µm. W and red star indicate wound.

Live imaging of cells after wounding revealed the gradual loss of p120ctn along the border between two epithelial cells followed by rounding of the resulting syncytium (Figure 1F,I, and Video 2), such that the initial two-cell morphology gradually became a one-cell morphology. We termed the visible loss of p120ctn “border breakdown”. We determined that as p120ctn was lost, the adherens junction protein E-cadherin was also lost (Figure 1—figure supplement 1E-H) indicating adherens junctions disassembly along the border between the cells. Border breakdowns were found spatially clustered in the first three to four rows of cells, 30-50 μm from the wound center, and they occurred primarily within the first hour after wounding, some within 10 minutes after wounding (Figure 1G,H). These events could be explained by wound-induced cell fusion or by wound-induced epithelial-to-mesenchymal transition. To ask whether border breakdowns represented cell fusions, we analyzed cytoplasmic mixing. Individual GFP-labeled cells (clones) were generated at random locations using the flip-out Gal4 technique (Pignoni and Zipursky, 1997). Before wounding, the level of cytoplasmic GFP fluorescence was stable yet exhibited cell-to-cell variability, allowing some differentiation of cells by intensity (Figure 1Ii). Minutes after a nearby laser ablation, GFP was observed to diffuse from labeled cells into neighboring unlabeled cells. GFP mixing between two cells was followed by the eventual loss of their shared p120ctn-labeled cell border, confirming that border breakdown is indeed cell fusion, but that cytoplasmic mixing occurs more than 10 minutes before the border breakdown is first observed to start (Figure 1I, Video 2). GFP mixing preceded border breakdowns in all observed cases (n=11). Cytoplasmic GFP mixing resulted in a consistent level of fluorescence across the two cells, indicating that GFP mixing is a reliable indicator of plasma membrane fusion. The order of events demonstrates that plasma membrane fusion occurs many minutes before the visible disassembly of the apically located adherens junctions. Thus, epithelial fusion is a rapid local response to wounding, and border breakdown is a lagging indicator of cell fusion.

Apical cell shrinking also represents cell-cell fusion, evident later during wound closure

Border breakdowns, representing cell fusion events, occurred mostly within 1 h after wounding, so it was unclear how syncytia grew in size and nuclear number between 1-2 h after wounding. However, an unexpected cell behavior was frequently observed during this time: the apical area of diploid epithelial cells shrank until they disappeared from the epithelial sheet, a phenomenon we termed “apical shrinking” (Figure 2A). Although the appearance of apical shrinking suggested cell extrusion, we were surprised at its high frequency around wounds (Figure 2E), as wounds have already lost many cells and would be expected to prioritize cell survival to speed barrier repair. To understand the fate of apical-shrinking cells, we tracked their nuclei and determined that they did not extrude; rather, all apical-shrinking cell nuclei moved laterally within the plane of the epithelium to join neighboring syncytia (n=7, Figure 2B), suggesting that apical shrinking is an indicator of cell fusion. To better understand this behavior, we analyzed individual GFP-labeled cells so that we could observe the fate of the cytoplasm. Like with border breakdowns, GFP mixing in the X-Y plane of the epithelium was observed to precede the initiation of apical shrinking, although the interval between GFP mixing and apical shrinking was longer, one or more hours (Figure 2C, Video 3). In X-Z projections, GFP labeled cytoplasm associated with apical shrinking shifted laterally to join with a neighboring cell as shown for two apical-shrinking cells in Figure 2D. For the cell labeled with the yellow arrow, before apical shrinking at 60 min post wound, its GFP-labeled cytoplasm extended from the plane of the adherens junctions (ZAJ) basally over 8 µm; when apical shrinking was underway at 120 min post wound, its cytoplasm extended less than 5 µm down from the apical plane; and when shrinking was almost complete at 150 min post wound, it extended only 1-2 µm down from the apical plane, appearing as a whisp of cytoplasm connecting the shrinking apical surface with the neighboring syncytia. A similar loss of basal depth over time is observed in the cell labeled with the white arrow. In both these cells, as the cell lost its basal volume, cytoplasm moved laterally to join the neighboring syncytia. Thus, analysis of both nuclei and cytoplasm demonstrated that apical shrinking represented cell fusion events. The spatial distribution of apical shrinking around wounds was similar to border breakdowns, but apical shrinking events occurred later and were more numerous (Figure 2E,F). Thus, cell fusion is evident as either border breakdown or apical shrinking, which both occur after GFP sharing.

Cell fusion often appears as apical shrinking.

A) The apical footprint of an epithelial cell shrinks in the epithelial plane after wounding (yellow arrow). B) Same sample as A, showing that the nucleus (asterisk) of the apically shrinking cell enters a neighboring syncytium, outlined in Biv. C) Cell 1 expresses actin-GFP before wounding. After wounding, cell 1 fuses with cells 2-6 evidenced by actin-GFP sharing, then the apical surface of cell 1 shrinks. D) Two cells undergoing apical shrinking are identified in the X-Y plane with white and yellow arrows, shown at three times after wounding. Below, two X-Z projections are shown for each, at Y1 and Y2 as indicated, with the Z-plane of the adherens junctions (ZAJ) indicated by dotted lines. Rather than extrude, cytoplasm moves to the right as the cell depth in the Z-axis diminishes, so that cytoplasm joins with the neighboring wound-proximal syncytium. These video frames are a continuation of the sample shown in Figure 1I. E) All cells displaying apical shrinking mapped to first frame after wounding. F) All apical shrinking and border breakdowns were tracked in 3 wounds. Border-breakdown fusions happen sooner after wounding than apical-shrinking fusions. A, B, C, D (top), and E show maximum intensity projections. Lower panels in D show X-Z projections. Scale bar for all panels A and B shown in Biv. A, B, C, D (top),E = 10 µm; D (X-Z projections) = 5 µm.

To determine what percentage of cells around a wound will fuse, we analyzed over 100 single epithelial cells randomly-labeled with cytoplasmic GFP within the radius of observed fusion (80 µm), tracking them for 6.5 h to assess their fate (Figure 3A,B): a full quarter of the cells fused (25%), sharing GFP before borders broke down or apical shrinking; 67% persisted as diploid cells, most with stable GFP, but infrequently (n=3) with GFP mixing and no subsequent cell fusion, suggesting that fusion was initiated then aborted; the remaining 7% could not be tracked (Figure 3—figure supplement 1A). All GFP-labeled cells with apical shrinking had previously shared cytoplasm, indicating that all apical shrinking represents fusion. For the three diploid cells with previous cytoplasmic mixing, we interpret those as cells that initiated fusion but were unable to productively stabilize and expand the site of fusion and instead returned to the diploid state. Fusing cells were strongly skewed toward the center of the wound: within 70 µm, about half the cells (47%) underwent fusion (Figure 3A, left), and fusion continued for over 300 min after wounding (Figure 3—figure supplement 1B,C). Apical shrinking accounts for more than half of all fusing cells (Figure 3C), and the spatial distribution of fusing cells that shrank vs. lost borders was similar (compare Figures 1G and 2E). However, apical shrinking began later than border breakdowns (Figure 2F, Figure 3—figure supplement 1C), continued for several hours after wounding (Figure 2F, Figure 3—figure supplement 1C), and took longer to complete (Figure 3—figure supplement 1D). Thus, fusion associated with apical shrinking accounted for continuing syncytial growth after border breakdowns subsided. Further, cell fusion is a persistent behavior over the course of wound closure.

Half the cells near the wound fuse to form syncytia, demonstrated by tracking individual cell fates.

A) All GFP labeled cells in the region of fusion (80 µm) were tracked in 5 wounds over 6.5 h after wounding to determine frequency of fusion initiation (GFP mixing) and later morphological fusion, i.e. a change in the adherens junction marker p120ctn-RFP, which outlined two cells before fusion and one large cell after fusion. All morphological fusion events were preceded by GFP sharing, and 90% of GFP sharing events were followed by morphological fusion. Untrackable cells lost GFP, see Figure 3—figure supplement 1A. B-C) Pre-wound location of persisting and fusing cells from panel A is shown with respect to the wound center. Panel B shows that fusion is common within 70 µm. Panel C shows that apical-shrinking fusion and border-breakdown fusion occur at similar distances from the wound.

Our individual cell tracking indicated that syncytia were the result of epithelial cell fusions. However, phagocytic hemocytes are known to migrate to the sites of pupal wounds (Antunes et al., 2013). To determine if hemocytes fused with epithelial syncytia around wounds, we labeled hemocytes with cytoplasmic GFP (hml>GFP) and observed their behavior in live imaging. In six wounds, we observed GFP-labeled hemocytes at the basal side of the wound margin; however, we did not observe any events where GFP was transferred from hemocytes to the epithelium, ruling out the fusion of hemocytes with epithelia in pupal wounds (Video 4) .

Syncytia close wounds faster

To evaluate the role of cell fusion and syncytia in repairing wounds, we wanted to analyze a mutant that blocked syncytia formation. It has been reported that several individual autophagy (Atg) genes are required for syncytia formation around small laser ablation wounds in the polyploid Drosophila larval epithelium (Kakanj et al., 2022), so we asked if the autophagy gene Atg1 promotes fusion around laser wounds in the diploid pupal epithelium. Using a dsRNA line known to disrupt syncytia in larvae, we knocked down Atg1 on one side of the wound expressing pnr-Gal4, as illustrated in Figure 4A, reserving the other side of the wound as an internal control, modifying an internally controlled wounding system we have previously reported (O’Connor et al., 2021b). Visualizing Ecad-labeled cell borders across the wound, we found that indeed Atg1 is required for syncytia formation in pupal wounds, like it is in larval wounds (Figure 4B). Specifically, we observed a reduction in cell fusion, both reductions of border loss and apical shrinking (Figure 4C,D). We next compared wound closure rates on the two sides of the wound (Figure 4E,F). Because leading edge migration is very slightly delayed on the pnr-Gal4 side of the wound even in control wounds (pnr>+, (Hua et al., 2026), we compared the difference in wound radii on the two sides, comparing control wounds to pnr>Atg1RNAi wounds. By 5h after wounding, the pnr>+ side lagged behind the other side by only about 10 µm, yet the pnr>Atg1RNAiside lagged behind the other side by an average of about 25 µm, a reproducible, significant, and large reduction in wound closure when syncytia formation was reduced (Figure 4F). These results suggest that syncytia speed wound closure. It is difficult, however, to disentangle the roles of autophagy and syncytia in this experiment, as the Atg1 knockdown cells may move more slowly because they are defective in the essential cellular function of autophagy and/or lacking syncytia. To address this question, we found a strong correlation between the size of the cell/syncytium and the distance it had moved from the start, measured at the time when the wound was half-closed. This correlation was true for cells/syncytia in control wounds as well as for those when Atg1 was knocked down (Figure 4G,H respectively). These data indicate that syncytia formed through cell fusion speed wound closure.

Atg1 promotes wound-induced fusion and faster wound closure.

A) Overview: Atg1 was knocked down in central pnr domain (pink), visualized by nls-mCherry (not shown). A laser pulse was targeted to the edge of the pnr domain (right or left) so that the wound extended into both Atg1 knockdown and control domains, allowing the symmetry of response to be analyzed. B) Inhibition of Atg1 decreased the frequency of wound-induced fusion in the pnr domain (right) where Atg1 is knocked down compared to the control domain (left). Some cell fusions are indicated by yellow arrows in the 30 min panel. The red dotted line at the center of each frame indicates the border between the domains, determined by the expression of UAS-nls-mCherry (not shown) under control of pnr-Gal4 on the right. C) Atg1 is required for syncytia to enlarge over time. In the pnr domain of control animals, the number of nuclei per syncytia increased over time after wounding. When Atg1 was knocked down, syncytia remained small over time. The number of nuclei was estimated based on area and nuclear density for the 3 largest syncytia of 3 different wounds, mean and SD. Compare to Fig. 1B. D) Inhibition of Atg1 decreased the number of wound-induced fusions in the pnr domain compared to the control domain in the same animal. For border loss, n = 11 pupae, p < 0.0001. For apical shrinking, n=9 pupae, p = 0.0008 (paired T-test). E-F) Wounds closed more slowly when Atg1 was knocked down. (E) Radius of the wound was calculated by the distance between the wound center and the leading edge in pnr or control domains. (F) Difference in the wound radius between pnr and control domains over time (i.e., wound asymmetry) is shown. Each dot represents one wound. With no genetic manipulation in either domain (black), wound radius was slightly larger in pnr than in control domain. When Atg1 was knocked down in the pnr domain, the asymmetry grew over time indicating slower wound closure on the Atg1 knockdown side. p = 0.0031 (5 hr) calculated from two-way ANOVA, fit full model, comparing each cell mean with the other cell mean in that row. Bar = 1 +/- SEM. G-H) Larger cells (i.e. syncytia) move further during wound closure, evident in control wounds (G) or with Atg1 knockdown (H). Each data point represents a cell or syncytium present at the leading edge when the wound is half closed. The curve of square root area was overlaid for comparison, as it represents the amount of movement expected based on reshaping of a cell/syncytium from an initial round state to one strongly elongated towards the wound.

Syncytia outcompete mononucleate cells at the leading edge of repair

To understand this phenomenon better, we further investigated control wounds. Because about half the cells fused to form syncytia around wounds, we were able to compare the behavior of syncytial to non-syncytial cells within the same wound, providing a well-controlled environment for assessing how syncytia contribute to wound closure. In live imaging, we observed that syncytia frequently moved faster than unfused cells toward the wound, overtaking unfused cells as the wound closed. Figure 5A tracks four GFP-labeled cells near a wound, with cell 1 furthest from and cell 4 closest to the wound (Figure 5Ai). Middle cells 2 and 3 fuse with five others into a large syncytium, which advances toward the wound faster than both non-fusing cells, and by 400 min after wounding the syncytium extends well beyond the unfused cells (Figure 5Aiii). This behavior was evident even without GFP labeling: Figure 5B-C show a group of cells at the leading edge of the wound (the wound is recognized by the lack of p120ctn). At 90 min after wounding, this region of the leading edge comprises three mononuclear cells flanked on either side by syncytia (outlined in yellow); the middle of the three mononuclear cells is outlined in orange (5B) and white (5C). By 150 min, the syncytia have pushed out all three mononuclear cells from the leading edge, and at 360 min two of these mononuclear cells are visible about 20 µm back from the wound edge. Indeed, we observed that several hours after wounding, the entire leading edge was occupied by syncytia. To investigate how syncytia came to occupy this position at the front lines of wound healing, we expressed MyoII-GFP/Zip-GFP, which along with actin forms the contractile purse string and makes the leading edge visible, along with p120ctn-RFP to label cell borders. We analyzed the persistence of all unfused and syncytial cells at the leading edge over the course of closure for three wounds, starting when the leading edge was first visible about 30 min after wounding (Figure 5D-G). At the start, 75% of the perimeter was occupied by syncytia with the rest occupied by 10-13 unfused cells (Figure 5D,G). As the wound closed, the syncytia became larger and displaced all the unfused cells, with the last unfused cell removed from the leading edge 20-160 min before closure (Figure 5E-G). No unfused cell persisted at the leading edge through wound closure, indicating that syncytia outcompete unfused cells to close the wound.

Syncytia outpace mononuclear cells.

A) Before wounding, two clusters of cells are labeled with GFP, 1-2 and 3-4 (white numbers). After wounding, cells 2 and 3 fuse with neighbors 5-9 to form a syncytium (dashed leading edge in Aiii), which advances toward the wound, passing unfused cell 4 (arrow). B-C) Unfused cell (outlined in orange and white) is replaced at leading edge by neighboring syncytia (S and yellow outline). D) Leading-edge perimeter was analyzed over time for three wounds, also analyzed in panel G. Percent leading edge occupied by syncytia increased over time to 100%. E) Images of sample 1 from graph D. Unfused cells (arrows) were tracked over the course of wound closure; 30 min after wounding, 10 unfused cells at the wound leading edge are indicated by arrows. F) At 120 min, the last unfused cell was ejected from leading edge (arrow). The wound closed at 140 min. G) The loss of unfused cells was analyzed over time in the three wounds from panel D. All unfused cells are excluded from leading edge by syncytia well before each wound closed 20-160 min later. Images are single Z slices in Ai-Aii and maximum intensity projections in Aiii-F. Scale bar for B,C shown in Civ and for E,F shown in F. Scale bars in A,E,F = 20 µm, B,C = 10 µm.

Radial border fusion reduces intercalation, the rate-limiting step of wound closure

Why do syncytia move faster? To address this question, we considered the geometry of fusion because different fusion orientations could provide different benefits for wound repair. We observed that fusions sometimes occur between two adjacent cells equidistant from the wound, as diagramed in the top panel of Figure 6A and exemplified in Figure 1Iiii-Iiv (cells 1,2). This breakdown of a radial border produces a syncytium elongated along the wound edge. Alternatively, fusion may occur between adjacent cells at different distances from the wound, as diagramed in the lower panel of Figure 6A and exemplified in Figure 1F (cells 3,4). This breakdown of a tangential border produces a spoke-like syncytium pointing into the wound. To analyze the frequency and timing of these two different axes of fusion, we calculated the orientations of all lost borders in four wounds. (The angle of the lost border can be determined for border breakdowns; it is difficult to determine for apical shrinking because the other fusion partner is hard to identify). In four analyzed wounds, we identified 235 border breakdowns: 39 radial borders and 196 tangential borders (Figure 6B). Thus, there were about five-fold more tangential borders lost to fusion than radial. Nonetheless, the two types had similar temporal and spatial distributions around the wounds (Figure 6—figure supplement 1).

Cells fuse along different axes at different frequencies, reducing the number of wound-proximal cell intercalations.

A) Illustration of tangential vs radial border breakdown. B) More tangential than radial borders break down after wounding. Data are from four wounds, with a total of 235 border breakdowns: 39 radial and 196 tangential. Individual counts are shown per wound with mean and SEM. C) Fusions across radial borders reduce intercalation at a wound. D) Quantification of intercalations and fusions around the three wounds of Figure 5D,G.

As a wound closes, fewer cells can occupy the ever-smaller leading edge, necessitating a rearrangement of cell adhesions to allow cell intercalation, as depicted in the top of Figure 6C. Indeed, when cells do not fuse after wounding, as in Drosophila wing discs, the impact of intercalations on tissue fluidity is known to be a rate-limiting step of wound repair (Tetley et al., 2019). As shown in the bottom of Figure 6C, the requirement for cell intercalation would be reduced specifically by radial border fusion.

To consider the contribution of radial border fusions to wound closure, we took three approaches. First, we analyzed existing data from the three wounds shown in Figure 5D-G, and asked how many leading-edge cells fused radially vs. intercalated: 17-41% of cells removed from the leading edge were removed through fusion, reducing the burden of intercalation substantially (Figure 6D). Interestingly, the larger the percentage of cells that fused rather than intercalated, the faster the wound closed (Figure 6D).

Second, we estimated tissue fluidity as measured by the cell shape index, ρ = P⁄√A, where P is cell perimeter and A is cell area. This dimensionless index characterizes the cell jamming transition in epithelia. In idealized vertex models, this transition occurs at ρ = 3.81: for lower values, the epithelium is in a jammed state with energy barriers that limit intercalations, giving rise to a solid-like tissue; for higher values, the energy barriers vanish, yielding an unjammed fluid-like tissue (Bi et al., 2016; Park et al., 2015). In real epithelia, ρ is increased by the fact that cell edges are not necessarily straight (as they are by construction in vertex models), and thus the solid-fluid transition may take place at different values of ρ (Wang et al., 2020), but higher ρ still implies greater tissue fluidity. To measure cell shape index, we segmented images of the pupal notum before and after wounding, extracted perimeter and area data for each cell, and calculated the distributions of cell shape index for both syncytia and non-syncytial cells (Figure 7A-B). The mean cell shape index increased significantly after wounding with syncytia having a higher shape index than nearby post-wound non-syncytial cells (p = 0.001) and both having higher shape indices than pre-wound cells (p < 10-6). The post-wound epithelium is thus substantially more fluid-like than it was before wounding

Computational modeling indicates that fusion speeds the rate of wound closure because of reduced cell intercalation.

(A-B) Experimental observations of cell shape index (perimeter divided by square root of area). By two hours after wounding, the cell shape index has increased for both syncytia and non-syncytial cells within two rows of the wound margin. These increases indicate an increase in tissue fluidity. (A) Cells and syncytia were segmented, and cell shape index calculated, as shown for this image at 115 min after wounding; cell shape index indicated by color-coding. (B) Pre-and post-wound cell-shape distributions were compiled from 6 pupae. The red dashed line represents the critical cell shape parameter, 3.81, for the solid-to-fluid transition in vertex models of epithelia. Grey boxes span the interquartile range with a line at the median; whiskers extend out 1.5x the median-to-closest-quartile range; and outliers beyond the whiskers are plotted as individual points. Distributions include 1304 pre-wound mononucleate cells, 168 post-wound mononucleate cells, and 52 post-wound syncytia; **** denotes p < 10-6 level; *** denotes p < 0.001. (C-E) Vertex model simulations of wound closure with cell fusion allowed or suppressed. When the spatial and temporal probabilities for cell fusion match experimental observations, the model yielded 55 ± 22 fusions per wound (mean ± standard deviation), with 15 ± 6 of these occurring at radially aligned edges. As shown in C, fusions in the model speed wound closure: four simulations per condition; shaded regions denote standard deviation; time normalized to the average time to closure when fusion is suppressed. Panels D-E show a series of still frames from a matched pair of simulations in which fusion is allowed or suppressed. Corresponding videos are available as Video 5.

Third, we adapted the computational vertex model from Tetley et al. (2019) to include probabilistic cell-cell fusions to directly investigate the role of cell fusion in tissue fluidity and wound closure. We generated four initial cell configurations with similarly sized wounds and used these to run four matched pairs of simulations – one that proceeded to close without any cell fusions and the other that allowed fusions to occur with spatiotemporal probabilities that matched our experimental observations (Figure 6B, Figure 6—figure supplement 1A). On average, the probabilistic fusion-allowed simulations had 68 ± 14 (mean ± SD) fusions and closed the wound in about 1/3 less time than those without fusions (Figure 7C). Selected frames from one of the matched simulation pairs are shown in Figure 7D-E and full videos of this pair are available as Video 5. Importantly, the modeled syncytia had no special properties: same contractility as other cells, same target shape index, and same target area as the sum of the fused pair. Just the presence of these larger cells reduced the need for intercalations during wound closure and thus sped up the process.

Tangential border fusions allow resources from distant cells to be mobilized to the wound edge

In contrast to radial border fusions, tangential border fusions might provide a way for cellular resources that would be trapped in distal cells to move toward the wound to contribute to closure. To test this hypothesis, we generated small flip-out clones expressing actin-GFP and wounded such that unlabeled cells intervened between the labeled cell and the leading edge (Figure 8). Like cytoplasmic GFP in Figures 1I and 2C, we observed actin-GFP to equilibrate between cells soon after fusing. For syncytia that did not have access to the leading edge (n =2), actin-GFP levels remained uniform (Figure 8A,B). In contrast, syncytia with access to the leading edge (n = 10) first uniformly distributed actin (Figure 8Cii), but once the leading edge was contacted, they redistributed actin-GFP to it (Figure 8C,D). Kymographs of actin-GFP confirm that regardless of location, actin equilibrates between fusing cells within 5 minutes (Figure 8Biii, Div); however, nearly all actin that originated in the distal cell is redistributed to the leading edge in a syncytium positioned there (Figure 8Div). In one striking instance, actin-GFP appeared to travel through three cells to arrive at the leading edge from its initial location three rows back (Figure 8E-G, Video 6). Importantly, although actin is labeled from only one of the fusing cells, it likely represents the location of actin from all fusing cells, explaining why syncytia are better able to occupy the leading edge. We envision that other resources would also be concentrated in subcellular regions as needed by syncytia. This ability of syncytia to concentrate pooled actin to the leading edge may allow syncytia to outcompete their smaller mononuclear neighbors and close wounds more rapidly.

Syncytia concentrate pooled resources at the leading edge.

Random scattered cells expressing actin-GFP were generated by heat-shock mediated flip-out expression of Gal4. A-B) Labeled actin is expressed in cell 1 before wounding (Ai,Bi). These cells are not close to the wound (inset in Aii). By 28 min after wounding, actin-GFP equilibrated between cells 1 and 2 (Aii,Bii), demonstrating cytoplasmic fusion. The resulting syncytium had no access to the leading edge, and actin remained equilibrated, as shown in the kymograph (Biii) generated from actin-GFP intensity over time at the yellow line in Bii. Fusion was confirmed at 180 min after wounding by apical shrinking of cell 2 (not shown). C-D) Labeled actin is expressed in cell 1 before wounding (Ci,Di) and equilibrates between cells 1 and 2 by 6 minutes after wounding, demonstrating cytoplasmic fusion (Cii,Dii). The resulting syncytium contacts the leading edge, and by 28 min after wounding actin-GFP from cell 1 is redistributed to the wound margin (Ciii,Diii), as shown in the kymograph (Div) of actin intensity over time at the yellow line in Diii. Fusion was confirmed at 55 min after wounding, by apical shrinking of cell 1 (not shown). E-F) Before wounding, actin-GFP in cell 1 is three cells away from the future leading edge. After wounding, fusion of cells 1-4 allows actin-GFP to be subcellularly localized to the leading-edge actin cable. Border breakdown is visible between cells 1-2 at 30 min after wounding (Ev) and apical shrinking occurs later (see Video 6). G) Mean profile plot of actin-GFP comparing cell 1 before wounding (from panel Fi, dotted line) to the syncytia of cells 1-4 at 30 min after wounding (from panel Fv, dark line), demonstrating that nearly all actin-GFP has been relocalized to the leading edge from its starting position 20-30 µm away. Single Z slices shown for Ei - ii, Fi - ii; maximum intensity projections for A-D, Eiii - v, Fiii - v. Scale bar for panels in A-B shown in Bii, 10 µm. Scale bar for panels in C-D shown in Diii, 5 µm. Scale bar for panels E-F shown in Fv, 10 µm.

Discussion

Previous work established that polyploid cells – both cells with multiple nuclei and cells with enlarged nuclei – are important for closing epithelial wounds in post-mitotic tissues (Losick et al., 2013). Here we have used live-imaging to analyze the response to wounds of a diploid mitotic tissue, the epithelial monolayer of the pupal notum. We found that syncytia form around wounds by cell fusion at a remarkably high rate, with almost half the epithelial cells within 5 cells of the wound fusing with neighbors over the course of hours. Syncytial size increases with wound size, indicating syncytia formation is a dynamic and scalable response to wounding. Several lines of evidence demonstrate that syncytia are faster at closing wounds than unfused cells. Knockdown of the autophagy gene Atg1 reduces fusion around pupal wounds as was previously reported for larval wounds; Atg1 knockdown also reduces the rate of wound closure. In both Atg1 knockdowns and in controls, there was a consistent observation that the bigger the syncytium, the faster it moved. Within control wounds, syncytia completely displace unfused cells of the same genotype at the leading edge of the wound, such that wounds are closed entirely by syncytia, even though there is a smaller number of syncytia than unfused cells. Finally, computational modeling of wound closure with and without cell fusion shows that wounds with fusion-based syncytia close faster because of the reduced need for cell intercalations.

Thus, wound-induced polyploidy is a generalized strategy for wound closure, used by both mitotic and non-mitotic tissues. Although one might have wondered if polyploidy was a wound response limited to tissues that cannot use mitosis to increase nuclear content, both endoreplication and cell fusion into syncytia occur after wounding in this mitotically active diploid tissue (Wang et al., 2015; White et al., 2024; this study). The requirement for Atg1 for fusion in both polyploid larval tissue and in diploid pupal tissue suggests that the fusion mechanisms are common, despite the altered cell cycle of the starting cells. Further, our finding that larger syncytia travel faster suggests that syncytia – and by extension perhaps cell enlargement via endoreplication –occur as adaptations to close wounds faster.

We identified two unique properties of syncytia that may improve their wound-closing ability. First, syncytia are larger than neighboring diploid cells and their presence enhances tissue fluidity. Previous studies have determined that the rate-limiting step in epithelial wound healing is the process of cell intercalation or rearrangement (Tetley et al., 2019). As the wound closes and the leading edge encircling the wound gets smaller, fewer cells can occupy the leading edge. The cells leaving the leading edge must remodel their cellular junctions, requiring time and energy. Overall, tissue solidity slows closure, so enhancing fluidity promotes faster closure. By fusing into larger syncytia, cells reduce the need for adhesion remodeling and intercalation and instead just change shape as the leading edge becomes smaller. Moreover, the resulting larger cells can have significantly more fluidity and less epithelial tension than their diploid progenitors, as recently demonstrated in syncytia formed by age-induced epidermal cell fusion (Dehn et al., 2023). As our computational modeling demonstrates, fusion can speed wound closure by ∼33% even when the resulting syncytia have no special properties besides being larger than neighboring diploid cells. Making the syncytia individually more fluid would be expected to speed closure even more.

The second unique property of syncytia is their ability to pool and concentrate the cellular resources of the many constituent cells into a subcellular location within the large syncytium. We explored this concept by monitoring the localization of actin within syncytia. Actin is important for wound healing, as epithelial wounds close through the action of two actin-dependent processes: the cinching of a multicellular actin cable at the leading edge that surrounds the wound, sometimes called a purse string (Abreu-Blanco et al., 2012; Bement et al., 1993; Martin and Lewis, 1992; Tamada et al., 2007), and the protrusion of actin-rich projections into the wound (Abreu-Blanco et al., 2012; Farooqui and Fenteany, 2005). Pupal wounds have been shown to use both processes, with the actin cable developing early and actin protrusions appearing later in wound closure (Antunes et al., 2013). By labeling actin in individual cells (clones), we were able to follow it redistribution in syncytia after fusion. As expected, at first labeled actin diffuses and equilibrates throughout the new large cell. Remarkably however, when the syncytium is in contact with the leading edge, labeled actin is concentrated at the leading edge, even if the actin originated several cells away from the leading edge. Without fusion, the labeled actin would have been trapped in a distal cell without access to the leading edge, but after fusion, the syncytium can concentrate the actin of its many component cells at the wound front. If N represents the number of cells that fused, our results suggests that syncytia can apply up to N times more actin to the leading edge; considering that we observed syncytia with dozens of nuclei, this could represent a significant enhancement of actin at the leading edge. Increased actin might explain the ability of syncytia to outcompete diploid cells at the leading edge. Presumably, other resources such as mitochondria and ribosomes could also be pooled and concentrated by syncytia at cellular locations where they promote wound healing. It is known that mitochondrial fragmentation is localized to the site of cellular injury (Horn et al., 2020), and that ribosomes are localized at the tips of severed neurons (Noma et al., 2017), and syncytia would have access to a larger pool of mitochondria and ribosomes than their component cells. By pooling the cellular resources of component cells, syncytia may also allow lethally damaged cells to survive by providing them with needed survival factors originating in cells further from the wound. Thus, the concept of resource sharing that we demonstrate with actin may have ramifications for many resources. Interestingly, it has been demonstrated that wound-induced endoreplicating cells also increase protein levels of some wound-repair resources, such non-muscle myosin, presumably possible because of the many copies present in the polyploid genome (Losick and Duhaime, 2021).

Two visibly distinct cellular behaviors attended fusion in the apical plane of the adherens junctions: border breakdowns and apical shrinking. Both processes were confirmed as fusion by expressing GFP in clones and analyzing GFP sharing at fusion. Border breakdowns occurred sooner after wounding than apical shrinking and appeared to be a faster process, as apical shrinking started later and took hours to complete. It is unclear whether these differences in appearance and timing indicate different mechanisms, as both require the autophagy gene Atg1. For border-breakdown, it is simple to envision that fusion initiates near the apical adherens junctions, which are under more tension than the basolateral membranes, and this may explain why fusion proceeds more quickly there. We recently found that although epithelial tension drops after laser wounding, it is restored within about 10 min (Han et al., 2023), consistent with the timing of apical border breakdown. Also consistent with this model, in the C. elegans hypodermis, developmental cell fusion has been observed to initiate at the apical side and travel in the apical-to-basal direction (Mohler et al., 1998). It is also possible that fusion initiates at a more basal site, and that fusion proceeds apically, with adherens junction disassembly occurring later. Regardless of the site of fusion initiation, fusion seems to be a relatively equal process when border breakdown is observed, with the resulting syncytium occupying the same footprint that the two fusion partners did, at least initially. For apical shrinking, however, one cell appears to transfer its cytoplasmic contents and nucleus into another recipient cell, with both cytoplasm and nucleus moving laterally to the location of the fusion recipient, while the apical surface of the donor cell shrinks. It is possible that apical shrinking may indicate a different location for the initial fusion pore below the plane of the adherens junctions; it is also possible that this may reflect a difference in the starting pressure of the two fusion partners, so that one subsumes the other.

It is unclear what triggers either border breakdown or apical shrinking fusion. Many developmentally programmed cell fusions are mediated by fusogens, cell surface proteins that bring opposing membranes into close contact with each other, as in the C. elegans hypodermis (Chernomordik and Kozlov, 2008; Iosilevskii and Podbilewicz, 2021; Markvoort and Marrink, 2011; Mohler et al., 2002; Podbilewicz et al., 2006; Sapir et al., 2007; Shemer et al., 2004). For other cell fusions, the fusogen is elusive and may not exist, as none has been identified yet in Drosophila myoblast fusions, which occur when a fusion competent myoblast generates actin-rich podosome-like membrane protrusions that invade a founder cell (Kim et al., 2015b; Lee and Chen, 2019; Petrany and Millay, 2019; Rushton et al., 1995; Sens et al., 2010). Wound-induced fusion may be a response to the plasma membrane damage that occurs around wounds; indeed, plasma membrane damage has been documented around both laser wounds and puncture wounds (McNeil and Steinhardt, 2003; Shannon et al., 2017). Interestingly, in the large syncytial cells of the C. elegans hypodermis, plasma membrane repair after a puncture wound requires the fusogen EFF-1 (Meng et al., 2020).

Polyploidy as a wound response has begun to get increased recognition. In adult Drosophila, epithelial puncture wounds are repaired by both endoreplication and syncytia formation (Bailey et al., 2020; Besen-McNally et al., 2021; Grendler et al., 2019; Losick, 2016; Losick et al., 2013; Losick et al., 2016). In the zebrafish epicardium, genetic ablation is repaired by a wavefront of multinucleated polyploid cells formed by endomitosis, and these lead diploid cells to encompass the heart (Cao et al., 2017). In adult mammalian cardiomyocytes, polyploidy may be an adaptive response to maintain growth after the cardiomyocytes lose their ability to complete mitosis. Mouse cornea endothelial cells endoreplicate to increase polyploidy to restore tissue ploidy following genetic ablation (Losick et al., 2016). Mammalian hepatocytes are known to become increasingly polyploid with age (Carriere, 1969; Wheatley, 2008) and in response to various types of injury and disease (Gentric et al., 2015; Madra et al., 1995; Muramatsu et al., 2000; Sigal et al., 1999; Toyoda et al., 2005; Wilkinson et al., 2019). All mechanisms that promote polyploidy – fusion, endomitosis, endoreplication – result in larger cells with the potential to localize more resources; of these, fusion would act the fastest after wounding because there is no need for DNA replication. Interestingly, in Drosophila embryos, wounds induce the surrounding cells to become larger by increasing their volume alone and not their ploidy (Scepanovic et al., 2021), suggesting that simply an increase in size is important. Although many examples of wound-induced polyploidy exist, it is still likely to be an underreported phenomenon, as endpoint analysis might miss a transient polyploid response to injury; live imaging is the surest way to identify a polyploid wounding response.

Another polyploid response to injury can occur after surgical implantation of biomaterials for the purpose of guiding regeneration, as reviewed previously (Al-Maawi et al., 2017). In some cases, only mononuclear cells of the immune system respond to the implant, and in these cases the biomaterial is integrated into the body; in other cases, the implant triggers the fusion of immune cells into multinucleated giant cells, and in these cases the material is degraded and rejected. These multinucleated giant cells seem to share properties with the syncytia of the pupal notum, as they are formed by fusion in response to an environmental trigger and they have an aggressive ability to protect the animal in response to wounding. As these studies highlight, understanding the formation, maintenance, and regulation of polyploid cells may improve our ability to successfully implant biomaterials to aid tissue regeneration.

It is often noted that wound responses are similar to cancer cell behaviors. This similarity extends to wound-induced syncytia and their counterparts, polyploid giant cancer cells, as both cell types are highly aggressive and invasive. Chemotherapeutics induce the formation of polyploid giant cancer cells (Illidge et al., 2000; Mosieniak et al., 2015; Ogden et al., 2015; Wang et al., 2013), and some studies indicate that they can form through cell-cell fusion in tumors (Melzer et al., 2018; Noubissi et al., 2015; Powell et al., 2011; Song et al., 2021; Zhang et al., 2021). Once formed polyploid giant cancer cells are hypothesized to escape further chemotherapy treatments due to increased resistance to genotoxic stress (Weihua et al., 2011). These polyploid giant cancer cells and their progeny also exhibit increased migration and invasion potential (Qu et al., 2013; Zhao et al., 2021). The parallels between the behaviors of polyploid giant cancer cells and the wound-induced syncytia of the pupal notum highlight the importance of understanding wound induced syncytia formation in a highly reproducible system, as a basic understanding of how these cells form in the Drosophila notum could inform how they become dysregulated in cancer.

Methods

Resource availability

Lead contact

Requests for fly lines, reagents, and additional questions should be directed to Dr. Andrea Page-McCaw (andrea.page-mccaw@vanderbilt.edu)

Materials availability

Fly lines generated in this study are available from the Bloomington Drosophila Stock Center or from the lead contact.

Experimental model and subject details

Drosophila melanogaster

Drosophila lines used in this study are in Table S1. All Drosophila lines were maintained on standard cornmeal-molasses media supplemented with dry yeast. All flies, except those used in heat-shock flip clonal analysis, were raised at room temperature. For clonal analysis experiments flies were raised at 18 degrees Celsius until the 3rd instar stage when they were heat shocked in a circulating water bath at 37°C for 3 minutes. They were then allowed to develop at room temperature to 15-18hr after puparium formation (APF) before wounding experiments (described below) were conducted. To knock down Atg1, Atg1RNAiwas crossed to the stock for internally controlled wounding analysis, expressing Ecad-GFP and TubP-Gal80ts ubiquitously and pnr-Gal4, UAS-nls-mCherry in the central region of the notum, similar to our previous system (O’Connor et al., 2021b). Crosses were moved to 29° C for 6 days to allow expression of the RNAi construct and dissected between 12-15 hr APF.

Method details

Pupal mounting

At room temperature, white prepupae were identified and marked within plastic food vials. 15-18 APF pupae were removed from the vial onto a piece of double-sided tape (Scotch brand, catalogue #665) applied to a microscope slide. Using fine forceps, the anterior pupal case was removed exposing the head and notum of all pupae applied to the tape, as previously described (O’Connor et al., 2022; Shannon et al., 2017). The tape was carefully removed from the microscope slide and inverted onto a pre-prepared cover glass (Corning 2980-246, 24 mm x 64 mm) (O’Connor et al., 2022). The pupae were carefully pressed down onto the cover glass by adhering the section of tape above the pupal head. Once the notum was visibly pressed onto the cover glass an oxygen permeable membrane (YSI, standard membrane kit, cat#1329882) was applied to prevent the pupae from drying out during imaging.

Pupal survival

Following imaging, pupae were kept mounted as described above and allowed to continue to develop and eclose for 3-4 days. Pupae that continued developing until they were able to crawl out of the partially dissected case were classified as ‘survived’ and their data acquired form these samples were used for analysis. If a pupae did not survive to eclosion, the associated datasets were not used in the study.

Live imaging

Images were collected using a Nikon Ti2 Eclipse with X-light V2 spinning disc (Nikon, Tokyo, Japan) with a 40X 1.3 NA oil-immersion objective or 60X 1.4 NA oil-immersion objective. Unless otherwise noted samples were imaged pre-wounding, immediately after wounding, every 2 min for 30 min, and then every 10 min for 6 h. Images were pre-processed in NIS-Elements using combinations of background subtraction, rolling ball correction, local contrast, and Denoise a.i. Assembly of figure panels was done using Affinity Designer and frames were centered on the entity in focus, compensating for frame shift due to wounding.

Laser ablation

A single pulse of a 3rd harmonic (355 nm) Q-switched Nd:YAG laser (5 ns pulse-width, Continuum Minilite II, Santa Clara, CA) was used for laser ablation. Laser pulse energies were kept to 1.9 μJ +/- 0.1 μJ, increased from our previous report (O’Connor et al., 2021b) to keep the wound size similar between old and new ablation rig.

Border breakdown and tangential vs radial assignment analysis

Individual border breakdown events were manually observed using FIJI by identifying syncytia late in the video and back-tracking to determine which borders broke down to form each syncytium. These borders were traced back to the first frame after wounding to develop the map on Fig 1G. For each border breakdown event, distance from the center of the wound and time after wounding were recorded in Microsoft Excel. These events were categorized as tangential or radial based on the orientation of the border-to-be-lost relative to a vector pointing outward from the center of the wound. Specifically, the line tool in FIJI was used to measure both the angle of the border with respect to the horizon, θ, and the angle of the line from the center of the wound to the center of the border with respect to the horizon, α. If |cos(α − θ)| ≥ cos 45°, then the border was classified as radial; otherwise, the border was classified as tangential.

Shrinking cell initiation / duration analysis

Shrinking cells were identified in live microscopy videos by beginning at the end of the video and playing backwards in FIJI. Backwards, a shrinking cell appears to bloom from the epithelial layer, characterized by the expansion of a bright puncta of p120ctn-RFP. Each cell that underwent this behavior was marked on a single frame of the video, then the distance from the center of the wound when shrinking started was denoted as well as the time that the shrinking started and completed. All cells that shrank were then manually traced back to the first frame after wounding to develop the map of shrinking cells (Fig 2E).

Approximating nuclei per syncytia

Nuclei and cell borders do not align in Z-projections of the pupal notum as the cells are non-prismatic. To estimate the number of nuclei per syncytium, the pre-wounding density of nuclei per unit area was determined for the circular region where syncytia form after wounding. Next, the apical area of the three largest syncytia/cells around a wound was measured at 0h, 1 h and 2 h post wounding. The area of each syncytia was multiplied by the nuclear density to yield the approximate number of nuclei per syncytia. The number of nuclei in each of the three syncytia was averaged to give a value for each of three samples in Figure 1B.

Measuring apical area of syncytia across varied wound sizes

The three largest syncytia were determined by eye in FIJI for six samples, three ablated at 1.9 µJ and three ablated at 3 µJ ablation. The apical area of syncytia was measured 3 h post wounding using the p120ctn-RFP signal. Initial wound size was calculated by measuring Myosin II marked leading edge when it became apparent 30-160 min post wounding.

Unfused cells at leading edge: count and percent analysis

Unfused cells at the leading edge were identified using FIJI by a lack of border breakdowns. Each unfused cell was manually observed over the duration of wound closure and the time at which it departed from the leading edge was noted. A count of unfused cells at the leading edge was created in Excel and formal graphs were generated using Prism 9. To measure percent of the leading edge comprised of unfused vs. syncytial cells, each unfused cell’s leading edge contact was measured in FIJI using the polygon line tool. The total circumference of the wound was measured using the polygon line tool and unfused cell measurements were subtracted from the total to infer the syncytial occupancy at the leading edge. Formal histograms were generated using Prism 9.

Analyzing GFP labeled cells

107 individually labeled GFP cells were analyzed across 5 wounds over 6.5 h. The position of the ablation was optimized to place as many individually labeled cells within 40-80 µm from the center of the wound as possible. After wounding it was possible to identify a mixing event by the decrease in intensity from the source cell with a corresponding increase in intensity of a previously unlabeled neighbor. Intensity differences made it possible to distinguish instances where two source cells were adjacent to each other but only one had a mixing event. However, large patches of source cells were not evaluated because inter-patch mixing was not distinguishable. To evaluate if border breakdowns were preceded by mixing, the 11 individually labeled cells that had border breakdowns were tracked back to the start of the video and confirmed to have a mixing event. There was never an instance where a labeled source cell had a border breakdown without a prior mixing event occurring.

Wound closure analysis

A pigmented scar forms at the site of laser ablation making identifying the exact moment a wound is closed difficult. Since the scar is approximately the same size in each sample, the time point at which the Myosin II signal disappeared below the scar was used as a proxy for closure.

Fusion analysis

A cell border was counted as a location of wound-induced fusion if it met three criteria: (1) adherens junctions along the cell border were lost; (2) the lost adherens junctions were not restored; and (3) the loss of adherens junctions was accompanied by a morphological change of the neighboring cells, for example, the moving apart of tri-cellular junctions previously connected by the lost cell border. Since border-breakdown fusions are observed most frequently within 1 hr after wounding, we counted all such fusions occurring within 1 hr 40 min of wounding.

When comparing the fusion frequency in pnr versus control domains, we used a landscape-oriented rectangle centered on the wound with its height defined by the inner radius of the zone of nuclear membrane damage visualized by the release of nls-mCherry from the nucleus, which is approximately the same as the zone of lysis (O’Connor et al., 2021a). Replacing the annular region with an extended rectangle excludes possible heterotypic fusions along the pnr-control border just above and below the wound. To calculate the frequency of fusion, in addition to the raw number of fusion borders, the total number of cell borders was also estimated by area-based scaling of the border count within a randomly selected 60-x-60-pixel (16.8-x-16.8-µm) region in the pre-wound image.

Since fusions can also appear as apical shrinking rather than border breakdown, we also quantified the number of apical-shrinking cell fusions in pnr versus control domains. Apical-shrinking cell fusions were defined as diploid cells that decreased in surface area over time and fully disappeared within a 5-hr movie. As with fusion analysis, cells near the pnr-control border were excluded. Apical shrinking cells were counted manually and their spatial locations marked using FIJI’s “Multi-point” tool.

Determining the effect of syncytia on wound closure

The distance between the leading edge and the wound center over time was used to calculate the speed of wound closure. The distance was calculated using the coordinates of the wound center and fixed points on the leading edge. To identify fixed positions on the leading edge that are unbiased and consistent over time, four reference lines, all crossing the center of the wound, were drawn using the red channel with only the nuc-mCherry signal: line 1 along the edge of the pnr domain labeled by nuc-mCherry, line 2 perpendicular to line 1, and line 3 and 4 are 30° from line 2. These references lines were then overlaid on the green channel with the E-cadherin signal.

To determine the speed of wound closure, the intersection of E-cadherin marked leading edge and reference lines 2-4 were used as the position of the leading edge. That is, at each time point, a total of 6 positions of the leading edge were recorded, among which 3 will be in the pnr domain and the other 3 in the ctrl domain. The average distance (d) between the leading edge and the center of the wound were calculated at each time point for both the pnr domain (dpnr) and the control domain (dctrl). The difference between control and pnr distance (Δd = dpnr – dctrl) was calculated over time for both the control pupae with no genetic manipulation in the pnr domain (pnr > +) and experimental animals with Atg1 knocked down in the pnr domain (pnr > Atg1RNAi). To determine whether Δd is statistically different between control and experimental pupae at each time point, two-way ANOVA (fit using full model) was used, comparing each cell mean with the other cell mean in that row.

To determine the distance syncytia covered during wound healing, syncytia were analyzed (1) at the leading edge when wound is half closed and (2) within the bow-tie region between reference line 3 and 4. ROIs were drawn along syncytia of interest using the “Freehand selections” tool of FIJI. In addition to syncytia location when wound is half closed, the starting position of cells that later fuse to form syncytia were also recorded in live-imaging videos, tracking the syncytia back in time. In addition to area, the coordinates of ROIs outlining syncytia at different time points were obtained using a self-written FIJI Macro. Within each ROI, the coordinate closest to the center of the wound was recorded using python, as the location of the syncytia. The distance a certain syncytium moved was calculated by subtracting the distance between the syncytium and the wound center when the wound was half closed from the distance between the syncytium and the wound center immediately after injury.

Calculating intercalation in live imaging

The change in the number of cells when the leading edge forms at 30 min (Nstart) to when the wound is closed (Nend) is equal to the number of intercalations plus the number of radial fusion events. Thus, (intercalations = ΔN – radial fusions). For the same three samples used in Fig 4D,G, we determined ΔN from Nstart at 30 min and Nend when the wound had closed. Radial fusions were tallied by manually observing border breakdown events between the leading-edge cells and intercalations were calculated. Each radial fusion was counted as one prevented intercalation event.

Calculating cell shape index

Images of pupae labeled with Ecad-GFP were segmented into individual cells with Cellpose (Stringer et al., 2021); cells that were not readily segmented were excluded from subsequent analysis. To avoid perimeter inflation due to pixelation, the boundaries of segmented cells were smoothed by taking the midpoint between every other pair of boundary pixels. Smoothing and calculation of perimeters, areas and cell shape indices (P⁄√A) were performed in Mathematica (Wolfram Research Inc., Champaign, IL).

Vertex models with cell fusion

Simulations of wound healing were run using vertex models implemented in Mathematica (Wolfram Research Inc., Champaign, IL) following the methods of Tetley et al. (2019)and using parameters that match experimental observations from the Drosophila pupal notum as detailed in Han et al. (2024) with the following exceptions: the mean line tension on each edge (Λ) was specified in terms of a target shape parameter (ρ0) such that Λ = −2ρ0Γ√A0, where the contractility (Γ) and target cell area (A0) were taken from Han et al. (2024) and ρ0 = 3.745 was chosen so that the actual cell shape index matched experimental pre-wound measurements; and the line tension variability (σm) was decreased from 1.0 x 10-2 to 9.3 x 10-4 so that the intercalation rate in unwounded cell sheets matched experimental observations. Pre-wound configurations were generated by performing a Voronoi tessellation of a circular area with 320 randomly distributed seed points and then equilibrating the cell configuration without stochastic fluctuations until it reached a steady state. Wounds were introduced by removing all cell borders within a circular patch ∼8 cells across and removing the wound’s contributions to the energy Hamiltonian. To handle wound closure and fusion, the following additions were made to the model: (1) purse-string tension on the leading edge of the wound was incorporated as in Tetley et al. (2019) with a normalized strength of 0.115; (2) traction/crawling forces of cells on the leading edge of the wound were incorporated as a uniform tensile stress acting inwardly on the edges and vertices of the wound area (similar to the uniform stress applied to the outer edge of the patch to represent tension from surrounding cells); this inward stress had a normalized strength of Σin⁄KA0 = 0.13, where K is the area elasticity modulus; (3) the cell-cell interfaces at which fusion would occur were selected stochastically with probabilities that depended on time, distance from the wound, and the radial or tangential orientation of the interface – all chosen to match experimental observations reported here; and (4) the fusion of two cells was handled by eliminating the interface between those cells and setting the target area of the fused cell to the sum of the target areas of its two progenitors. All cells, fused or not, had the same contractility and target shape parameter.

Kymograph and plot profile analysis

Actin-GFP intensity was analyzed using the kymograph tool in FIJI after drawing a 11-pixel line through the middle of the syncytia. Profile plot values were exported from NIS elements to Excel and graphs were generated using Prism.